### 2021-08-30 docs/holodeck-paper/src/ms.tex
Start adding boilerplate for holodeck paper

% =================================================================================================
% Holodeck Paper Manuscript
% -------------------------
%
%
%
% =================================================================================================

% general package imports and setup
\input{packages}
% definitions of custom commands
\input{commands}
% journal abbreviations for bibliography
\input{abbrevs}

% =================================================================================================
% ====    Front Matter
% =================================================================================================

\newtotcounter{citnum} % From the package documentation
\def\oldbibitem{} \let\oldbibitem=\bibitem
\def\bibitem{\stepcounter{citnum}\oldbibitem}

\newcommand{\orcidauthorA}{0000-0002-6625-6450} % For author A

\title[Short Title]{Very long and verbose title}
\author[NANOGrav]{the NANOGrav collaboration}
% \author[L.Z.~Kelley et al.]{
%     Luke Zoltan Kelley$^{1}$\thanks{E-mail:lzkelley@northwestern.edu}\orcidA{} \\
%     $^{1}$ \begin{minipage}[t]{\linewidth} Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA), and\\Department of Physics \& Astronomy, Northwestern University, Evanston, IL 60208
%     \end{minipage}
% }

\begin{document}

\maketitle

\begin{abstract}
    abstract
\end{abstract}

\begin{keywords}
    quasars: supermassive black holes
\end{keywords}

% =================================================================================================
% ====    Section 1 - Introduction
% =================================================================================================

\section{Introduction}
    \label{sec:intro}

% =================================================================================================
% ====    Section 2 - Methods
% =================================================================================================

\section{Methods}
    \label{sec:meth}

    Here, the GW strain from a circular binary is \citep[][Eq.~7]{Sesana+2008} (sky and polarization averaged),
        \begin{equation}
        \label{eq:source_strain}
        h_s = \frac{8}{10^{1/2}} \frac{\left(G\mchirp\right)^{5/3}}{c^4 \, d_L}
            \left(2 \pi f_r \right)^{2/3},
        \end{equation}
    for a luminosity distance $d_L$.

    \subsection{Source Distributions and Gravitational Wave Calculations}

        Define the (comoving-) number density of sources as, $n \equiv dN / dV_c$.

        From \citet[][Eqs.~5/8]{Phinney-2001},
        	\begin{align}
        	h_c^2(f) = \int_0^\infty \!\! dz \; \frac{d^2 N}{dz \, d\ln f_r} \; h_s^2\lr{f_r}.
        	\end{align}
        From \citet[][Eq.~6]{Sesana+2008} we can write,
        	\begin{align}
        	\frac{d^2 N}{dz \, d\ln f_r} = \frac{d n_c}{dz} \frac{dz}{dt} \frac{dt}{d\ln f_r} \frac{d V_c}{dz}.
        	\end{align}
        From cosmography \citep[e.g.][]{Hogg-1999},
        	\begin{align}
        	\frac{dz}{dt} = H_0 \lr{1+z} E(z) \\
        	\frac{d V_c}{dz} = 4\pi \frac{c}{H_0} \frac{d_c^2}{E(z)} \\
          d_L = d_c \, (1+z)
        	\end{align}
        Putting this together we have,
        	\begin{align}
        	h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \frac{f_r}{df_r / dt}.
        	\end{align}
        % If this is being calculated from a finite volume (e.g.~simulation), then we can associate the comoving-number-density of sources at a given redshift with the number of binaries per-unit comoving volume, i.e.,
        % 	\begin{align}
        % 	\frac{dn_c}{dz} \, dz = N_\trt{fv}(z)/V_\trt{c,fv}.
        % 	\end{align}

    \subsection{Semi-Analytic Modeling}

        Define the (comoving-) number density of sources as, $n \equiv dN / dV_c$, and the distribution function of sources as $f(M,a,z) = d^2 n(M,a,z) / dM da$.  Here $M$ is the total mass of each systems, $a$ is the binary separation, and $z$ is the redshift.  Note that binary frequency $f$ could be used almost interchangeably with $a$, and that both evolve as a function of time or equivalently redshift.  This formalism can easily be extended to include other parameters, such as mass-ratio and eccentricity ($q$ and $e$).  We can write the conservation equation for binaries as of function of redshift as,
        \begin{equation}
        \label{eq:conservation}
        \diffp{f}{z} + \diffp{}{m} \lrs{f \diffp{m}{z}} + \diffp{}{a} \lrs{f \diffp{a}{z}} = S_f(m, a, z).
        \end{equation}
        Here $S_f$ is a source/sink function that can account for the creation or destruction of binaries.  Here we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts).  Thus we set $S_f = 0$.

        We consider the standard \citep[e.g.~Sesana style; see][]{Chen+2019} semi-analytic model (SAM) formalism of MBH binary populations.  In this style of calculation, $f$ is determined in a region of parameter space that can be observed/estimated, and this is \textit{evolved} to find the distribution in a different region of parameter space that is of interest.  In practice, the observed parameter space is galaxies and galaxy mergers, and the parameter space of interest is close separated MBH binaries that could be GW detectable.  We will express the distribution function as a product of a mass-function, and a pair fraction:
        \begin{equation}
        \label{eq:dist_func}
        f(M,a,z) = \Phi(M, z) \cdot F_a(M, a, z).
        \end{equation}
        In \citet{Chen+2019}, for example, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$f(M, z) = \Phi(M, z) \cdot F(M, z)$ s.t.~$F = \int_{a_0}^{a_1} F_a \, da$.  From $\eqref{eq:conservation}$, we use the chain rule to mix time and redshift evolution, and assume that the mass-change of binaries is negligible, i.e.~$\diffp{m}{t} = 0$, giving:
        \begin{equation}
        \diffp{f}{z} = - \diffp{t}{z} \diffp{}{a} \lrs{f \diffp{a}{t}}.
        \end{equation}
        Plugging-in \eqref{eq:dist_func}:
        \begin{equation}
        \diffp{f}{z} = - \diffp{t}{z} \Phi(M,z) \diffp{}{a} \lrs{F(M,a,z) \diffp{a}{t}}.
        \end{equation}

        The binary population is assumed to be changing only in separation and redshift, which are related by $\partial a / \partial z = (\partial a / \partial t) (\partial t / \partial z)$.  Because the overall number-density is conserved, we can take a finite step in separation and time/redshift, $a\rightarrow a'$ and $z\rightarrow z'$.  Here the time it takes for a binary to go from $a \rightarrow a'$ is $\tau(M,a,z|a')$ which leads to a redshift at the later time of $z' = z'(t + \tau)$.  The standard expression \citep[e.g.~][~Eq.~5]{Chen+2019}, requires making the approximation that,
        \begin{equation}
        \diffp{}{a} \lrs{F(M,a,z) \diffp{a}{t}} \approx \frac{F}{\Delta t} = \frac{F}{\tau(M,a,z|a')}.
        \end{equation}
        Thus giving,
        \begin{equation}
        \diffp{f(M,a',z')}{{z'}} = \diffp{n}{{\log_{10}\!M}{a'}{z'}} = - \diffp{t}{z} \frac{\Phi(M,z) \, F(M,a,z)}{\tau(M,a,z|a')}.
        \end{equation}

    % % Fig -
    % \begin{figure}
    %     \centering
    %     \includegraphics[width=1\columnwidth]{{{example-image-c}}}
    %     \caption{\textbf{header.} body.}
    %     \label{fig:fig1}
    % \end{figure}
    %
    %
    % \begin{equation}\begin{split}
    %     \label{eq:eq1}
    %     2+2 & = 4 \\
    %         & = 10 - 6
    % \end{split}\end{equation}
    %
    % % Fig -
    % \begin{figure*}
	% 	\centering
    %     \subfloat[first]{{
    %         \includegraphics[width=1.75\columnwidth]{{{example-image-a}}}
    %     }}
	% 	\\
    %     % \qquad
    %     \subfloat[second]{{
    %     	\includegraphics[width=1.75\columnwidth]{{{example-image-b}}}
    %     }}
    %     \caption{\textbf{Head.}  Body.}
    %     \label{fig:fig2}
    % \end{figure*}

% =================================================================================================
% ====    Section 3 - Results
% =================================================================================================
\section{Results}
    \label{sec:res}

    % Fig -
    % \begin{figure*}
    %     \centering
    %     \includegraphics[width=1.5\columnwidth]{{{figs/det-nums_fedd-1.00_voff+3.00_dvel+2.00_tobs+0.70}}}
    %     \caption{\textbf{Distributions of parameters for all binaries (black) and those with a detectable kinematic signature in the secondary AGN (colors).}  Less than $10^{-7}$ of primaries are observable, which are not shown.  Secondary AGN with observable velocity offsets ($\voff$) are shown in blue assuming a sensitivity of \mbox{$\voffsens = 10^3 \kmps$}.  Secondaries with observable changing velocities ($\dvel$) are shown in orange (solid) for a sensitivity $\dvelsens = 10^2 \kmps$, and observing baseline of \mbox{$\tobs = 5\, \yr$}.  Systems where both offsets and changes are detectable (orange, dotted) are mostly a uniform subset of the $\dvel$ sample, and are almost indistinguishable in the 2D contour plots.  An optical flux cutoff of \mbox{$6\E{-14} \mathrm{erg/s/cm}^2$} (AB Mag $\approx 21$) is also imposed, where all AGN are assumed to be accreting at an Eddington fraction of $0.1$.}
    %     \label{fig:det_bins_num}
    % \end{figure*}

% =================================================================================================
% ====    Section 4 - Conclusions
% =================================================================================================
\section{Discussion \& Conclusions}
    \label{sec:disc}

% =================================================================================================
% ====    End Matter
% =================================================================================================

% ====    Acknowledgments

\section*{Acknowledgments}
	I am very thankful to.

    This research made use of \astropy, a community-developed core Python package for Astronomy \citep{astropy2013}, in addition to \scipy~\citep{scipy}, \ipython~\citep{ipython}, \jupyter~notebook~\citep{jupyter}, \numpy~\citep{numpy2011} \& \sympy~\citep{sympy2017}.  All figures were generated using \matplotlib~\citep{matplotlib2007}.  Kernel density estimation was performed using the \kalepy{} package (\href{https://github.com/lzkelley/kalepy}{github.com/lzkelley/kalepy}) \citep{kalepy2020}.

    The Illustris data is available online at \href{https://www.illustris-project.org/}{www.illustris-project.org} \citep{Nelson+2015}, and Illustris-TNG data at \href{https://www.tng-project.org/}{www.tng-project.org} \citep{Nelson+2019}.

% ====    Bibliography

\let\oldUrl\url
\renewcommand{\url}[1]{\href{#1}{Link}}

\quad{}
\bibliographystyle{mnras}
\bibliography{refs}

\onecolumn
\clearpage

% ====    Appendices

\appendix

    \section{Additional Material}
        \label{sec:app}

        \noindent\begin{minipage}{\linewidth}

            % Fig -
            % \centering
            % \includegraphics[width=0.55\columnwidth]{{{figs/doan+2019_jitter}}}
            % \captionof{figure}{Distribution of velocity-offset variations (`jitters') from \citep{Doan+2019}, for the red and blue components of the observed BLs.  Measurements are shown as ticks at $y=0$, and KDE distributions are shown using a Scott's factor bandwidth, and Gaussian kernel.  The purple curve is the KDE distribution taking both red and blue data as independent.  The KDE $50\%$ (median) and $99\%$ (Gaussian standard deviation $\sigma \approx 2.3$) jitter for each component are shown with dashed vertical lines and corresponding labels.}
            % \label{fig:jitter_doan2019}
			%
            % \setlength{\tabcolsep}{8pt}
            % \begin{tabular}{p{1.5cm} c c c c c c c}
            %      & & & $5\%$ & $25\%$ & $50\%$ & $75\%$ & $95\%$ \\ [0.5ex]
            %     \hline
            %     \multirow{4}{*}{All}  & \multirow{4}{*}{-}
            %         & $M \; [M_\odot]$ & $3.4 \times10^{ 6 }$ & $7.2 \times10^{ 6 }$ & $1.8 \times10^{ 7 }$ & $6.0 \times10^{ 7 }$ & $4.1 \times10^{ 8 }$ \\
            %         & & q & $2.9 \times10^{ -2 }$ & $1.7 \times10^{ -1 }$ & $4.0 \times10^{ -1 }$ & $6.8 \times10^{ -1 }$ & $9.3 \times10^{ -1 }$ \\
            %         & & z & 0.24 & 0.41 & 0.53 & 0.62 & 0.69 \\
            %         & & $p \; [\mathrm{yr}]$ & $1.4 \times10^{ 2 }$ & $1.5 \times10^{ 3 }$ & $9.6 \times10^{ 3 }$ & $4.0 \times10^{ 4 }$ & $2.0 \times10^{ 5 }$ \\
            %         & & $a \; [\mathrm{pc}]$ & $2.1 \times10^{ -2 }$ & $1.3 \times10^{ -1 }$ & $4.9 \times10^{ -1 }$ & $1.3 \times10^{ 0 }$ & $4.1 \times10^{ 0 }$ \\
            %         & & $a \; [r_g]$ & $1.2 \times10^{ 4 }$ & $1.4 \times10^{ 5 }$ & $4.5 \times10^{ 5 }$ & $1.2 \times10^{ 6 }$ & $4.1 \times10^{ 6 }$ \\
            %     \hline
            %     \multirow{4}{1.5cm}{Secondary\newline Offset\newline($\voff$)} & \multirow{4}{*}{$0.49\%$}
            %         & $M \; [M_\odot]$ & $1.4 \times10^{ 8 }$ & $4.0 \times10^{ 8 }$ & $7.6 \times10^{ 8 }$ & $1.4 \times10^{ 9 }$ & $3.7 \times10^{ 9 }$ \\
            %         & & q & $9.6 \times10^{ -4 }$ & $4.2 \times10^{ -3 }$ & $1.2 \times10^{ -2 }$ & $3.6 \times10^{ -2 }$ & $1.5 \times10^{ -1 }$ \\
            %         & & z & 0.25 & 0.42 & 0.54 & 0.63 & 0.69 \\
            %         & & $p \; [\mathrm{yr}]$ & $8.3 \times10^{ 2 }$ & $1.5 \times10^{ 3 }$ & $2.3 \times10^{ 3 }$ & $3.7 \times10^{ 3 }$ & $9.0 \times10^{ 3 }$ \\
            %         & & $a \; [\mathrm{pc}]$ & 0.22 & 0.41 & 0.58 & 0.87 & 1.82 \\
            %         & & $a \; [r_g]$ & $5.0 \times10^{ 3 }$ & $1.1 \times10^{ 4 }$ & $1.8 \times10^{ 4 }$ & $2.9 \times10^{ 4 }$ & $4.9 \times10^{ 4 }$ \\
            % \end{tabular}
            % \captionof{table}{\textbf{Parameters of detectable binary systems with redshift $z < 0.7$.}  The indicated quantiles are given for total mass ($M$), mass ratio ($q$), separation ($a$), and orbital period ($p$).  In only four binaries ($2\E{-8}$ of systems) is the primary detectable, and those have parameters: \mbox{$M\approx 2$--$5\E{9} \, \msol$},
            % \hspace{0.5ex} \mbox{$q \approx 0.7$--$0.9$}, \hspace{0.5ex} \mbox{$z \approx 0.6$--$0.8$}, \hspace{0.5ex} \mbox{$a \approx 2.2$--$2.8 \, \pc$}.}
            % \label{tab:obs_pars}

        \end{minipage}

    \twocolumn

\end{document}

### 2021-08-30 docs/holodeck-paper/src/abbrevs.tex
Start adding boilerplate for holodeck paper

% \def\apj{ApJ}
% \def\mnras{MNRAS}
% \def\nat{Nat}
% \def\physrevB{Phys. Rev. B}
% \def\prd{Phys. Rev. D}
% \def\araa{ARA\&A}                % "Ann. Rev. Astron. Astrophys."
% \def\aap{A\&A}                   % "Astron. Astrophys."
% \def\aaps{A\&AS}                 % "Astron. Astrophys. Suppl. Ser."
% \def\aj{AJ}                      % "Astron. J."
% \def\apjs{ApJS}                  % "Astrophys. J. Suppl. Ser."
% \def\pasp{PASP}                  % "Publ. Astron. Soc. Pac."
% \def\apjl{ApJ}                   % letter at ApJ
% \def\pasj{PASJ}
% \def\ssr{Space Science Reviews}
% \def\physrep{Physics Reports}
% \def\qjras{Quarterly Journal of the Royal Astronomical Society}

\def\lt{<}

\def\aapr{A\&AR}                                     % Astronomy and Astrophysics Review, the
\def\aj{Astron. J.}              		   		% Astronomical Journal
\def\apj{Astrophys. J.}       		        	 	% Astrophysical Journal
\def\apjl{Astrophys. J. Lett.}             		% Astrophysical Journal, Letters
\def\pasj{PASJ}
\def\physrep{Phys. Rep.}
\def\pasp{PASP}
\def\pasa{PASA}
\def\ssr{Space Science Rev.}			% Space Science Reviews
\def\apjs{Astrophys. J., Suppl. Ser.}            % Astrophysical Journal, Supplement
\def\mnras{Mon. Not. R. Astron. Soc.}        % Monthly Notices of the RAS
\def\prd{Phys. Rev. D}      				% Physical Review D
\def\prl{Phys. Rev. Lett.}   				% Physical Review Letters
\def\cqg{Class. Quant. Grav.}			%Classical and Quantum Gravity
\def\araa{Annu. Rev. Astron. Astrophys.}  % Annual Review of Astron and Astrophys
\def\nat{Nature}              				% Nature
\def\na{Nature Astron.}                     % Nature Astronomy
\def\aap{Astron. Astrophys.}               		% Astronomy and Astrophysics
\def\jasa{J. Am. Stat. Assoc.}
\def\jrssb{J. R. Stat. Soc. B}
\def\aipcs{AIP Conf. Ser.}
\def\jgr{J. Geophys. Res.}                      % Journal of Geophysical Research
\def\sovast{Soviet Astronomy}
\def\planss{Planet.~Space~Sci.}   % Planetary Space Science
\def\memsai{Mem. Societa Astronomica Italiana}

### 2021-08-30 docs/holodeck-paper/src/packages.tex
Start adding boilerplate for holodeck paper

\documentclass[useAMS, usenatbib]{mnras}
\usepackage{color}
\usepackage{booktabs}   % Fancier tables
\usepackage{tabularx}
\usepackage{float}
\usepackage{totcount}
\usepackage{cuted}
% adds colors to colorspace (https://en.wikibooks.org/wiki/LaTeX/Colors)
\usepackage[dvipsnames, table]{xcolor}
\usepackage{graphicx}
\usepackage{multirow}
% Force figures to stay in their sections (tex.stackexchange.com/q/279/22806)
\usepackage[section]{placeins}
\usepackage{subfig}
\usepackage{multicol}
\usepackage{widetext}
\usepackage{upgreek}    % allow alternative greek letters like $\uptau$
\usepackage{enumitem}   % customize enumerate symbols
\setlist[enumerate]{leftmargin=*}   % Set left-margin of enumerate lists to match the edge
\setlist[itemize]{leftmargin=*}   % Set left-margin of enumerate lists to match the edge
\usepackage[T1]{fontenc}

\usepackage{lipsum} % generate dummy text
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[thinc]{esdiff}  % easier/convenient derivatives and partial derivatives

% `txfonts` : Changes math-greek fonts (I like 'sigma'), NOTE: this must be after ams imports
\usepackage{txfonts}

\usepackage{xfrac}

% For ORCID iDs
\usepackage{tikz,xcolor,hyperref}
\definecolor{lime}{HTML}{A6CE39}
\DeclareRobustCommand{\orcidicon}{
	\begin{tikzpicture}
	\draw[lime, fill=lime] (0,0)
	circle [radius=0.16]
	node[white] {{\fontfamily{qag}\selectfont \tiny ID}};
	\draw[white, fill=white] (-0.0625,0.095)
	circle [radius=0.007];
	\end{tikzpicture}
	\hspace{-2mm}
}

\foreach \x in {A, ..., Z}{\expandafter\xdef\csname orcid\x\endcsname{\noexpand\href{https://orcid.org/\csname orcidauthor\x\endcsname}
			{\noexpand\orcidicon}}
}

\graphicspath{ {figs/} }

% shift around the text on the page so looks good on Letter paper
\voffset-.6in
\hoffset0.2in

### 2021-10-16 docs/holodeck-paper/src/ms.tex
Update SAM formalism

@@
 \input{abbrevs}
+
+\newcommand{\thardf}{\tau_f}
 
@@
         	\end{align}
+        The binary hardening timescale is given by the expression,
+        \begin{equation}
+            \thardf \equiv \diffp{t}{{\ln\!f_r}} = \frac{f_r}{\partial f_r / \partial t}
+        \end{equation}
         From cosmography \citep[e.g.][]{Hogg-1999},
@@
         	\begin{align}
-        	h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \frac{f_r}{df_r / dt}.
+        	h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
         	\end{align}
@@
             \label{eq:dist_func}
-        f(M,a,z) = \Phi(M, z) \cdot F_a(M, a, z).
+            f(M,a,z) = \frac{\Phi(M, z)}{M \ln\!10} \cdot F_a(M, a, z),
         \end{equation}
+        where the mass-function is, $\Phi(M, z) \equiv \diffp{n_g}{{\logten{M}}}$.
         In \citet{Chen+2019}, for example, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$f(M, z) = \Phi(M, z) \cdot F(M, z)$ s.t.~$F = \int_{a_0}^{a_1} F_a \, da$.
@@
             \diffp{f(M,a',z')}{{z'}} = \diffp{n}{{\log_{10}\!M}{a'}{z'}} = - \diffp{t}{z} \frac{\Phi(M,z) \, F(M,a,z)}{\tau(M,a,z|a')}.
+        \end{equation}
+
+        Utimately we are interested in a total number of binaries $N$ such that $dN \equiv n \, dV_c$ for a comoving volume $V_c(z)$.
+To find the number of binaries, we use the chain rule to connect the time-evolution of the universe with the evolution of each binary over frequency,
+        \begin{equation}
+            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = \diffp{n}{{M}{q}{z}} \diffp{z}{t} \diffp{t}{ {\ln\!f_r} } \diffp{V_c}{z}.
+        \end{equation}
+        Putting this together, we can finally write,
+        \begin{equation}
+            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = - \Phi(M,z) \, F(M,a,z) \frac{\thardf}{\tau(M,a,z|a')} \diffp{V_c}{z}.
         \end{equation}

### 2021-10-16 holodeck/utils.py
Allow cumulative-summation to be turned on/off in trapz methods.  Add frequency conversion functions.

@@ -346,7 +346,8 @@ def trapz_loglog(
     bounds: Optional[Tuple[float, float]] = None,
     axis: int = -1,
     dlogx: Optional[float] = None,
-    lntol: float = 1e-2
+    lntol: float = 1e-2,
+    cumsum: bool = True,
     ) -> npt.ArrayLike:
     """Calculate integral, given `y = dA/dx` or `y = dA/dlogx` w/ trapezoid rule in log-log space.
 
@@ -446,8 +447,13 @@ def trapz_loglog(
         trapz[idx] = aa[idx] * delta_logx[idx]
 
     # integ = np.log(log_base) * np.cumsum(trapz, axis=axis)
-    integ = np.cumsum(trapz, axis=axis) / np.log(log_base)   # FIX: I think this is divided by base... 2021-10-05
+    # integ = np.cumsum(trapz, axis=axis) / np.log(log_base)   # FIX: I think this is divided by base... 2021-10-05
+    integ = trapz / np.log(log_base)
+    if cumsum:
+        integ = np.cumsum(integ, axis=axis)
     if bounds is not None:
+        if not cumsum:
+            logging.warning(f"WARNING: bounds is not None, but cumsum is False!")
         integ = np.moveaxis(integ, axis, 0)
         lo, hi = integ[ii-1, ...]
         integ = hi - lo
@@ -455,7 +461,7 @@ def trapz_loglog(
     return integ
 
 
-def trapz(yy: npt.ArrayLike, xx: npt.ArrayLike, axis: int = -1):
+def trapz(yy: npt.ArrayLike, xx: npt.ArrayLike, axis: int = -1, cumsum: bool = True):
     """Perform a cumulative integration along the given axis.
 
     Arguments
@@ -486,7 +492,8 @@ def trapz(yy: npt.ArrayLike, xx: npt.ArrayLike, axis: int = -1):
     ct = 0.5 * (ct[1:] + ct[:-1])
     ct = np.moveaxis(ct, 0, -1)
     ct = ct * np.diff(xx)
-    ct = np.cumsum(ct, axis=-1)
+    if cumsum:
+        ct = np.cumsum(ct, axis=-1)
     ct = np.moveaxis(ct, -1, axis)
     return ct
 
@@ -552,12 +559,28 @@ def m1m2_from_mtmr(mt, mr):
     return np.array([m1, m2])
 
 
+def frst_from_fobs(fobs, redz):
[…]

### 2021-10-16 holodeck/sam.py
BUG: fix issue in new calculation of number of binaries from density of binaries --- was performing cumulative sum when it shouldnt have been

@@ -23,7 +23,7 @@ import kalepy as kale
 
 import holodeck as holo
 from holodeck import cosmo, utils, log
-from holodeck.constants import GYR, SPLC, MSOL, MPC
+from holodeck.constants import GYR, SPLC, MSOL, MPC, YR
 
 _AGE_UNIVERSE_GYR = cosmo.age(0.0).to('Gyr').value  # [Gyr]  ~ 13.78
 
@@ -66,10 +66,10 @@ class GSMF_Schechter(_Galaxy_Stellar_Mass_Function):
     """
 
     def __init__(self, phi0=-2.77, phiz=-0.27, mref0=1.74e11*MSOL, mrefz=0.0, alpha0=-1.24, alphaz=-0.03):
-        self._phi0 = phi0         # - 2.77  +/- [-0.29, +0.27]
-        self._phiz = phiz         # - 0.27  +/- [-0.21, +0.23]
+        self._phi0 = phi0         # - 2.77  +/- [-0.29, +0.27]  [1/Mpc^3]
+        self._phiz = phiz         # - 0.27  +/- [-0.21, +0.23]  [1/Mpc^3]
         self._mref0 = mref0       # +11.24  +/- [-0.17, +0.20]  Msol
-        self._mrefz = mrefz
+        self._mrefz = mrefz       #  0.0                        Msol
         self._alpha0 = alpha0     # -1.24   +/- [-0.16, +0.16]
         self._alphaz = alphaz     # -0.03   +/- [-0.14, +0.16]
         return
@@ -335,6 +335,12 @@ class _MMBulge_Relation(abc.ABC):
 
 
 class MMBulge_Simple(_MMBulge_Relation):
+    """Mbh--MBulge relation as a simple power-law.
+
+    M_bh = M_0 * (M_bulge / M_ref)^alpha
+    M_bulge = f_bulge * M_star
+
+    """
 
     def __init__(self, mass_norm=1.48e8*MSOL, mref=1.0e11*MSOL, malpha=1.01, bulge_mfrac=0.615):
         self._mass_norm = mass_norm   # log10(M/Msol) = 8.17  +/-  [-0.32, +0.35]
@@ -433,12 +439,14 @@ class Semi_Analytic_Model:
         shape = [len(ee) for ee in self.edges]
         return tuple(shape)
 
+    '''
     def mass_stellar(self):
         # total-mass, mass-ratio ==> (M1, M2)
[…]

### 2021-10-21 holodeck/sam.py
Update SAM sampling to use new dens/mass parameters from kalepy sampling;update GW calculations appropriately

@@ -1,22 +1,27 @@
-"""
+"""Holodeck - Semi Analytic Modeling submodule
 
-Chen, Sesana, Conselice 2019 = [Chen19]
-Constraining astrophysical observables of galaxy and supermassive black hole binary mergers using pulsar timing arrays
-https://ui.adsabs.harvard.edu/abs/2019MNRAS.488..401C/abstract
-
+References
+----------
+-   Chen, Sesana, Conselice 2019 = [Chen19]
+    https://ui.adsabs.harvard.edu/abs/2019MNRAS.488..401C/abstract
+    Constraining astrophysical observables of galaxy and supermassive black hole binary mergers
+        using pulsar timing arrays
 
 To-Do
 -----
-* [ ] Check that _GW_ frequencies and _orbital_ frequencies are being used in the correct places.
+*[ ]Check that _GW_ frequencies and _orbital_ frequencies are being used in the correct places.
     Check `number_at_gw_fobs` and related methods.
-* [ ] Expand SAM mass-ratios to wider range, change to log-space.
+*[ ]Expand SAM mass-ratios to wider range, change to log-space.
+*[ ]Incorporate arbitrary hardening mechanisms into SAM construction, sample self-consistently.
+*[ ]When using `sample_outliers` check whether the density (used for intrabin sampling) should be
+    the log(dens) instead of just `dens`.
 
 """
 
 import abc
 import inspect
 
-# import numba
+import numba
 import numpy as np
 
 import kalepy as kale
@@ -530,8 +535,10 @@ class Semi_Analytic_Model:
 
         Returns
         -------
+        edges : (4,) list of 1darrays of scalar
+            A list containing the edges along each dimension.
         number : (M, Q, Z, F) ndarray of scalar
-
+            Number density of binaries.
 
         Notes
         -----
@@ -629,11 +636,25 @@ class Semi_Analytic_Model:
 
         return edges, number
 
-    def gwb(self, fobs, hard=holo.evolution.Hard_GW, realize=False, **kwargs):
-        """
+    def gwb(self, fobs, hard=holo.evolution.Hard_GW, realize=False):
+        """Calculate the (smooth/semi-analytic) GWB at the given observed frequencies.
+
         Arguments
         ---------
-        fobs : units of [1/s]
+        fobs : (F,) array_like of scalar, [1/sec]
[…]

### 2021-10-21 holodeck/utils.py
Cleanup

@@ -1,6 +1,7 @@
 """
 
-References:
+References
+----------
 - Peters-1964 : [Peters 1964](https://ui.adsabs.harvard.edu/abs/1964PhRv..136.1224P/abstract)
 - EN07 : [Enoki & Nagashima 2007](https://ui.adsabs.harvard.edu/abs/2007PThPh.117..241E/abstract)
 - Enoki+2004 : [Enoki et al. 2004](https://ui.adsabs.harvard.edu/abs/2004ApJ...615...19E/abstract)
@@ -734,19 +735,6 @@ def gw_lum_circ(mchirp, freq_orb_rest):
     """
     lgw_circ = _GW_LUM_CONST * np.power(2.0*np.pi*freq_orb_rest*mchirp, 10.0/3.0)
     return lgw_circ
-
-
-'''
-def gw_strain_source(mchirp, dlum, freq_orb_rest):
-    """GW Strain from a single source in a circular orbit.
-
-    e.g. Sesana+2004 Eq.36
-    e.g. EN07 Eq.17
-    """
-    #
-    hs = _GW_SRC_CONST * mchirp * np.power(2*mchirp*freq_orb_rest, 2/3) / dlum
-    return hs
-'''
 
 
 def gw_strain_source(mchirp, dcom, freq_orb_rest):

### 2021-10-21 holodeck/sam.py
Cleanup

@@ -800,7 +800,6 @@ def _gws_from_samples(vals, weights, fobs):
     return gff, gwf, gwb
 
 
-'''
 def sampled_gws_from_sam(sam, fobs, hard=holo.evolution.Hard_GW, **kwargs):
     """
 
@@ -810,10 +809,9 @@ def sampled_gws_from_sam(sam, fobs, hard=holo.evolution.Hard_GW, **kwargs):
         Target frequencies of interest in units of [1/yr]
 
     """
-    vals, weights = sample_sam_with_hardening(sam, hard, fobs=fobs, **kwargs)
+    vals, weights, edges, dens = sample_sam_with_hardening(sam, hard, fobs=fobs, **kwargs)
     gff, gwf, gwb = _gws_from_samples(vals, weights, fobs)
     return gff, gwf, gwb
-'''
 
 
 @numba.njit

### 2021-10-21 holodeck/sam.py
BUG: f-string for py3.7 compat

@@ -11,7 +11,7 @@ To-Do
 -----
 *[ ]Check that _GW_ frequencies and _orbital_ frequencies are being used in the correct places.
     Check `number_at_gw_fobs` and related methods.
-*[ ]Expand SAM mass-ratios to wider range, change to log-space.
+*[ ]Change mass-ratios and redshifts (1+z) to log-space; expand q parameter range.
 *[ ]Incorporate arbitrary hardening mechanisms into SAM construction, sample self-consistently.
 *[ ]When using `sample_outliers` check whether the density (used for intrabin sampling) should be
     the log(dens) instead of just `dens`.

### 2022-03-24 holodeck/observations.py
Move scaling-relationships to 'relations.py', implement updated, cleaned up MMBulge relationship classes

@@ -2,47 +2,22 @@
 
 To-Do
 -----
-* [ ] Pass concentration-relation (or other method to calculate) to NFW classes on instantiation
-
-
-References
-----------
-* [NFW-97] : Navarro, Frenk & White 1997
-    A Universal Density Profile from Hierarchical Clustering
-    https://ui.adsabs.harvard.edu/abs/1997ApJ...490..493N/abstract
-
-* [Guo-2010] : Guo, White, Li & Boylan-Kolchin 2010
-    How do galaxies populate dark matter haloes?
-    https://ui.adsabs.harvard.edu/abs/2010MNRAS.404.1111G/abstract
-
-* [Kormendy+2013] : Kormendy & Ho 2013
-    Coevolution (Or Not) of Supermassive Black Holes and Host Galaxies
-    https://ui.adsabs.harvard.edu/abs/2013ARA%26A..51..511K/abstract
-
-* [McConnell+2013] : McConnell & Ma 2013
-    Revisiting the Scaling Relations of Black Hole Masses and Host Galaxy Properties
-    https://ui.adsabs.harvard.edu/abs/2013ApJ...764..184M/abstract
-
-* [Behroozi+2013] : Behroozi, Wechsler & Conroy 2013
-    The Average Star Formation Histories of Galaxies in Dark Matter Halos from z = 0-8
-    https://ui.adsabs.harvard.edu/abs/2013ApJ...770...57B/abstract
-
-* [Klypin+2016] : Klypin et al. 2016
-    MultiDark simulations: the story of dark matter halo concentrations and density profiles
-    https://ui.adsabs.harvard.edu/abs/2016MNRAS.457.4340K/abstract
 
 """
-
-import abc
+import os
 
 import numpy as np
-import scipy as sp
+# import scipy as sp
 
-from holodeck import log, cosmo, utils
-from holodeck.constants import MSOL, NWTG
+import holodeck as holo
+# from holodeck import log, cosmo, utils
+# from holodeck.constants import MSOL, NWTG
+
+FNAME_MM13 = "mcconnell+ma-2013_1211.2816.txt"
 
 
-def load_mcconnell_ma_2013(fname):
+def load_mcconnell_ma_2013():
+    fname = os.path.join(holo._PATH_DATA, FNAME_MM13)
     header = []
     data_raw = []
     hcnt = 0
@@ -81,612 +56,3 @@ def load_mcconnell_ma_2013(fname):
     data['reffv'] = np.array([dr[17] for dr in data_raw])
 
     return data
-
-
[…]

### 2022-03-24 holodeck/relations.py
Move scaling-relationships to 'relations.py', implement updated, cleaned up MMBulge relationship classes

@@ -2,28 +2,53 @@
 
 References
 ----------
--
+-   Kormendy+Ho-2013 (1304.7762) = [KH13]
+    Coevolution (Or Not) of Supermassive Black Holes and Host Galaxies
+    https://ui.adsabs.harvard.edu/abs/2013ARA%26A..51..511K/abstract
+
+-   McConnell+Ma-2013 (1211.2816) = [MM13]
+    Revisiting the Scaling Relations of Black Hole Masses and Host Galaxy Properties
+    https://ui.adsabs.harvard.edu/abs/2013ApJ...764..184M/abstract
+
+* [NFW-97] : Navarro, Frenk & White 1997
+    A Universal Density Profile from Hierarchical Clustering
+    https://ui.adsabs.harvard.edu/abs/1997ApJ...490..493N/abstract
+
+* [Guo-2010] : Guo, White, Li & Boylan-Kolchin 2010
+    How do galaxies populate dark matter haloes?
+    https://ui.adsabs.harvard.edu/abs/2010MNRAS.404.1111G/abstract
+
+* [Behroozi+2013] : Behroozi, Wechsler & Conroy 2013
+    The Average Star Formation Histories of Galaxies in Dark Matter Halos from z = 0-8
+    https://ui.adsabs.harvard.edu/abs/2013ApJ...770...57B/abstract
+
+* [Klypin+2016] : Klypin et al. 2016
+    MultiDark simulations: the story of dark matter halo concentrations and density profiles
+    https://ui.adsabs.harvard.edu/abs/2016MNRAS.457.4340K/abstract
+
 
 To-Do
 -----
+*[ ]Pass concentration-relation (or other method to calculate) to NFW classes on instantiation
 *[ ]
 
 """
 
 import abc
-# import inspect
 
-# import numba
 import numpy as np
+import scipy as sp
 
 # import kalepy as kale
 
-import holodeck as holo
+# import holodeck as holo
 from holodeck import cosmo, utils, log
-from holodeck.constants import GYR, SPLC, MSOL, MPC
+from holodeck.constants import MSOL, NWTG
 
 
 class _MMBulge_Relation(abc.ABC):
+
+    # ---- Abstract Methods : must be overridden in subclasses  ----
 
     @abc.abstractmethod
     def __init__(self, *args, **kwargs):
@@ -49,6 +74,8 @@ class _MMBulge_Relation(abc.ABC):
     def dmbulge_dmstar(self, mstar):
         return
 
+    # ---- Internal Methods ----
+
[…]

### 2022-03-24 holodeck/sam.py

@@ -29,6 +29,7 @@ import kalepy as kale
 import holodeck as holo
 from holodeck import cosmo, utils, log
 from holodeck.constants import GYR, SPLC, MSOL, MPC
+from holodeck import relations
 
 _AGE_UNIVERSE_GYR = cosmo.age(0.0).to('Gyr').value  # [Gyr]  ~ 13.78
 
@@ -305,7 +306,7 @@ class Semi_Analytic_Model:
 
     def __init__(
         self, mtot=[2.75e5*MSOL, 1.0e11*MSOL, 46], mrat=[0.02, 1.0, 50], redz=[0.0, 6.0, 61],
-        gsmf=GSMF_Schechter, gpf=GPF_Power_Law, gmt=GMT_Power_Law, mmbulge=MMBulge_Simple
+        gsmf=GSMF_Schechter, gpf=GPF_Power_Law, gmt=GMT_Power_Law, mmbulge=relations.MMBulge_Standard
     ):
 
         if inspect.isclass(gsmf):
@@ -325,7 +326,7 @@ class Semi_Analytic_Model:
 
         if inspect.isclass(mmbulge):
             mmbulge = mmbulge()
-        elif not isinstance(mmbulge, _MMBulge_Relation):
+        elif not isinstance(mmbulge, relations._MMBulge_Relation):
             raise ValueError("`mmbulge` must be an instance or subclass of `_MMBulge_Relation`!")
 
         # NOTE: the spacing (log vs lin) is important.  e.g. in integrating from differential-number to (total) number

### 2022-03-25 holodeck/sam.py
DOCS: to-do item added

@@ -15,6 +15,8 @@ To-Do
 *[ ]Incorporate arbitrary hardening mechanisms into SAM construction, sample self-consistently.
 *[ ]When using `sample_outliers` check whether the density (used for intrabin sampling) should be
     the log(dens) instead of just `dens`.
+*[ ]Should there be an additional dimension of parameter space for galaxy properties?  This way
+    variance in scaling relations is incorporated directly before calculating realizations.
 
 """
 

### 2022-03-25 docs/holodeck-paper/src/ms.tex
MS: manuscript updates

@@
 \newcommand{\thardf}{\tau_f}
+\newcommand{\holodeck}{\texttt{holodeck}}
+\newcommand{\python}{\texttt{python}}
+\newcommand{\mmbulge}{{$M_\textrm{BH}$--$M_\textrm{bulge}$}}
 
@@
         \begin{equation}
+            \label{eq:sam_final}
             \diffp{N}{{M}{q}{z}{\ln\!f_r}} = - \Phi(M,z) \, F(M,a,z) \frac{\thardf}{\tau(M,a,z|a')} \diffp{V_c}{z}.
@@
     % \end{figure*}
+
+\section{Implementation}
+    \label{sec:imp}
+
+    The \holodeck{} framework is written primarily in \python{} using a class-based structure for modularity and extensibility.
+
+    \subsection{Semi-Analytic Models}
+        \label{sec:imp_sam}
+
+        The core of the semi-analytic modeling framework in \holodeck{} is implementing \eqref{eq:sam_final}.
+There are classes for each of the key components: The galaxy stellar-mass function (GSMF, $\Phi$), the galaxy pair fraction ($F$), the merger timescale ($\tau$), and additional a scaling relation for mapping host galaxies via to MBH properties --- typically via an \mmbulge{} relation.
+Each component class is implemented generically via basic API methods (e.g.~a GSMF that returns the number density of galaxies at a given stellar-mass and redshift), and can be subclassed to provide a particular implementation (e.g.~an analytic Schechter-function GSMF).
+The number (or number density) of binaries is specified over a grid of parameters, typically in a four dimensional space of total mass ($M$), mass ratio ($q$), redshift ($z$), and binary rest-frame orbital frequency ($f_r$).
+
+        The semi-analytic model formalism defines a continuous distribution of MBH binaries.
+For many calculations, the discrete nature of MBHB sources is important, not only for examining individual continuous wave sources but also because of the important effects of discrete sources on the shape of the GWB due to the non-linear scaling of GW strain \citep{Sesana+2008}.
+Additional, a component of the uncertainty in expected GW signals is due to `cosmic variance', which can be treated as Poisson noise when creating multiple samples (or `realizations') of our binary populations.
+To handle discretization and resampling we use the \kalepy{} package \citep{kalepy2021}, in particular, a method of outlier sampling to capture regions of parameter space where finite number effects are important.
+The \holodeck{} semi-analytic models specify the number (or number-density) of binaries over a grid, i.e.~$N_{ijkl}$.
+By definition, grid cells which have an expected total number of binaries $\gg 1$ will not produce appreciable finite-number effects, so instead they are given a weighting based on a Poisson distribution centered around their expectation value, i.e.~$\mathcal{P}(N_{ijkl})$.
+Grid cells with expected numbers of order unity or less do produce important finite-number effects, and thus are directly resampled
 

### 2022-03-25 holodeck/observations.py
Tweaks

@@ -1,4 +1,4 @@
-"""
+"""Observational Data
 
 To-Do
 -----

### 2022-03-25 holodeck/relations.py

@@ -101,12 +101,12 @@ class MMBulge_Standard(_MMBulge_Relation):
     """
     """
 
-    MASS_AMP = np.nan
-    MASS_REF = 1.0
-    MASS_PLAW = np.nan
-    SCATTER_DEX = np.nan
+    MASS_AMP = 3.0e8 * MSOL
+    MASS_PLAW = 1.0
+    MASS_REF = 1.0e11 * MSOL
+    SCATTER_DEX = 0.0
 
-    def __init__(self, mamp=None, mplaw=None, mref=None, bulge_mfrac=0.615, scatter_dex=None, scatter=None):
+    def __init__(self, mamp=None, mplaw=None, mref=None, bulge_mfrac=0.615, scatter_dex=None, scatter=False):
         if mamp is None:
             mamp = self.MASS_AMP
         if mplaw is None:
@@ -119,6 +119,7 @@ class MMBulge_Standard(_MMBulge_Relation):
         if scatter is True:
             pass
         elif scatter is False:
+            log.debug(f"Setting ``scatter_dex=None`` in {self.__class__} (arg ``scatter={scatter}``)")
             scatter_dex = None
         else:
             utils.error("`scatter` must be either True or False!")

### 2022-03-25 holodeck/sam.py

@@ -306,7 +306,9 @@ class Semi_Analytic_Model:
 
     def __init__(
         self, mtot=[2.75e5*MSOL, 1.0e11*MSOL, 46], mrat=[0.02, 1.0, 50], redz=[0.0, 6.0, 61],
-        gsmf=GSMF_Schechter, gpf=GPF_Power_Law, gmt=GMT_Power_Law, mmbulge=relations.MMBulge_Standard
+        # self, mtot=[2.75e5*MSOL, 1.0e11*MSOL, 11], mrat=[0.02, 1.0, 10], redz=[0.0, 6.0, 9],
+        shape=None,
+        gsmf=GSMF_Schechter, gpf=GPF_Power_Law, gmt=GMT_Power_Law, mmbulge=relations.MMBulge_MM13
     ):
 
         if inspect.isclass(gsmf):
@@ -328,6 +330,14 @@ class Semi_Analytic_Model:
             mmbulge = mmbulge()
         elif not isinstance(mmbulge, relations._MMBulge_Relation):
             raise ValueError("`mmbulge` must be an instance or subclass of `_MMBulge_Relation`!")
+
+        if shape is not None:
+            if len(shape) == 3:
+                mtot[2] = shape[0]
+                mrat[2] = shape[1]
+                redz[2] = shape[2]
+            else:
+                raise
 
         # NOTE: the spacing (log vs lin) is important.  e.g. in integrating from differential-number to (total) number
         self.mtot = np.logspace(*np.log10(mtot[:2]), mtot[2])
@@ -607,17 +617,12 @@ class Semi_Analytic_Model:
         # get the correct shape for frequencies
         # NOTE: coms[:, :, :, i] == coms[0, 0, 0, i]  i.e. only last dimension varies
         coms[-1] = coms[-1][1:, 1:, 1:, :]
-
-        # shape_grid = coms[0].shape
         coms = [cc.flat[:] for cc in coms]   # use `[:]` to avoid issues with flatiter instance
 
         # ---- calculate GW strain at bin centroids
         mc = utils.chirp_mass(*utils.m1m2_from_mtmr(coms[0], coms[1]))
-        print(f"shape coms[2] = {np.shape(coms[2])}", type(coms[2]))
         dc = cosmo.comoving_distance(coms[2]).cgs.value
         fr = utils.frst_from_fobs(coms[3], coms[2])
[…]

### 2022-03-28 docs/holodeck-paper/src/ms.tex
manuscript updates

@@
 \input{abbrevs}
-
-\newcommand{\thardf}{\tau_f}
-\newcommand{\holodeck}{\texttt{holodeck}}
-\newcommand{\python}{\texttt{python}}
-\newcommand{\mmbulge}{{$M_\textrm{BH}$--$M_\textrm{bulge}$}}
 
@@
 
-    Here, the GW strain from a circular binary is \citep[][Eq.~7]{Sesana+2008} (sky and polarization averaged),
-        \begin{equation}
-        \label{eq:source_strain}
-        h_s = \frac{8}{10^{1/2}} \frac{\left(G\mchirp\right)^{5/3}}{c^4 \, d_L}
-            \left(2 \pi f_r \right)^{2/3},
-        \end{equation}
-    for a luminosity distance $d_L$.
-
     \subsection{Source Distributions and Gravitational Wave Calculations}
 
-        Define the (comoving-) number density of sources as, $n \equiv dN / dV_c$.
-
-        From \citet[][Eqs.~5/8]{Phinney-2001},
+        Let $N(z)$ denote the number of binaries out to a redshift $z$, corresponding to a comoving volume $V_c$.
+        We can calculate the characteristic strain from an ensemble of GW sources as \citep[][Eqs.~5/8]{Phinney-2001},
         	\begin{align}
@@
         	\end{align}
-        From \citet[][Eq.~6]{Sesana+2008} we can write,
+        In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$.  These can be related as \citep[][Eq.~6]{Sesana+2008},
         	\begin{align}
@@
         	\end{align}
-        The binary hardening timescale is given by the expression,
+        Note that the total number of binaries in a given frequency interval depends not only on the number density, but also their rate of frequency-evolution.
+The faster binaries evolve through a given frequency interval, the fewer binaries will occupy that interval on average.
+This is typically described by the binary `hardening' timescale (with respect to frequency), given by the expression,
         \begin{equation}
-            \thardf \equiv \diffp{t}{{\ln\!f_r}} = \frac{f_r}{\partial f_r / \partial t}
+            \thardf \equiv \diffp{t}{{\ln\!f_r}} = \frac{f_r}{\partial f_r / \partial t}.
         \end{equation}
-        From cosmography \citep[e.g.][]{Hogg-1999},
-        	\begin{align}
-        	\frac{dz}{dt} = H_0 \lr{1+z} E(z) \\
-        	\frac{d V_c}{dz} = 4\pi \frac{c}{H_0} \frac{d_c^2}{E(z)} \\
-          d_L = d_c \, (1+z)
-        	\end{align}
-        Putting this together we have,
+        Plugging in the relevant cosmographic relationships (see \secref{sec:app_eqs}) we have,
         	\begin{align}
@@
         	\end{align}
-        % If this is being calculated from a finite volume (e.g.~simulation), then we can associate the comoving-number-density of sources at a given redshift with the number of binaries per-unit comoving volume, i.e.,
-        % 	\begin{align}
-        % 	\frac{dn_c}{dz} \, dz = N_\trt{fv}(z)/V_\trt{c,fv}.
-        % 	\end{align}
+        The challenge in modeling astrophysical sources of GWs is in calculating their number density.  Numerous approaches are possible, described in the following sections.
 
@@
 
-        Define the (comoving-) number density of sources as, $n \equiv dN / dV_c$, and the distribution function of sources as $f(M,a,z) = d^2 n(M,a,z) / dM da$.
+        Define the distribution function of sources as $f(M,a,z) = d^2 n(M,a,z) / dM da$.
 Here $M$ is the total mass of each systems, $a$ is the binary separation, and $z$ is the redshift.
-Note that binary frequency $f$ could be used almost interchangeably with $a$, and that both evolve as a function of time or equivalently redshift.
+Note that binary frequency $f_r$ could be used almost interchangeably with $a$, and that both evolve as a function of time or equivalently redshift.
 This formalism can easily be extended to include other parameters, such as mass-ratio and eccentricity ($q$ and $e$).
@@
         The core of the semi-analytic modeling framework in \holodeck{} is implementing \eqref{eq:sam_final}.
-There are classes for each of the key components: The galaxy stellar-mass function (GSMF, $\Phi$), the galaxy pair fraction ($F$), the merger timescale ($\tau$), and additional a scaling relation for mapping host galaxies via to MBH properties --- typically via an \mmbulge{} relation.
+There are classes for each of the key components: The galaxy stellar-mass function (GSMF, $\Phi$), the galaxy pair fraction \(F\), the merger timescale ($\tau$), and additional a scaling relation for mapping host galaxies via to MBH properties --- typically via an \mmbulge{} relation.
 Each component class is implemented generically via basic API methods (e.g.~a GSMF that returns the number density of galaxies at a given stellar-mass and redshift), and can be subclassed to provide a particular implementation (e.g.~an analytic Schechter-function GSMF).
@@
 Additional, a component of the uncertainty in expected GW signals is due to `cosmic variance', which can be treated as Poisson noise when creating multiple samples (or `realizations') of our binary populations.
-To handle discretization and resampling we use the \kalepy{} package \citep{kalepy2021}, in particular, a method of outlier sampling to capture regions of parameter space where finite number effects are important.
+To handle discretization and resampling we use the \texttt{kalepy} package \citep{kalepy2021}, in particular, a method of outlier sampling to capture regions of parameter space where finite number effects are important.
 The \holodeck{} semi-analytic models specify the number (or number-density) of binaries over a grid, i.e.~$N_{ijkl}$.
@@
 
-    This research made use of \astropy, a community-developed core Python package for Astronomy \citep{astropy2013}, in addition to \scipy~\citep{scipy}, \ipython~\citep{ipython}, \jupyter~notebook~\citep{jupyter}, \numpy~\citep{numpy2011} \& \sympy~\citep{sympy2017}.
-All figures were generated using \matplotlib~\citep{matplotlib2007}.
-Kernel density estimation was performed using the \kalepy{} package (\href{https://github.com/lzkelley/kalepy}{github.com/lzkelley/kalepy}) \citep{kalepy2020}.
+    This research made use of \texttt{astropy}, a community-developed core Python package for Astronomy \citep{astropy2013}, in addition to \texttt{Scipy}~\citep{scipy}, \texttt{ipython}~\citep{ipython}, \texttt{jupyter}~notebook~\citep{jupyter}, \texttt{Numpy}~\citep{numpy2011} \& \texttt{SymPy}~\citep{sympy2017}.
+All figures were generated using \texttt{matplotlib}~\citep{matplotlib2007}.
+Kernel density estimation was performed using the \texttt{kalepy}{} package (\href{https://github.com/lzkelley/kalepy}{github.com/lzkelley/kalepy}) \citep{kalepy2020}.
 
@@
 \appendix
+
+    \section{Additional Equations}
+        \label{sec:app_eqs}
+        A binary is described by its total mass $M$, mass ratio $q \equiv m_2 / m_1$ (s.t.~$q\leq1$), (proper) semi-major axis $a$, and rest-frame orbital frequency $f_r$.
+The characteristic gravitational wave mass is the chirp mass, $\mchirp \equiv M \, q^{3/5} \, \lr[-6/5]{1+q}$.
+
+        The GW strain from a circular binary is \citep[][Eq.~7; sky and polarization averaged]{Sesana+2008},
+            \begin{equation}
+            \label{eq:source_strain}
+            h_s = \frac{8}{10^{1/2}} \frac{\left(G\mchirp\right)^{5/3}}{c^4 \, d_L}
+                \left(2 \pi f_r \right)^{2/3},
+            \end{equation}
+        for a luminosity distance $d_L$.
+Recall that the luminosity distance and comoving distance are related as $d_L = d_c \, (1+z)$, that the comoving volume of the Universe is given by \citep{Hogg-1999},
+        \begin{equation}
+            \frac{d V_c}{dz} = 4\pi \frac{c}{H_0} \frac{d_c^2}{E(z)},
+        \end{equation}
+        and that redshift can be related to time (or age of the universe) as,
+        \begin{align}
+            \frac{dz}{dt} = H_0 \lr{1+z} E(z).
+        \end{align}
+        The evolution of the Hubble parameter over redshift can be calculated from the matter-energy constituents of the universe as,
+        \begin{align}
+            H(z) & = H_0(z) E(z), \\
+            E(z) & \equiv \lrs[1/2]{\Omega_R \lr[4]{1+z} + \Omega_m \lr[3]{1+z} + \Omega_k \lr[2]{1+z} + \Omega_\Lambda}.
+        \end{align}
+        The $\Omega$ terms are the usual redshift-zero fractions of the critical density for radiation, non-relativistic matter, curvature, and dark energy.
+We adopt the results of WMAP9 \citep{Hinshaw+2013} for our fiducial cosmology, assuming zero curvature: $H_0 = 69.33 \textrm{ km/s/Mpc}$, $\Omega_m = 0.2880$, $\Omega_\Lambda = 0.7120$.
 

### 2022-03-28 holodeck/observations.py
--amend

@@ -7,11 +7,8 @@ To-Do
 import os
 
 import numpy as np
-# import scipy as sp
 
 import holodeck as holo
-# from holodeck import log, cosmo, utils
-# from holodeck.constants import MSOL, NWTG
 
 FNAME_MM13 = "mcconnell+ma-2013_1211.2816.txt"
 

### 2022-03-28 holodeck/cosmology.py
--amend

@@ -1,4 +1,9 @@
-"""
+"""Cosmology related code.
+
+[WMAP9] = Hinshaw et al. 2013 (1212.5226)
+    Nine-year Wilkinson Microwave Anisotropy Probe (WMAP) Observations: Cosmological Parameter Results
+    https://ui.adsabs.harvard.edu/abs/2013ApJS..208...19H/abstract
+
 """
 
 import astropy as ap
@@ -20,17 +25,13 @@ class Cosmology(ap.cosmology.FlatLambdaCDM):
 
     API Methods
     -----------
-    -   tage_to_z                      -   Convert from universe-age to redshift.
-    -   dcom_to_z                      -   Convert from comoving-dist to redshift.
-    -   dlum_to_z                      -   Convert from luminosity-dist to redshift.
-    -   get_grid                       -   Retrieve the underlying interpolation grid.
 
     """
 
-    # These are WMAP9 parameters
-    Omega0 = 0.286
-    OmegaBaryon = 0.0463
-    HubbleParam = 0.693
+    # These are WMAP9 parameters, see [WMAP9], Table 3, WMAP+BAO+H0
+    Omega0 = 0.2880
+    OmegaBaryon = 0.0472
+    HubbleParam = 0.6933
     _H0_sub = HubbleParam * 100.0  # NOTE: this *cannot* be `H0` or `_H0` --- conflicts with astropy internals
     SPLC = 29979245800.0
 

### 2022-03-28 setup.py
Bump to version 0.2

@@ -43,6 +43,6 @@ setup(
 
     # Additional entries you may want simply uncomment the lines you want and fill in the data
     # url='http://www.my_package.com',  # Website
-    python_requires=">=3.5",          # Python version restrictions
+    python_requires=">=3.7",          # Python version restrictions
 
 )

### 2022-03-29 docs/holodeck-paper/src/ms.tex
maint

@@
         	\begin{align}
+                \label{eq:num_num_dens}
             	\frac{d^2 N}{dz \, d\ln f_r} = \frac{d n_c}{dz} \frac{dz}{dt} \frac{dt}{d\ln f_r} \frac{d V_c}{dz}.
@@
 
-        Define the distribution function of sources as $f(M,a,z) = d^2 n(M,a,z) / dM da$.
+        Define the distribution function of sources as $f(M,q,a,z) = d^3 n(M,q,a,z) / dM dq da$.
 Here $M$ is the total mass of each systems, $a$ is the binary separation, and $z$ is the redshift.
@@
             \label{eq:conservation}
-            \diffp{f}{z} + \diffp{}{m} \lrs{f \diffp{m}{z}} + \diffp{}{a} \lrs{f \diffp{a}{z}} = S_f(m, a, z).
+            \diffp{f}{z} +
+                \diffp{}{M} \lrs{f \diffp{M}{z}} +
+                \diffp{}{q} \lrs{f \diffp{q}{z}} +
+                \diffp{}{a} \lrs{f \diffp{a}{z}} = S_f(M, q, a, z).
         \end{equation}
         Here $S_f$ is a source/sink function that can account for the creation or destruction of binaries.
-Here we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts).
-Thus we set $S_f = 0$.
 
@@
 In practice, the observed parameter space is galaxies and galaxy mergers, and the parameter space of interest is close separated MBH binaries that could be GW detectable.
-We will express the distribution function as a product of a mass-function, and a pair fraction:
+\note{Thus we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts), and we set $S_f = 0$.}
+
+        We will express the distribution function as a product of a mass function, and a pair fraction:
         \begin{equation}
             \label{eq:dist_func}
-            f(M,a,z) = \frac{\Phi(M, z)}{M \ln\!10} \cdot F_a(M, a, z),
+            f(M,q,a,z) = \frac{\Phi(M, z)}{M \ln\!10} \cdot F_a(M,q,a,z),
         \end{equation}
-        where the mass-function is, $\Phi(M, z) \equiv \diffp{n_g}{{\logten{M}}}$.
-        In \citet{Chen+2019}, for example, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$f(M, z) = \Phi(M, z) \cdot F(M, z)$ s.t.~$F = \int_{a_0}^{a_1} F_a \, da$.
-From $\eqref{eq:conservation}$, we use the chain rule to mix time and redshift evolution, and assume that the mass-change of binaries is negligible, i.e.~$\diffp{m}{t} = 0$, giving:
+        where the mass function, $\Phi(M, q, z) \equiv \diffp{n_g}{{\logten{M}}}$, is calculated based on the number density of galaxies ($n_g$).
+We assume that there is a one-to-one mapping from galaxy mass to MBH mass, such that the galaxy mass-function can still be used to uniquely define the mass distribution of MBHs.
+Typically the MBH--galaxy relation is given in terms of an `$M_\trt{BH}$--$M_\trt{bulge}$' relation \citep[e.g.][]{McConnell+Ma-2013}, which is a observationally derived relation between the mass of MBHs and the mass of the stellar bulge component of their host galaxy.
+        In \citet{Chen+2019}, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$f(M,q,z) = \Phi(M, z) \cdot F(M, q, z)$ s.t.~$F = \int_{a_0}^{a_1} F_a \, da$.
+From $\eqref{eq:conservation}$, we use the chain rule to mix time and redshift evolution, and assume that the mass-change of binaries is negligible, i.e.~$\diffp{m}{t} = 0$ and $\diffp{q}{t} = 0$, giving:
         \begin{equation}
@@
         \begin{equation}
-            \diffp{f}{z} = - \diffp{t}{z} \Phi(M,z) \diffp{}{a} \lrs{F(M,a,z) \diffp{a}{t}}.
+            \diffp{f}{z} = - \diffp{t}{z} \Phi(M,z) \diffp{}{a} \lrs{F(M,q,a,z) \diffp{a}{t}}.
         \end{equation}
@@
 Because the overall number-density is conserved, we can take a finite step in separation and time/redshift, $a\rightarrow a'$ and $z\rightarrow z'$.
-Here the time it takes for a binary to go from $a \rightarrow a'$ is $\tau(M,a,z|a')$ which leads to a redshift at the later time of $z' = z'(t + \tau)$.
-The standard expression \citep[e.g.~][~Eq.~5]{Chen+2019}, requires making the approximation that,
+Here the time it takes for a binary to go from $a \rightarrow a'$ is $\tau(M,a,z|a')$, which leads to a redshift at the later time of $z' = z'(t + \tau)$.
+The standard expression \citep[e.g.][~Eq.~5]{Chen+2019}, requires making the approximation that,
         \begin{equation}
-            \diffp{}{a} \lrs{F(M,a,z) \diffp{a}{t}} \approx \frac{F}{\Delta t} = \frac{F}{\tau(M,a,z|a')}.
+            \diffp{}{a} \lrs{F(M,q,a,z) \diffp{a}{t}} \approx \frac{F}{\tau(M,q,a,z|a')}.
         \end{equation}
@@
         \begin{equation}
-            \diffp{f(M,a',z')}{{z'}} = \diffp{n}{{\log_{10}\!M}{a'}{z'}} = - \diffp{t}{z} \frac{\Phi(M,z) \, F(M,a,z)}{\tau(M,a,z|a')}.
+            \label{eq:cont_eq_result}
+            \diffp{f(M,q,a',z')}{{z'}} = \diffp{n}{{M}{q}{a'}{z'}} = - \diffp{t}{z} \frac{\Phi(M,z) \, F(M,a,z)}{\tau(M,a,z|a')}.
         \end{equation}
 
-        Utimately we are interested in a total number of binaries $N$ such that $dN \equiv n \, dV_c$ for a comoving volume $V_c(z)$.
-To find the number of binaries, we use the chain rule to connect the time-evolution of the universe with the evolution of each binary over frequency,
-        \begin{equation}
-            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = \diffp{n}{{M}{q}{z}} \diffp{z}{t} \diffp{t}{ {\ln\!f_r} } \diffp{V_c}{z}.
-        \end{equation}
-        Putting this together, we can finally write,
+        Combining \eqref{eq:cont_eq_result} with \eqref{eq:num_num_dens}, we can finally write,
         \begin{equation}
             \label{eq:sam_final}
-            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = - \Phi(M,z) \, F(M,a,z) \frac{\thardf}{\tau(M,a,z|a')} \diffp{V_c}{z}.
+            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = - \Phi(M,z) \, F(M,q,a,z) \frac{\thardf}{\tau(M,q,a,z|a')} \diffp{V_c}{z}.
         \end{equation}

### 2022-03-29 holodeck/sam.py
maint

@@ -472,7 +472,7 @@ class Semi_Analytic_Model:
         Notes
         -----
 
-        d N / d ln f_r = (dn/dz) * (dz/dt) * (dt/d ln f_r) * (dVc/dz)
+        d N / d ln f_r = (dn/dz) * (dt/d ln f_r) * (dz/dt) * (dVc/dz)
                        = (dn/dz) * (f_r / [df_r/dt]) * 4 pi c D_c^2 (1+z) * dz
 
         d N / d ln a   = (dn/dz) * (dz/dt) * (dt/d ln a) * (dVc/dz)
@@ -696,6 +696,7 @@ def sample_sam_with_hardening(
 
     # edges: Mtot [grams], mrat (q), redz (z), {fobs (f) [1/s] OR sepa (a) [cm]}
     edges, dnum = sam.number_from_hardening(hard, fobs=fobs, sepa=sepa, limit_merger_time=limit_merger_time)
+
     log_edges = [np.log10(edges[0]), edges[1], edges[2], np.log(edges[3])]
 
     if cut_below_mass is not None:
@@ -706,8 +707,17 @@ def sample_sam_with_hardening(
     # integrate each bin to convert from probability- density to mass
     # NOTE: _integrate_differential_number() has log-vs-lin spacings hardcoded! use `edges` as is
     mass = holo.sam._integrate_differential_number(edges, dnum, freq=True)
+
+    # ====================================
+    log.warning("Storing values for now!")
+    sam._edges = edges
+    sam._dnum = dnum
+    sam._mass = mass
+    sam._log_edges = log_edges
+    # mass = None
+    # ====================================
+
     # sample binaries from distribution, using appropriate spacing as needed
-    # BUG: should the density used for proportional sampling `dnum` be log(density) ?!
     if (sample_threshold is None) or (sample_threshold == 0.0):
         log.warning(f"Sampling *all* binaries (~{mass.sum():.2e}).")
         log.warning("Set `sample_threshold` to only sample outliers.")

### 2022-04-01 holodeck/sam.py
BUG: differential-number of binaries must be integrated over frequency before poisson sampling
then it needs to be re-converted back to 1/dlogf for proper characteristic strain calculation.

@@ -35,6 +35,8 @@ _AGE_UNIVERSE_GYR = cosmo.age(0.0).to('Gyr').value  # [Gyr]  ~ 13.78
 
 
 class _Galaxy_Stellar_Mass_Function(abc.ABC):
+    """Galaxy Stellar-Mass Function base-class.  Used to calculate number-density of galaxies.
+    """
 
     @abc.abstractmethod
     def __init__(self, *args, **kwargs):
@@ -123,6 +125,8 @@ class GSMF_Schechter(_Galaxy_Stellar_Mass_Function):
 
 
 class _Galaxy_Pair_Fraction(abc.ABC):
+    """Galaxy Pair Fraction base class, used to describe the fraction of galaxies in mergers/pairs.
+    """
 
     @abc.abstractmethod
     def __init__(self, *args, **kwargs):
@@ -151,6 +155,8 @@ class _Galaxy_Pair_Fraction(abc.ABC):
 
 
 class GPF_Power_Law(_Galaxy_Pair_Fraction):
+    """Galaxy Pair Fraction - Single Power-Law
+    """
 
     def __init__(self, frac_norm_allq=0.025, frac_norm=None, mref=1.0e11*MSOL,
                  malpha=0.0, zbeta=0.8, qgamma=0.0, obs_conv_qlo=0.25):
@@ -203,6 +209,8 @@ class GPF_Power_Law(_Galaxy_Pair_Fraction):
 
 
 class _Galaxy_Merger_Time(abc.ABC):
+    """Galaxy Merger Time base class, used to model merger timescale of galaxy pairs.
+    """
 
     @abc.abstractmethod
     def __init__(self, *args, **kwargs):
@@ -230,6 +238,8 @@ class _Galaxy_Merger_Time(abc.ABC):
         return
 
     def zprime(self, mtot, mrat, redz, **kwargs):
+        """Return the redshift after merger (i.e. input `redz` delayed by merger time).
+        """
         tau0 = self(mtot, mrat, redz)  # sec
         age = cosmo.age(redz).to('s').value
         new_age = age + tau0
@@ -249,7 +259,7 @@ class _Galaxy_Merger_Time(abc.ABC):
 
 
 class GMT_Power_Law(_Galaxy_Merger_Time):
-    """
+    """Galaxy Merger Time - simple power law prescription
     """
 
[…]

### 2022-04-01 docs/holodeck-paper/src/ms.tex
MS: text updates

@@
 
+\begin{itemize}
+    \item Change $f_r \rightarrow f$
+    \item holodeck uses $f_\trt{obs}$ as grid variable, not $f_r$, mesh this with the analytic formalism and description
+\end{itemize}
+
 \section{Introduction}
@@
         	\end{align}
-        In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$.  These can be related as \citep[][Eq.~6]{Sesana+2008},
+        In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$, extrapolated into an observer's entire past light-cone.
+These can be related as \citep[][Eq.~6]{Sesana+2008},
         	\begin{align}
@@
 The faster binaries evolve through a given frequency interval, the fewer binaries will occupy that interval on average.
-This is typically described by the binary `hardening' timescale (with respect to frequency), given by the expression,
+This is typically described by the binary `hardening' timescale, given by the expression,
         \begin{equation}
@@
         \end{equation}
+
+        From the SAM formalism, the number of binaries in each bin is calculated by `integrating' each finite element of \eqref{eq:sam_final}, i.e.,
+        \begin{equation}
+            \diffp{N_{ijk}(f_r)}{{\ln\!f_r}} \equiv \diffp{N(M_i, q_j, z_k, {f_r})}{{M}{q}{z}{\ln\!f_r}} \, \Delta M \,\Delta q \, \Delta z.
+        \end{equation}
+        The characteristic strain of the GWB can then be calculated
 

### 2022-04-01 holodeck/sam.py
BUG: do not allow single frequencies to be requrested from SAM.gwb() --- need a full bin (i.e. two edges)

@@ -611,10 +611,16 @@ class Semi_Analytic_Model:
 
         """
 
-        squeeze = False
-        if np.isscalar(fobs):
-            fobs = np.atleast_1d(fobs)
-            squeeze = True
+        # squeeze = False
+        squeeze = True
+        fobs = np.atleast_1d(fobs)
+        if np.isscalar(fobs) or np.size(fobs) == 1:
+            err = "single values of `fobs` are not allowed, can only calculated GWB within some bin of frequency!"
+            err += "  e.g. ``fobs = 1/YR; fobs = [0.9*fobs, 1.1*fobs]``"
+            utils.error(err)
+
+            # fobs = np.atleast_1d(fobs)
+            # squeeze = True
 
         # ---- Get the differential-number of binaries for each bin
         # this is  ``d^4 N / [dlog10(M) dq dz dln(f_r)]``
@@ -639,6 +645,7 @@ class Semi_Analytic_Model:
 
         # ---- find weighted bin centers
         # get unweighted centers
+        print(f"{dnum.shape=}")
         cent = kale.utils.midpoints(dnum, log=False, axis=(0, 1, 2, 3))
         # get weighted centers for each dimension
         for ii, cc in enumerate(coms):

### 2022-04-01 holodeck/sam.py
BUG: do not allow single frequencies to be requrested from SAM.gwb() --- need a full bin (i.e. two edges)

@@ -613,10 +613,16 @@ class Semi_Analytic_Model:
 
         """
 
-        squeeze = False
-        if np.isscalar(fobs):
-            fobs = np.atleast_1d(fobs)
-            squeeze = True
+        # squeeze = False
+        squeeze = True
+        fobs = np.atleast_1d(fobs)
+        if np.isscalar(fobs) or np.size(fobs) == 1:
+            err = "single values of `fobs` are not allowed, can only calculated GWB within some bin of frequency!"
+            err += "  e.g. ``fobs = 1/YR; fobs = [0.9*fobs, 1.1*fobs]``"
+            utils.error(err)
+
+            # fobs = np.atleast_1d(fobs)
+            # squeeze = True
 
         # ---- Get the differential-number of binaries for each bin
         # this is  ``d^4 N / [dlog10(M) dq dz dln(f_r)]``
@@ -641,6 +647,7 @@ class Semi_Analytic_Model:
 
         # ---- find weighted bin centers
         # get unweighted centers
+        print(f"{dnum.shape=}")
         cent = kale.utils.midpoints(dnum, log=False, axis=(0, 1, 2, 3))
         # get weighted centers for each dimension
         for ii, cc in enumerate(coms):

### 2022-07-18 docs/source/conf.py
DOCS: lots of cleanup, update configurations for building sphinx docs

@@ -1,0 +1,115 @@
+# Configuration file for the Sphinx documentation builder.
+#
+# This file does only contain a selection of the most common options. For a
+# full list see the documentation:
+# http://www.sphinx-doc.org/en/stable/config
+
+import os
+import sys
+sys.path.insert(0, os.path.abspath('../..'))
+
+import holodeck   # noqa
+
+
+# -- Project information -----------------------------------------------------
+
+project = 'holodeck'
+copyright = holodeck.__copyright__
+author = holodeck.__author__
+# The short X.Y version
+version = holodeck.__version__
+# The full version, including alpha/beta/rc tags
+release = holodeck.__version__
+
+
+# -- General configuration ---------------------------------------------------
+
+extensions = [
+    'sphinx.ext.autodoc',
+    'sphinx.ext.autosummary',
+    'sphinx.ext.extlinks',
+    'sphinx.ext.intersphinx',
+    'sphinx.ext.mathjax',
+    # 'sphinx.ext.napoleon',
+    'sphinx.ext.viewcode',
+    'numpydoc',     # allow numpy/google style docstrings
+]
+
+autosummary_generate = True
+
+source_suffix = ['.rst', '.md']
+
+master_doc = 'index'
+
+exclude_patterns = ['_build', 'Thumbs.db', '.DS_Store']
+
+# NOTE: `numpy` is actually needed, otherwise things break
+autodoc_mock_imports = ['pytest', 'kalepy']
+
+html_theme = 'sphinx_rtd_theme'
+
+nitpick_ignore = [
+    ('py:class', 'numpy.typing._array_like._SupportsArray'),
+    ('py:class', 'numpy.typing._nested_sequence._NestedSequence'),
+]
+
+# ---- Extensions ------------------------------------------------------------
+
+intersphinx_mapping = {
+    'numpy': ('https://numpy.org/doc/stable/', None),
+}
+
+numpydoc_show_class_members = False
+
+# Report warnings for all validation checks
+# numpydoc_validation_checks = {"all"}
+
+# Report warnings for all checks *except* those listed
+numpydoc_validation_checks = {
+    "all",
+    "GL08",   # The object does not have a docstring
+    "ES01",   # No extended summary found
+    "PR01",   # Parameters { ... } not documented
+    "PR02",   # Unknown parameters { ... }
+    "PR07",   # Parameter has no description
+    "PR10",   # Parameter "___" requires a space before the colon separating the parameter name and type
+    "RT01",   # No Returns section found
+    "RT03",   # Return value has no description
[…]

### 2022-07-18 holodeck/__init__.py
DOCS: update docstrings for sphinx/readthedocs/numpydoc conventions

@@ -3,6 +3,10 @@ holodeck
 Supermassive binary black hole simulator for pulsar timing array signals and galaxy population statistics.
 """
 
+__author__ = "NANOGrav"
+__copyright__ = "Copyright (c) 2022 NANOGrav"
+__license__ = "MIT"
+
 import os
 import logging
 

### 2022-07-18 holodeck/evolution.py

@@ -3,43 +3,55 @@ Holodeck - evolution submodule.
 
 To-Do
 -----
-*   General
-    -   [ ] evolution modifiers should act at each step, instead of after all steps?  This would be
-        a way to implement a changing accretion rate, for example; or to set a max/min hardening
-        rate.
-    -   [ ] re-implement "magic" hardening models that coalesce in zero change-of-redshift or fixed
-        amounts of time.
-*   Dynamical_Friction_NFW
-    -   [ ] Allow stellar-density profiles to also be specified (instead of using a hard-coded
-        Dehnen profile)
-    -   [ ] Generalize calculation of stellar characteristic radius.  Make self-consistent with
-        stellar-profile, and user-specifiable.
-*   Sesana_Scattering
-    -   [ ] Allow stellar-density profile (or otherwise the binding-radius) to be user-specified
-        and flexible.  Currently hard-coded to Dehnen profile estimate.
-*   _SHM06
-    -   [ ] Interpolants of hardening parameters return 2D arrays which we then take the diagonal
-        of, but there should be a better way of doing this.
+
+* General
+
+  * evolution modifiers should act at each step, instead of after all steps?  This would be
+    a way to implement a changing accretion rate, for example; or to set a max/min hardening rate.
+  * re-implement "magic" hardening models that coalesce in zero change-of-redshift or fixed
+    amounts of time.
+
+* Dynamical_Friction_NFW
+
+  * Allow stellar-density profiles to also be specified (instead of using a hard-coded
+    Dehnen profile)
+  * Generalize calculation of stellar characteristic radius.  Make self-consistent with
+    stellar-profile, and user-specifiable.
+
+* Sesana_Scattering
+
+  * Allow stellar-density profile (or otherwise the binding-radius) to be user-specified
+    and flexible.  Currently hard-coded to Dehnen profile estimate.
+
+* _SHM06
+
+  * Interpolants of hardening parameters return 2D arrays which we then take the diagonal
+    of, but there should be a better way of doing this.
+
+* Fixed_Time
+
+  * Handle `rchar` better with respect to interpolation.  Currently not an interpolation
+    variable, which restricts it's usage.
 
 References
 ----------
-*   [Quinlan96] :: Quinlan 1996
[…]

### 2022-07-18 holodeck/log.py

@@ -11,16 +11,16 @@ def get_logger(name='holodeck', level_stream=logging.WARNING, tostr=sys.stdout, 
 
     If logging to output stream (stdout) is enabled, an `_Indent_Formatter` object is used.
 
-    Arguments
-    ---------
+    Parameters
+    ----------
     name : str,
         Handle for this logger, must be distinct for a distinct logger.
+    level_stream : int,
+        Logging level for stream.
     tostr : bool,
         Log to stdout stream.
     tofile : str or `None`,
         Filename to log to (turned off if `None`).
-    level_stream : int,
-        Logging level for stream.
     level_file : int,
         Logging level for file.
 

### 2022-07-18 holodeck/plot.py

@@ -184,8 +184,8 @@ def smap(args=[0.0, 1.0], cmap=None, log=False, norm=None, midpoint=None,
          under='0.8', over='0.8', left=None, right=None):
     """Create a colormap from a scalar range to a set of colors.
 
-    Arguments
-    ---------
+    Parameters
+    ----------
     args : scalar or array_like of scalar
         Range of valid scalar values to normalize with
     cmap : None, str, or ``matplotlib.colors.Colormap`` object
@@ -464,8 +464,8 @@ def plot_gwb(gwb, color=None, uniform=False, nreals=5):
 
     Plots samples, confidence intervals, power-law, and adds twin-Hz axis (x2).
 
-    Arguments
-    ---------
+    Parameters
+    ----------
     gwb : `gravwaves.Grav_Waves` (subclass) instance
 
     Returns

### 2022-07-18 holodeck/relations.py

@@ -2,35 +2,33 @@
 
 References
 ----------
--   Kormendy+Ho-2013 (1304.7762) = [KH13]
-    Coevolution (Or Not) of Supermassive Black Holes and Host Galaxies
-    https://ui.adsabs.harvard.edu/abs/2013ARA%26A..51..511K/abstract
+* Kormendy+Ho-2013 (1304.7762) = [KH13]
+  Coevolution (Or Not) of Supermassive Black Holes and Host Galaxies
+  https://ui.adsabs.harvard.edu/abs/2013ARA%26A..51..511K/abstract
 
--   McConnell+Ma-2013 (1211.2816) = [MM13]
-    Revisiting the Scaling Relations of Black Hole Masses and Host Galaxy Properties
-    https://ui.adsabs.harvard.edu/abs/2013ApJ...764..184M/abstract
+* McConnell+Ma-2013 (1211.2816) = [MM13]
+  Revisiting the Scaling Relations of Black Hole Masses and Host Galaxy Properties
+  https://ui.adsabs.harvard.edu/abs/2013ApJ...764..184M/abstract
 
 * [NFW-97] : Navarro, Frenk & White 1997
-    A Universal Density Profile from Hierarchical Clustering
-    https://ui.adsabs.harvard.edu/abs/1997ApJ...490..493N/abstract
+  A Universal Density Profile from Hierarchical Clustering
+  https://ui.adsabs.harvard.edu/abs/1997ApJ...490..493N/abstract
 
 * [Guo-2010] : Guo, White, Li & Boylan-Kolchin 2010
-    How do galaxies populate dark matter haloes?
-    https://ui.adsabs.harvard.edu/abs/2010MNRAS.404.1111G/abstract
+  How do galaxies populate dark matter haloes?
+  https://ui.adsabs.harvard.edu/abs/2010MNRAS.404.1111G/abstract
 
 * [Behroozi+2013] : Behroozi, Wechsler & Conroy 2013
-    The Average Star Formation Histories of Galaxies in Dark Matter Halos from z = 0-8
-    https://ui.adsabs.harvard.edu/abs/2013ApJ...770...57B/abstract
+  The Average Star Formation Histories of Galaxies in Dark Matter Halos from z = 0-8
+  https://ui.adsabs.harvard.edu/abs/2013ApJ...770...57B/abstract
 
 * [Klypin+2016] : Klypin et al. 2016
-    MultiDark simulations: the story of dark matter halo concentrations and density profiles
[…]

### 2022-07-18 holodeck/utils.py

@@ -107,8 +107,8 @@ def tqdm(*args, **kwargs):
 def get_file_size(fnames, precision=1):
     """Return a human-readable size of a file or set of files.
 
-    Arguments
-    ---------
+    Parameters
+    ----------
     fnames : str or list
         Paths to target file(s)
     precisions : int,
@@ -197,8 +197,8 @@ def frac_str(vals: npt.ArrayLike, prec: int = 2) -> str:
 
     e.g. [0, 1, 2, 0, 0] ==> "2/5 = 4.0e-1"
 
-    Arguments
-    ---------
+    Parameters
+    ----------
     vals : (N,) array_like,
         Input array to find non-zero elements of.
     prec : int
@@ -370,8 +370,8 @@ def nyquist_freqs(
 ) -> np.ndarray:
     """Calculate Nyquist frequencies for the given timing parameters.
 
-    Arguments
-    ---------
+    Parameters
+    ----------
     dur : float,
         Duration of observations
     cad : float,
@@ -415,12 +415,12 @@ def quantiles(
 
     NOTE: if `values` is a masked array, then only unmasked values are used!
 
-    Arguments
-    ---------
+    Parameters
+    ----------
     values: (N,)
         input data
-    percs: (M,) scalar [0.0, 1.0]
-        Desired quantiles of the data.
+    percs: (M,) scalar
+        Desired quantiles of the data.  Within range of [0.0, 1.0].
     weights: (N,) or `None`
         Weights for each input data point in `values`.
     axis: int or `None`,
@@ -544,8 +544,8 @@ def trapz_loglog(
     For each interval (x[i+1], x[i]), calculate the integral assuming that y is of the form,
         `y = a * x^gamma`
 
-    Arguments
-    ---------
+    Parameters
+    ----------
     yy : ndarray
     xx : (X,) array_like of scalar,
     bounds : (2,) array_like of scalar,
@@ -648,8 +648,8 @@ def trapz_loglog(
 def trapz(yy: npt.ArrayLike, xx: npt.ArrayLike, axis: int = -1, cumsum: bool = True):
     """Perform a cumulative integration along the given axis.
 
-    Arguments
-    ---------
+    Parameters
+    ----------
     yy : ArrayLike of scalar,
         Input to be integrated.
     xx : ArrayLike of scalar,
[…]

### 2022-07-18 holodeck/cosmology.py
DOCS: update docstrings for sphinx/readthedocs/numpydoc conventions

@@ -18,8 +18,8 @@ class Cosmology(ap.cosmology.FlatLambdaCDM):
     given a derived quantitiy (i.e. to come from d_L to redshift), the numerical grid can then be
     inverted and interpolated to find the corresponding redshift.
 
-    API Methods
-    -----------
+    Methods
+    -------
     -   tage_to_z                      -   Convert from universe-age to redshift.
     -   dcom_to_z                      -   Convert from comoving-dist to redshift.
     -   dlum_to_z                      -   Convert from luminosity-dist to redshift.
@@ -73,8 +73,8 @@ class Cosmology(ap.cosmology.FlatLambdaCDM):
     def _init_interp_grid(z_pnts, num_pnts):
         """Create a grid in redshift for interpolation.
 
-        Arguments
-        ---------
+        Parameters
+        ----------
         z_pnts : (N,) array_like of scalar
             Monotonically decreasing redshift values greater than zero.  Inbetween each pair of
             values, `num_pnts` points are evenly added in log-space.  Also, `num_pnts` are added

### 2022-07-18 holodeck/sam.py

@@ -2,16 +2,16 @@
 
 References
 ----------
--   Chen, Sesana, Conselice 2019 = [Chen19]
-    https://ui.adsabs.harvard.edu/abs/2019MNRAS.488..401C/abstract
-    Constraining astrophysical observables of galaxy and supermassive black hole binary mergers
-        using pulsar timing arrays
+* Chen, Sesana, Conselice 2019 = [Chen19]
+  Constraining astrophysical observables of galaxy and supermassive black hole binary mergers
+  using pulsar timing arrays
+  https://ui.adsabs.harvard.edu/abs/2019MNRAS.488..401C/abstract
 
 To-Do
 -----
-*[ ]Check that _GW_ frequencies and _orbital_ frequencies are being used in the correct places.
-    Check `number_at_gw_fobs` and related methods.
-*[ ]Incorporate arbitrary hardening mechanisms into SAM construction, sample self-consistently.
+* Check that *GW* frequencies and *orbital* frequencies are being used in the correct places.
+  Check `number_at_gw_fobs` and related methods.
+* Incorporate arbitrary hardening mechanisms into SAM construction, sample self-consistently.
 
 """
 
@@ -48,8 +48,8 @@ class _Galaxy_Stellar_Mass_Function(abc.ABC):
 
         i.e. Phi = dn / dlog10(M)
 
-        Arguments
-        ---------
+        Parameters
+        ----------
         mstar : scalar or ndarray
             Galaxy stellar-mass in units of [grams]
         redz : scalar or ndarray
@@ -87,8 +87,8 @@ class GSMF_Schechter(_Galaxy_Stellar_Mass_Function):
 
         See: [Chen19] Eq.8
 
-        Arguments
-        ---------
+        Parameters
+        ----------
         mstar : scalar or ndarray
             Galaxy stellar-mass in units of [grams]
         redz : scalar or ndarray
@@ -136,8 +136,8 @@ class _Galaxy_Pair_Fraction(abc.ABC):
     def __call__(self, mtot, mrat, redz):
         """Return the fraction of galaxies in pairs of the given parameters.
 
-        Arguments
-        ---------
+        Parameters
+        ----------
         mtot : scalar or ndarray,
             Total-mass of combined system, units of [grams].
         mrat : scalar or ndarray,
@@ -182,8 +182,8 @@ class GPF_Power_Law(_Galaxy_Pair_Fraction):
     def __call__(self, mtot, mrat, redz):
[…]

### 2022-07-18 docs/source/conf.py
BUG: issue in readthedocs.yml configuration

@@ -44,7 +44,9 @@ master_doc = 'index'
 exclude_patterns = ['_build', 'Thumbs.db', '.DS_Store']
 
 # NOTE: `numpy` is actually needed, otherwise things break
-autodoc_mock_imports = ['pytest', 'kalepy']
+autodoc_mock_imports = [
+    # 'pytest', 'kalepy', 'astropy', 'h5py', 'kalepy', 'matplotlib', 'scipy', 'tqdm',
+]
 
 html_theme = 'sphinx_rtd_theme'
 

### 2022-07-18 docs/holodeck-paper/src/sam_text.tex

--- 2022-03-29 docs/holodeck-paper/src/ms.tex
@@
-% =================================================================================================
-% Holodeck Paper Manuscript
-% -------------------------
-%
-%
-%
-% =================================================================================================
-
-% general package imports and setup
-\input{packages}
-% definitions of custom commands
-\input{commands}
-% journal abbreviations for bibliography
-\input{abbrevs}
-
-% =================================================================================================
-% ====    Front Matter
-% =================================================================================================
-
-\newtotcounter{citnum} % From the package documentation
-\def\oldbibitem{} \let\oldbibitem=\bibitem
-\def\bibitem{\stepcounter{citnum}\oldbibitem}
-
-\newcommand{\orcidauthorA}{0000-0002-6625-6450} % For author A
-
-\title[Short Title]{Very long and verbose title}
-\author[NANOGrav]{the NANOGrav collaboration}
-% \author[L.Z.~Kelley et al.]{
-%     Luke Zoltan Kelley$^{1}$\thanks{E-mail:lzkelley@northwestern.edu}\orcidA{} \\
-%     $^{1}$ \begin{minipage}[t]{\linewidth} Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA), and\\Department of Physics \& Astronomy, Northwestern University, Evanston, IL 60208
-%     \end{minipage}
-% }
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-    abstract
-\end{abstract}
-
-\begin{keywords}
-    quasars: supermassive black holes
-\end{keywords}
-
-% =================================================================================================
-% ====    Section 1 - Introduction
-% =================================================================================================
-
-\section{Introduction}
-    \label{sec:intro}
-
-% =================================================================================================
-% ====    Section 2 - Methods
-% =================================================================================================
-
-\section{Methods}
-    \label{sec:meth}
-
 \subsection{Source Distributions and Gravitational Wave Calculations}
@@
         \end{align}
-        In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$.  These can be related as \citep[][Eq.~6]{Sesana+2008},
+    In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$, extrapolated into an observer's entire past light-cone.
+These can be related as \citep[][Eq.~6]{Sesana+2008},
         \begin{align}
@@
 The faster binaries evolve through a given frequency interval, the fewer binaries will occupy that interval on average.
-This is typically described by the binary `hardening' timescale (with respect to frequency), given by the expression,
+This is typically described by the binary `hardening' timescale (or `residence time'), given by the expression,
     \begin{equation}
-            \thardf \equiv \diffp{t}{{\ln\!f_r}} = \frac{f_r}{\partial f_r / \partial t}.
+        \thardf \equiv - \diffp{t}{{\ln\!f_r}} = - \frac{f_r}{\partial f_r / \partial t}.
     \end{equation}
@@
 
-        Define the distribution function of sources as $f(M,q,a,z) = d^3 n(M,q,a,z) / dM dq da$.
-Here $M$ is the total mass of each systems, $a$ is the binary separation, and $z$ is the redshift.
-Note that binary frequency $f_r$ could be used almost interchangeably with $a$, and that both evolve as a function of time or equivalently redshift.
-This formalism can easily be extended to include other parameters, such as mass-ratio and eccentricity ($q$ and $e$).
+    Define the distribution function of sources as $F(M,q,a,z) = d^2 n(M,q,a,z) / dM dq$.
+Here $M$ is the total mass of each systems, the mass-ratio is $q\equiv m_2/m_1 \leq 1$, $a$ is the binary separation, and $z$ is the redshift.
+    % Define the distribution function of sources as $F(M,q,f_r,z) = d^2 n(M,q,f_r,z) / dM dq$.
+Here $M$ is the total mass of each systems, the mass-ratio is $q\equiv m_2/m_1 \leq 1$, $f_r$ is the binary orbital frequency in the rest frame, and $z$ is the redshift.
     We can write the conservation equation for binaries as of function of redshift as,
@@
         \label{eq:conservation}
-            \diffp{f}{z} +
-                \diffp{}{M} \lrs{f \diffp{M}{z}} +
-                \diffp{}{q} \lrs{f \diffp{q}{z}} +
-                \diffp{}{a} \lrs{f \diffp{a}{z}} = S_f(M, q, a, z).
+        \diffp{F}{z} +
+            \diffp{}{M} \lrs{F \diffp{M}{z}} +
+            \diffp{}{q} \lrs{F \diffp{q}{z}} +
+            \diffp{}{a} \lrs{F \diffp{a}{z}} = S_{\!F}(M, q, a, z).
     \end{equation}
-        Here $S_f$ is a source/sink function that can account for the creation or destruction of binaries.
+    Here $S_{\!F}$ is a source/sink function that can account for the creation or destruction of binaries.
 
-        We consider the standard \citep[e.g.~Sesana style; see][]{Chen+2019} semi-analytic model (SAM) formalism of MBH binary populations.
-In this style of calculation, $f$ is determined in a region of parameter space that can be observed/estimated, and this is \textit{evolved} to find the distribution in a different region of parameter space that is of interest.
+    We consider the standard \citep[e.g.~][]{Sesana+2008, Chen+2019} semi-analytic model (SAM) formalism of MBH binary populations.
+In this style of calculation, $F$ is determined in a region of parameter space that can be observed/estimated, and this is evolved to find the distribution in a different region of parameter space that is of interest.
 In practice, the observed parameter space is galaxies and galaxy mergers, and the parameter space of interest is close separated MBH binaries that could be GW detectable.
-\note{Thus we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts), and we set $S_f = 0$.}
+Thus we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts), and we set $S_{\!F}$.
 
@@
         \label{eq:dist_func}
-            f(M,q,a,z) = \frac{\Phi(M, z)}{M \ln\!10} \cdot F_a(M,q,a,z),
+        F(M,q,z) = \frac{\Phi(M, z)}{M \ln\!10} \cdot P(M,q,z),
     \end{equation}
@@
 Typically the MBH--galaxy relation is given in terms of an `$M_\trt{BH}$--$M_\trt{bulge}$' relation \citep[e.g.][]{McConnell+Ma-2013}, which is a observationally derived relation between the mass of MBHs and the mass of the stellar bulge component of their host galaxy.
-        In \citet{Chen+2019}, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$f(M,q,z) = \Phi(M, z) \cdot F(M, q, z)$ s.t.~$F = \int_{a_0}^{a_1} F_a \, da$.
+    In \citet{Chen+2019}, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$P = \int_{a_0}^{a_1} P_a \, da$.
+
     From $\eqref{eq:conservation}$, we use the chain rule to mix time and redshift evolution, and assume that the mass-change of binaries is negligible, i.e.~$\diffp{m}{t} = 0$ and $\diffp{q}{t} = 0$, giving:
     \begin{equation}
-            \diffp{f}{z} = - \diffp{t}{z} \diffp{}{a} \lrs{f \diffp{a}{t}}.
-        \end{equation}
-        Plugging-in \eqref{eq:dist_func}:
-        \begin{equation}
-            \diffp{f}{z} = - \diffp{t}{z} \Phi(M,z) \diffp{}{a} \lrs{F(M,q,a,z) \diffp{a}{t}}.
+        \diffp{F}{z} = - \diffp{t}{z} \diffp{}{a} \lrs{F \diffp{a}{t}}.
     \end{equation}
@@
 Because the overall number-density is conserved, we can take a finite step in separation and time/redshift, $a\rightarrow a'$ and $z\rightarrow z'$.
-Here the time it takes for a binary to go from $a \rightarrow a'$ is $\tau(M,a,z|a')$, which leads to a redshift at the later time of $z' = z'(t + \tau)$.
-The standard expression \citep[e.g.][~Eq.~5]{Chen+2019}, requires making the approximation that,
+Here the time it takes for a binary to go from $a \rightarrow a'$ is $T(M,q,a,z|a')$, which leads to a redshift at the later time of $z' = z'(t + T)$.
+So far, we have left the binary separation $a$ as implicit in the expression for F.
+To obtain the standard expression \citep[e.g.][~Eq.~5]{Chen+2019}, we make the approximation that,
     \begin{equation}
-            \diffp{}{a} \lrs{F(M,q,a,z) \diffp{a}{t}} \approx \frac{F}{\tau(M,q,a,z|a')}.
+        \diffp{}{a} \lrs{F(M,q,z) \diffp{a}{t}} \approx \frac{F}{T(M,q,a,z|a')}.
     \end{equation}
@@
         \label{eq:cont_eq_result}
-            \diffp{f(M,q,a',z')}{{z'}} = \diffp{n}{{M}{q}{a'}{z'}} = - \diffp{t}{z} \frac{\Phi(M,z) \, F(M,a,z)}{\tau(M,a,z|a')}.
+        \diffp{F(M,q,a',z')}{{z'}} = \diffp{n}{{M}{q}{z'}} = - \diffp{t}{z} \frac{\Phi(M,z) \, P(M,q,z)}{T(M,q,a,z|a')}.
     \end{equation}
@@
         \label{eq:sam_final}
-            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = - \Phi(M,z) \, F(M,q,a,z) \frac{\thardf}{\tau(M,q,a,z|a')} \diffp{V_c}{z}.
+        \diffp{N}{{M}{q}{z}{\ln\!f_r}} = \frac{\Phi(M,z) \, P(M,q,z)}{T(M,q,a,z|a')} \, \thardf \, \diffp{V_c}{z}.
     \end{equation}
 
-    % % Fig -
-    % \begin{figure}
-    %     \centering
-    %     \includegraphics[width=1\columnwidth]{{{example-image-c}}}
-    %     \caption{\textbf{header.} body.}
-    %     \label{fig:fig1}
-    % \end{figure}
-    %
-    %
-    % \begin{equation}\begin{split}
-    %     \label{eq:eq1}
-    %     2+2 & = 4 \\
-    %         & = 10 - 6
-    % \end{split}\end{equation}
-    %
-    % % Fig -
-    % \begin{figure*}
-	% 	\centering
-    %     \subfloat[first]{{
-    %         \includegraphics[width=1.75\columnwidth]{{{example-image-a}}}
-    %     }}
-	% 	\\
-    %     % \qquad
-    %     \subfloat[second]{{
-    %     	\includegraphics[width=1.75\columnwidth]{{{example-image-b}}}
-    %     }}
-    %     \caption{\textbf{Head.}  Body.}
-    %     \label{fig:fig2}
-    % \end{figure*}
-
-\section{Implementation}
-    \label{sec:imp}
-
-    The \holodeck{} framework is written primarily in \python{} using a class-based structure for modularity and extensibility.
-
-    \subsection{Semi-Analytic Models}
-        \label{sec:imp_sam}
-
-        The core of the semi-analytic modeling framework in \holodeck{} is implementing \eqref{eq:sam_final}.
-There are classes for each of the key components: The galaxy stellar-mass function (GSMF, $\Phi$), the galaxy pair fraction \(F\), the merger timescale ($\tau$), and additional a scaling relation for mapping host galaxies via to MBH properties --- typically via an \mmbulge{} relation.
-Each component class is implemented generically via basic API methods (e.g.~a GSMF that returns the number density of galaxies at a given stellar-mass and redshift), and can be subclassed to provide a particular implementation (e.g.~an analytic Schechter-function GSMF).
-The number (or number density) of binaries is specified over a grid of parameters, typically in a four dimensional space of total mass ($M$), mass ratio ($q$), redshift ($z$), and binary rest-frame orbital frequency ($f_r$).
-
-        The semi-analytic model formalism defines a continuous distribution of MBH binaries.
-For many calculations, the discrete nature of MBHB sources is important, not only for examining individual continuous wave sources but also because of the important effects of discrete sources on the shape of the GWB due to the non-linear scaling of GW strain \citep{Sesana+2008}.
-Additional, a component of the uncertainty in expected GW signals is due to `cosmic variance', which can be treated as Poisson noise when creating multiple samples (or `realizations') of our binary populations.
-To handle discretization and resampling we use the \texttt{kalepy} package \citep{kalepy2021}, in particular, a method of outlier sampling to capture regions of parameter space where finite number effects are important.
-The \holodeck{} semi-analytic models specify the number (or number-density) of binaries over a grid, i.e.~$N_{ijkl}$.
-By definition, grid cells which have an expected total number of binaries $\gg 1$ will not produce appreciable finite-number effects, so instead they are given a weighting based on a Poisson distribution centered around their expectation value, i.e.~$\mathcal{P}(N_{ijkl})$.
-Grid cells with expected numbers of order unity or less do produce important finite-number effects, and thus are directly resampled
-
-% =================================================================================================
-% ====    Section 3 - Results
-% =================================================================================================
-\section{Results}
-    \label{sec:res}
-
-    % Fig -
-    % \begin{figure*}
-    %     \centering
-    %     \includegraphics[width=1.5\columnwidth]{{{figs/det-nums_fedd-1.00_voff+3.00_dvel+2.00_tobs+0.70}}}
-    %     \caption{\textbf{Distributions of parameters for all binaries (black) and those with a detectable kinematic signature in the secondary AGN (colors).}  Less than $10^{-7}$ of primaries are observable, which are not shown.
-Secondary AGN with observable velocity offsets ($\voff$) are shown in blue assuming a sensitivity of \mbox{$\voffsens = 10^3 \kmps$}.
-Secondaries with observable changing velocities ($\dvel$) are shown in orange (solid) for a sensitivity $\dvelsens = 10^2 \kmps$, and observing baseline of \mbox{$\tobs = 5\, \yr$}.
-Systems where both offsets and changes are detectable (orange, dotted) are mostly a uniform subset of the $\dvel$ sample, and are almost indistinguishable in the 2D contour plots.
-An optical flux cutoff of \mbox{$6\E{-14} \mathrm{erg/s/cm}^2$} (AB Mag $\approx 21$) is also imposed, where all AGN are assumed to be accreting at an Eddington fraction of $0.1$.}
-    %     \label{fig:det_bins_num}
-    % \end{figure*}
-
-% =================================================================================================
-% ====    Section 4 - Conclusions
-% =================================================================================================
-\section{Discussion \& Conclusions}
-    \label{sec:disc}
-
-% =================================================================================================
-% ====    End Matter
-% =================================================================================================
-
-% ====    Acknowledgments
-
-\section*{Acknowledgments}
-	I am very thankful to.
-
-    This research made use of \texttt{astropy}, a community-developed core Python package for Astronomy \citep{astropy2013}, in addition to \texttt{Scipy}~\citep{scipy}, \texttt{ipython}~\citep{ipython}, \texttt{jupyter}~notebook~\citep{jupyter}, \texttt{Numpy}~\citep{numpy2011} \& \texttt{SymPy}~\citep{sympy2017}.
-All figures were generated using \texttt{matplotlib}~\citep{matplotlib2007}.
-Kernel density estimation was performed using the \texttt{kalepy}{} package (\href{https://github.com/lzkelley/kalepy}{github.com/lzkelley/kalepy}) \citep{kalepy2020}.
-
-    The Illustris data is available online at \href{https://www.illustris-project.org/}{www.illustris-project.org} \citep{Nelson+2015}, and Illustris-TNG data at \href{https://www.tng-project.org/}{www.tng-project.org} \citep{Nelson+2019}.
-
-% ====    Bibliography
-
-\let\oldUrl\url
-\renewcommand{\url}[1]{\href{#1}{Link}}
-
-\quad{}
-\bibliographystyle{mnras}
-\bibliography{refs}
-
-\onecolumn
-\clearpage
-
-% ====    Appendices
-
-\appendix
-
-    \section{Additional Equations}
-        \label{sec:app_eqs}
-        A binary is described by its total mass $M$, mass ratio $q \equiv m_2 / m_1$ (s.t.~$q\leq1$), (proper) semi-major axis $a$, and rest-frame orbital frequency $f_r$.
-The characteristic gravitational wave mass is the chirp mass, $\mchirp \equiv M \, q^{3/5} \, \lr[-6/5]{1+q}$.
-
-        The GW strain from a circular binary is \citep[][Eq.~7; sky and polarization averaged]{Sesana+2008},
-            \begin{equation}
-            \label{eq:source_strain}
-            h_s = \frac{8}{10^{1/2}} \frac{\left(G\mchirp\right)^{5/3}}{c^4 \, d_L}
-                \left(2 \pi f_r \right)^{2/3},
-            \end{equation}
-        for a luminosity distance $d_L$.
-Recall that the luminosity distance and comoving distance are related as $d_L = d_c \, (1+z)$, that the comoving volume of the Universe is given by \citep{Hogg-1999},
-        \begin{equation}
-            \frac{d V_c}{dz} = 4\pi \frac{c}{H_0} \frac{d_c^2}{E(z)},
-        \end{equation}
-        and that redshift can be related to time (or age of the universe) as,
-        \begin{align}
-            \frac{dz}{dt} = H_0 \lr{1+z} E(z).
-        \end{align}
-        The evolution of the Hubble parameter over redshift can be calculated from the matter-energy constituents of the universe as,
-        \begin{align}
-            H(z) & = H_0(z) E(z), \\
-            E(z) & \equiv \lrs[1/2]{\Omega_R \lr[4]{1+z} + \Omega_m \lr[3]{1+z} + \Omega_k \lr[2]{1+z} + \Omega_\Lambda}.
-        \end{align}
-        The $\Omega$ terms are the usual redshift-zero fractions of the critical density for radiation, non-relativistic matter, curvature, and dark energy.
-We adopt the results of WMAP9 \citep{Hinshaw+2013} for our fiducial cosmology, assuming zero curvature: $H_0 = 69.33 \textrm{ km/s/Mpc}$, $\Omega_m = 0.2880$, $\Omega_\Lambda = 0.7120$.
-
-    \section{Additional Material}
-        \label{sec:app}
-
-        \noindent\begin{minipage}{\linewidth}
-
-            % Fig -
-            % \centering
-            % \includegraphics[width=0.55\columnwidth]{{{figs/doan+2019_jitter}}}
-            % \captionof{figure}{Distribution of velocity-offset variations (`jitters') from \citep{Doan+2019}, for the red and blue components of the observed BLs.
-Measurements are shown as ticks at $y=0$, and KDE distributions are shown using a Scott's factor bandwidth, and Gaussian kernel.
-The purple curve is the KDE distribution taking both red and blue data as independent.
-The KDE $50\%$ (median) and $99\%$ (Gaussian standard deviation $\sigma \approx 2.3$) jitter for each component are shown with dashed vertical lines and corresponding labels.}
-            % \label{fig:jitter_doan2019}
-			%
-            % \setlength{\tabcolsep}{8pt}
-            % \begin{tabular}{p{1.5cm} c c c c c c c}
-            %      & & & $5\%$ & $25\%$ & $50\%$ & $75\%$ & $95\%$ \\ [0.5ex]
-            %     \hline
-            %     \multirow{4}{*}{All}  & \multirow{4}{*}{-}
-            %         & $M \; [M_\odot]$ & $3.4 \times10^{ 6 }$ & $7.2 \times10^{ 6 }$ & $1.8 \times10^{ 7 }$ & $6.0 \times10^{ 7 }$ & $4.1 \times10^{ 8 }$ \\
-            %         & & q & $2.9 \times10^{ -2 }$ & $1.7 \times10^{ -1 }$ & $4.0 \times10^{ -1 }$ & $6.8 \times10^{ -1 }$ & $9.3 \times10^{ -1 }$ \\
-            %         & & z & 0.24 & 0.41 & 0.53 & 0.62 & 0.69 \\
-            %         & & $p \; [\mathrm{yr}]$ & $1.4 \times10^{ 2 }$ & $1.5 \times10^{ 3 }$ & $9.6 \times10^{ 3 }$ & $4.0 \times10^{ 4 }$ & $2.0 \times10^{ 5 }$ \\
-            %         & & $a \; [\mathrm{pc}]$ & $2.1 \times10^{ -2 }$ & $1.3 \times10^{ -1 }$ & $4.9 \times10^{ -1 }$ & $1.3 \times10^{ 0 }$ & $4.1 \times10^{ 0 }$ \\
-            %         & & $a \; [r_g]$ & $1.2 \times10^{ 4 }$ & $1.4 \times10^{ 5 }$ & $4.5 \times10^{ 5 }$ & $1.2 \times10^{ 6 }$ & $4.1 \times10^{ 6 }$ \\
-            %     \hline
-            %     \multirow{4}{1.5cm}{Secondary\newline Offset\newline($\voff$)} & \multirow{4}{*}{$0.49\%$}
-            %         & $M \; [M_\odot]$ & $1.4 \times10^{ 8 }$ & $4.0 \times10^{ 8 }$ & $7.6 \times10^{ 8 }$ & $1.4 \times10^{ 9 }$ & $3.7 \times10^{ 9 }$ \\
-            %         & & q & $9.6 \times10^{ -4 }$ & $4.2 \times10^{ -3 }$ & $1.2 \times10^{ -2 }$ & $3.6 \times10^{ -2 }$ & $1.5 \times10^{ -1 }$ \\
-            %         & & z & 0.25 & 0.42 & 0.54 & 0.63 & 0.69 \\
-            %         & & $p \; [\mathrm{yr}]$ & $8.3 \times10^{ 2 }$ & $1.5 \times10^{ 3 }$ & $2.3 \times10^{ 3 }$ & $3.7 \times10^{ 3 }$ & $9.0 \times10^{ 3 }$ \\
-            %         & & $a \; [\mathrm{pc}]$ & 0.22 & 0.41 & 0.58 & 0.87 & 1.82 \\
-            %         & & $a \; [r_g]$ & $5.0 \times10^{ 3 }$ & $1.1 \times10^{ 4 }$ & $1.8 \times10^{ 4 }$ & $2.9 \times10^{ 4 }$ & $4.9 \times10^{ 4 }$ \\
-            % \end{tabular}
-            % \captionof{table}{\textbf{Parameters of detectable binary systems with redshift $z < 0.7$.}  The indicated quantiles are given for total mass ($M$), mass ratio ($q$), separation ($a$), and orbital period ($p$).
-In only four binaries ($2\E{-8}$ of systems) is the primary detectable, and those have parameters: \mbox{$M\approx 2$--$5\E{9} \, \msol$},
-            % \hspace{0.5ex} \mbox{$q \approx 0.7$--$0.9$}, \hspace{0.5ex} \mbox{$z \approx 0.6$--$0.8$}, \hspace{0.5ex} \mbox{$a \approx 2.2$--$2.8 \, \pc$}.}
-            % \label{tab:obs_pars}
-
-        \end{minipage}
-
-    \twocolumn
-
-\end{document}
+    % From the SAM formalism, the number of binaries in each bin is calculated by `integrating' each finite element of \eqref{eq:sam_final}, i.e.,
+    % \begin{equation}
+    %     \diffp{N_{ijk}(f_r)}{{\ln\!f_r}} \equiv \diffp{N(M_i, q_j, z_k, {f_r})}{{M}{q}{z}{\ln\!f_r}} \, \Delta M \,\Delta q \, \Delta z.
+    % \end{equation}
+    % The characteristic strain of the GWB can then be calculated

### 2022-07-19 docs/holodeck-paper/src/ms.tex

--- 2022-03-28 docs/holodeck-paper/src/ms.tex
@@
 
-    \subsection{Source Distributions and Gravitational Wave Calculations}
-
-        Let $N(z)$ denote the number of binaries out to a redshift $z$, corresponding to a comoving volume $V_c$.
-        We can calculate the characteristic strain from an ensemble of GW sources as \citep[][Eqs.~5/8]{Phinney-2001},
-        	\begin{align}
-        	h_c^2(f) = \int_0^\infty \!\! dz \; \frac{d^2 N}{dz \, d\ln f_r} \; h_s^2\lr{f_r}.
-        	\end{align}
-        In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$.  These can be related as \citep[][Eq.~6]{Sesana+2008},
-        	\begin{align}
-        	\frac{d^2 N}{dz \, d\ln f_r} = \frac{d n_c}{dz} \frac{dz}{dt} \frac{dt}{d\ln f_r} \frac{d V_c}{dz}.
-        	\end{align}
-        Note that the total number of binaries in a given frequency interval depends not only on the number density, but also their rate of frequency-evolution.
-The faster binaries evolve through a given frequency interval, the fewer binaries will occupy that interval on average.
-This is typically described by the binary `hardening' timescale (with respect to frequency), given by the expression,
-        \begin{equation}
-            \thardf \equiv \diffp{t}{{\ln\!f_r}} = \frac{f_r}{\partial f_r / \partial t}.
-        \end{equation}
-        Plugging in the relevant cosmographic relationships (see \secref{sec:app_eqs}) we have,
-        	\begin{align}
-        	h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
-        	\end{align}
-        The challenge in modeling astrophysical sources of GWs is in calculating their number density.  Numerous approaches are possible, described in the following sections.
-
-    \subsection{Semi-Analytic Modeling}
-
-        Define the distribution function of sources as $f(M,a,z) = d^2 n(M,a,z) / dM da$.
-Here $M$ is the total mass of each systems, $a$ is the binary separation, and $z$ is the redshift.
-Note that binary frequency $f_r$ could be used almost interchangeably with $a$, and that both evolve as a function of time or equivalently redshift.
-This formalism can easily be extended to include other parameters, such as mass-ratio and eccentricity ($q$ and $e$).
-We can write the conservation equation for binaries as of function of redshift as,
-        \begin{equation}
-            \label{eq:conservation}
-            \diffp{f}{z} + \diffp{}{m} \lrs{f \diffp{m}{z}} + \diffp{}{a} \lrs{f \diffp{a}{z}} = S_f(m, a, z).
-        \end{equation}
-        Here $S_f$ is a source/sink function that can account for the creation or destruction of binaries.
-Here we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts).
-Thus we set $S_f = 0$.
-
-        We consider the standard \citep[e.g.~Sesana style; see][]{Chen+2019} semi-analytic model (SAM) formalism of MBH binary populations.
-In this style of calculation, $f$ is determined in a region of parameter space that can be observed/estimated, and this is \textit{evolved} to find the distribution in a different region of parameter space that is of interest.
-In practice, the observed parameter space is galaxies and galaxy mergers, and the parameter space of interest is close separated MBH binaries that could be GW detectable.
-We will express the distribution function as a product of a mass-function, and a pair fraction:
-        \begin{equation}
-            \label{eq:dist_func}
-            f(M,a,z) = \frac{\Phi(M, z)}{M \ln\!10} \cdot F_a(M, a, z),
-        \end{equation}
-        where the mass-function is, $\Phi(M, z) \equiv \diffp{n_g}{{\logten{M}}}$.
-        In \citet{Chen+2019}, for example, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$f(M, z) = \Phi(M, z) \cdot F(M, z)$ s.t.~$F = \int_{a_0}^{a_1} F_a \, da$.
-From $\eqref{eq:conservation}$, we use the chain rule to mix time and redshift evolution, and assume that the mass-change of binaries is negligible, i.e.~$\diffp{m}{t} = 0$, giving:
-        \begin{equation}
-            \diffp{f}{z} = - \diffp{t}{z} \diffp{}{a} \lrs{f \diffp{a}{t}}.
-        \end{equation}
-        Plugging-in \eqref{eq:dist_func}:
-        \begin{equation}
-            \diffp{f}{z} = - \diffp{t}{z} \Phi(M,z) \diffp{}{a} \lrs{F(M,a,z) \diffp{a}{t}}.
-        \end{equation}
-
-        The binary population is assumed to be changing only in separation and redshift, which are related by $\partial a / \partial z = (\partial a / \partial t) (\partial t / \partial z)$.
-Because the overall number-density is conserved, we can take a finite step in separation and time/redshift, $a\rightarrow a'$ and $z\rightarrow z'$.
-Here the time it takes for a binary to go from $a \rightarrow a'$ is $\tau(M,a,z|a')$ which leads to a redshift at the later time of $z' = z'(t + \tau)$.
-The standard expression \citep[e.g.~][~Eq.~5]{Chen+2019}, requires making the approximation that,
-        \begin{equation}
-            \diffp{}{a} \lrs{F(M,a,z) \diffp{a}{t}} \approx \frac{F}{\Delta t} = \frac{F}{\tau(M,a,z|a')}.
-        \end{equation}
-        Thus giving,
-        \begin{equation}
-            \diffp{f(M,a',z')}{{z'}} = \diffp{n}{{\log_{10}\!M}{a'}{z'}} = - \diffp{t}{z} \frac{\Phi(M,z) \, F(M,a,z)}{\tau(M,a,z|a')}.
-        \end{equation}
-
-        Utimately we are interested in a total number of binaries $N$ such that $dN \equiv n \, dV_c$ for a comoving volume $V_c(z)$.
-To find the number of binaries, we use the chain rule to connect the time-evolution of the universe with the evolution of each binary over frequency,
-        \begin{equation}
-            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = \diffp{n}{{M}{q}{z}} \diffp{z}{t} \diffp{t}{ {\ln\!f_r} } \diffp{V_c}{z}.
-        \end{equation}
-        Putting this together, we can finally write,
-        \begin{equation}
-            \label{eq:sam_final}
-            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = - \Phi(M,z) \, F(M,a,z) \frac{\thardf}{\tau(M,a,z|a')} \diffp{V_c}{z}.
-        \end{equation}
+    \input{sam_text}
 

### 2022-07-19 docs/holodeck-paper/src/sam.tex

--- 2022-07-19 docs/holodeck-paper/src/ms.tex
@@
-% =================================================================================================
-% Holodeck Paper Manuscript
-% -------------------------
-%
-%
-%
-% =================================================================================================
-
 % general package imports and setup
@@
 
-\newtotcounter{citnum} % From the package documentation
-\def\oldbibitem{} \let\oldbibitem=\bibitem
-\def\bibitem{\stepcounter{citnum}\oldbibitem}
-
-\newcommand{\orcidauthorA}{0000-0002-6625-6450} % For author A
-
-\title[Short Title]{Very long and verbose title}
-\author[NANOGrav]{the NANOGrav collaboration}
-% \author[L.Z.~Kelley et al.]{
-%     Luke Zoltan Kelley$^{1}$\thanks{E-mail:lzkelley@northwestern.edu}\orcidA{} \\
-%     $^{1}$ \begin{minipage}[t]{\linewidth} Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA), and\\Department of Physics \& Astronomy, Northwestern University, Evanston, IL 60208
-%     \end{minipage}
-% }
-
 \begin{document}
 
-\maketitle
-
-\begin{abstract}
-    abstract
-\end{abstract}
-
-\begin{keywords}
-    quasars: supermassive black holes
-\end{keywords}
-
-% =================================================================================================
-% ====    Section 1 - Introduction
-% =================================================================================================
-
-\begin{itemize}
-    \item Change $f_r \rightarrow f$
-    \item holodeck uses $f_\trt{obs}$ as grid variable, not $f_r$, mesh this with the analytic formalism and description
-\end{itemize}
-
-\section{Introduction}
-    \label{sec:intro}
-
-% =================================================================================================
-% ====    Section 2 - Methods
-% =================================================================================================
-
-\section{Methods}
-    \label{sec:meth}
+% \maketitle
 
 \input{sam_text}
-
-    % % Fig -
-    % \begin{figure}
-    %     \centering
-    %     \includegraphics[width=1\columnwidth]{{{example-image-c}}}
-    %     \caption{\textbf{header.} body.}
-    %     \label{fig:fig1}
-    % \end{figure}
-    %
-    %
-    % \begin{equation}\begin{split}
-    %     \label{eq:eq1}
-    %     2+2 & = 4 \\
-    %         & = 10 - 6
-    % \end{split}\end{equation}
-    %
-    % % Fig -
-    % \begin{figure*}
-	% 	\centering
-    %     \subfloat[first]{{
-    %         \includegraphics[width=1.75\columnwidth]{{{example-image-a}}}
-    %     }}
-	% 	\\
-    %     % \qquad
-    %     \subfloat[second]{{
-    %     	\includegraphics[width=1.75\columnwidth]{{{example-image-b}}}
-    %     }}
-    %     \caption{\textbf{Head.}  Body.}
-    %     \label{fig:fig2}
-    % \end{figure*}
-
-\section{Implementation}
-    \label{sec:imp}
-
-    The \holodeck{} framework is written primarily in \python{} using a class-based structure for modularity and extensibility.
-
-    \subsection{Semi-Analytic Models}
-        \label{sec:imp_sam}
-
-        The core of the semi-analytic modeling framework in \holodeck{} is implementing \eqref{eq:sam_final}.
-There are classes for each of the key components: The galaxy stellar-mass function (GSMF, $\Phi$), the galaxy pair fraction \(F\), the merger timescale ($\tau$), and additional a scaling relation for mapping host galaxies via to MBH properties --- typically via an \mmbulge{} relation.
-Each component class is implemented generically via basic API methods (e.g.~a GSMF that returns the number density of galaxies at a given stellar-mass and redshift), and can be subclassed to provide a particular implementation (e.g.~an analytic Schechter-function GSMF).
-The number (or number density) of binaries is specified over a grid of parameters, typically in a four dimensional space of total mass ($M$), mass ratio ($q$), redshift ($z$), and binary rest-frame orbital frequency ($f_r$).
-
-        The semi-analytic model formalism defines a continuous distribution of MBH binaries.
-For many calculations, the discrete nature of MBHB sources is important, not only for examining individual continuous wave sources but also because of the important effects of discrete sources on the shape of the GWB due to the non-linear scaling of GW strain \citep{Sesana+2008}.
-Additional, a component of the uncertainty in expected GW signals is due to `cosmic variance', which can be treated as Poisson noise when creating multiple samples (or `realizations') of our binary populations.
-To handle discretization and resampling we use the \texttt{kalepy} package \citep{kalepy2021}, in particular, a method of outlier sampling to capture regions of parameter space where finite number effects are important.
-The \holodeck{} semi-analytic models specify the number (or number-density) of binaries over a grid, i.e.~$N_{ijkl}$.
-By definition, grid cells which have an expected total number of binaries $\gg 1$ will not produce appreciable finite-number effects, so instead they are given a weighting based on a Poisson distribution centered around their expectation value, i.e.~$\mathcal{P}(N_{ijkl})$.
-Grid cells with expected numbers of order unity or less do produce important finite-number effects, and thus are directly resampled
-
-% =================================================================================================
-% ====    Section 3 - Results
-% =================================================================================================
-\section{Results}
-    \label{sec:res}
-
-    % Fig -
-    % \begin{figure*}
-    %     \centering
-    %     \includegraphics[width=1.5\columnwidth]{{{figs/det-nums_fedd-1.00_voff+3.00_dvel+2.00_tobs+0.70}}}
-    %     \caption{\textbf{Distributions of parameters for all binaries (black) and those with a detectable kinematic signature in the secondary AGN (colors).}  Less than $10^{-7}$ of primaries are observable, which are not shown.
-Secondary AGN with observable velocity offsets ($\voff$) are shown in blue assuming a sensitivity of \mbox{$\voffsens = 10^3 \kmps$}.
-Secondaries with observable changing velocities ($\dvel$) are shown in orange (solid) for a sensitivity $\dvelsens = 10^2 \kmps$, and observing baseline of \mbox{$\tobs = 5\, \yr$}.
-Systems where both offsets and changes are detectable (orange, dotted) are mostly a uniform subset of the $\dvel$ sample, and are almost indistinguishable in the 2D contour plots.
-An optical flux cutoff of \mbox{$6\E{-14} \mathrm{erg/s/cm}^2$} (AB Mag $\approx 21$) is also imposed, where all AGN are assumed to be accreting at an Eddington fraction of $0.1$.}
-    %     \label{fig:det_bins_num}
-    % \end{figure*}
-
-% =================================================================================================
-% ====    Section 4 - Conclusions
-% =================================================================================================
-\section{Discussion \& Conclusions}
-    \label{sec:disc}
-
-% =================================================================================================
-% ====    End Matter
-% =================================================================================================
-
-% ====    Acknowledgments
-
-\section*{Acknowledgments}
-	I am very thankful to.
-
-    This research made use of \texttt{astropy}, a community-developed core Python package for Astronomy \citep{astropy2013}, in addition to \texttt{Scipy}~\citep{scipy}, \texttt{ipython}~\citep{ipython}, \texttt{jupyter}~notebook~\citep{jupyter}, \texttt{Numpy}~\citep{numpy2011} \& \texttt{SymPy}~\citep{sympy2017}.
-All figures were generated using \texttt{matplotlib}~\citep{matplotlib2007}.
-Kernel density estimation was performed using the \texttt{kalepy}{} package (\href{https://github.com/lzkelley/kalepy}{github.com/lzkelley/kalepy}) \citep{kalepy2020}.
-
-    The Illustris data is available online at \href{https://www.illustris-project.org/}{www.illustris-project.org} \citep{Nelson+2015}, and Illustris-TNG data at \href{https://www.tng-project.org/}{www.tng-project.org} \citep{Nelson+2019}.
 
@@
 
-    \section{Additional Material}
-        \label{sec:app}
-
-        \noindent\begin{minipage}{\linewidth}
-
-            % Fig -
-            % \centering
-            % \includegraphics[width=0.55\columnwidth]{{{figs/doan+2019_jitter}}}
-            % \captionof{figure}{Distribution of velocity-offset variations (`jitters') from \citep{Doan+2019}, for the red and blue components of the observed BLs.
-Measurements are shown as ticks at $y=0$, and KDE distributions are shown using a Scott's factor bandwidth, and Gaussian kernel.
-The purple curve is the KDE distribution taking both red and blue data as independent.
-The KDE $50\%$ (median) and $99\%$ (Gaussian standard deviation $\sigma \approx 2.3$) jitter for each component are shown with dashed vertical lines and corresponding labels.}
-            % \label{fig:jitter_doan2019}
-			%
-            % \setlength{\tabcolsep}{8pt}
-            % \begin{tabular}{p{1.5cm} c c c c c c c}
-            %      & & & $5\%$ & $25\%$ & $50\%$ & $75\%$ & $95\%$ \\ [0.5ex]
-            %     \hline
-            %     \multirow{4}{*}{All}  & \multirow{4}{*}{-}
-            %         & $M \; [M_\odot]$ & $3.4 \times10^{ 6 }$ & $7.2 \times10^{ 6 }$ & $1.8 \times10^{ 7 }$ & $6.0 \times10^{ 7 }$ & $4.1 \times10^{ 8 }$ \\
-            %         & & q & $2.9 \times10^{ -2 }$ & $1.7 \times10^{ -1 }$ & $4.0 \times10^{ -1 }$ & $6.8 \times10^{ -1 }$ & $9.3 \times10^{ -1 }$ \\
-            %         & & z & 0.24 & 0.41 & 0.53 & 0.62 & 0.69 \\
-            %         & & $p \; [\mathrm{yr}]$ & $1.4 \times10^{ 2 }$ & $1.5 \times10^{ 3 }$ & $9.6 \times10^{ 3 }$ & $4.0 \times10^{ 4 }$ & $2.0 \times10^{ 5 }$ \\
-            %         & & $a \; [\mathrm{pc}]$ & $2.1 \times10^{ -2 }$ & $1.3 \times10^{ -1 }$ & $4.9 \times10^{ -1 }$ & $1.3 \times10^{ 0 }$ & $4.1 \times10^{ 0 }$ \\
-            %         & & $a \; [r_g]$ & $1.2 \times10^{ 4 }$ & $1.4 \times10^{ 5 }$ & $4.5 \times10^{ 5 }$ & $1.2 \times10^{ 6 }$ & $4.1 \times10^{ 6 }$ \\
-            %     \hline
-            %     \multirow{4}{1.5cm}{Secondary\newline Offset\newline($\voff$)} & \multirow{4}{*}{$0.49\%$}
-            %         & $M \; [M_\odot]$ & $1.4 \times10^{ 8 }$ & $4.0 \times10^{ 8 }$ & $7.6 \times10^{ 8 }$ & $1.4 \times10^{ 9 }$ & $3.7 \times10^{ 9 }$ \\
-            %         & & q & $9.6 \times10^{ -4 }$ & $4.2 \times10^{ -3 }$ & $1.2 \times10^{ -2 }$ & $3.6 \times10^{ -2 }$ & $1.5 \times10^{ -1 }$ \\
-            %         & & z & 0.25 & 0.42 & 0.54 & 0.63 & 0.69 \\
-            %         & & $p \; [\mathrm{yr}]$ & $8.3 \times10^{ 2 }$ & $1.5 \times10^{ 3 }$ & $2.3 \times10^{ 3 }$ & $3.7 \times10^{ 3 }$ & $9.0 \times10^{ 3 }$ \\
-            %         & & $a \; [\mathrm{pc}]$ & 0.22 & 0.41 & 0.58 & 0.87 & 1.82 \\
-            %         & & $a \; [r_g]$ & $5.0 \times10^{ 3 }$ & $1.1 \times10^{ 4 }$ & $1.8 \times10^{ 4 }$ & $2.9 \times10^{ 4 }$ & $4.9 \times10^{ 4 }$ \\
-            % \end{tabular}
-            % \captionof{table}{\textbf{Parameters of detectable binary systems with redshift $z < 0.7$.}  The indicated quantiles are given for total mass ($M$), mass ratio ($q$), separation ($a$), and orbital period ($p$).
-In only four binaries ($2\E{-8}$ of systems) is the primary detectable, and those have parameters: \mbox{$M\approx 2$--$5\E{9} \, \msol$},
-            % \hspace{0.5ex} \mbox{$q \approx 0.7$--$0.9$}, \hspace{0.5ex} \mbox{$z \approx 0.6$--$0.8$}, \hspace{0.5ex} \mbox{$a \approx 2.2$--$2.8 \, \pc$}.}
-            % \label{tab:obs_pars}
-
-        \end{minipage}
-
     \twocolumn

### 2022-07-19 setup.py
Try allow python3.7 to see if that works

@@ -43,6 +43,6 @@ setup(
 
     # Additional entries you may want simply uncomment the lines you want and fill in the data
     # url='http://www.my_package.com',  # Website
-    python_requires=">=3.8",          # Python version restrictions
+    python_requires=">=3.7",          # Python version restrictions
 
 )

### 2022-07-19 scripts/run_tox.sh
Enable tests to work on py37, they are currently succeeding.

@@ -1,27 +1,0 @@
-#!/bin/zsh
-#
-# `pyenv` can be installed with homebrew, e.g. `brew install pyenv`
-# NOTE: the versions of python (in `VERSIONS`) need to match those in the `tox.ini` file's `envlist` variable
-#
-
-set -e
-VERSIONS=("3.7:latest" "3.8:latest" "3.9:latest" "3.10:latest")
-
-# Iterate the string array using for loop
-for val in ${VERSIONS[@]}; do
-   echo $val
-   pyenv install -s $val
-done
-
-# versions=("${(@f)$(pyenv versions)}")
-# versions=("${versions[@]:1}")
-# echo ${versions}
-# clean=()
-# for val in ${versions[@]}; do
-#    val=${val// }
-#    clean+=($val)
-# done
-# echo $clean
-# pyenv local ${clean[@]}
-
-tox

### 2022-07-26 holodeck/population.py
DOCS: added getting started material, abbrevs

@@ -194,7 +194,7 @@ class Pop_Illustris(_Population_Discrete):
 
     Takes as input a data file that includes BH and subhalo data for BH and/or galaxy mergers.
 
-    NOTES
+    Notes
     -----
     * Parameters required in input hdf5 file:
         * `box_volume_mpc`:

### 2022-07-26 holodeck/relations.py
DOCS: cleanup references, more 'relations.py' references to 'biblio.rst'

@@ -1,34 +1,17 @@
 """Holodeck - Scaling Relations
 
-References
-----------
-* Kormendy+Ho-2013 (1304.7762) = [KH13]
-  Coevolution (Or Not) of Supermassive Black Holes and Host Galaxies
-  https://ui.adsabs.harvard.edu/abs/2013ARA%26A..51..511K/abstract
-
-* McConnell+Ma-2013 (1211.2816) = [MM13]
-  Revisiting the Scaling Relations of Black Hole Masses and Host Galaxy Properties
-  https://ui.adsabs.harvard.edu/abs/2013ApJ...764..184M/abstract
-
-* [NFW-97] : Navarro, Frenk & White 1997
-  A Universal Density Profile from Hierarchical Clustering
-  https://ui.adsabs.harvard.edu/abs/1997ApJ...490..493N/abstract
-
-* [Guo-2010] : Guo, White, Li & Boylan-Kolchin 2010
-  How do galaxies populate dark matter haloes?
-  https://ui.adsabs.harvard.edu/abs/2010MNRAS.404.1111G/abstract
-
-* [Behroozi+2013] : Behroozi, Wechsler & Conroy 2013
-  The Average Star Formation Histories of Galaxies in Dark Matter Halos from z = 0-8
-  https://ui.adsabs.harvard.edu/abs/2013ApJ...770...57B/abstract
-
-* [Klypin+2016] : Klypin et al. 2016
-  MultiDark simulations: the story of dark matter halo concentrations and density profiles
-  https://ui.adsabs.harvard.edu/abs/2016MNRAS.457.4340K/abstract
-
 To-Do
 -----
 * Pass concentration-relation (or other method to calculate) to NFW classes on instantiation
+
+References
+----------
+* [Behroozi2013]_ : Behroozi, Wechsler & Conroy 2013.
+* [Guo2010]_ Guo, White, Li & Boylan-Kolchin 2010.
+* [Klypin2016]_ Klypin et al. 2016.
+* [KH13]_ Kormendy & Ho 2013.
+* [MM13]_ McConnell & Ma 2013.
+* [NFW97]_ Navarro, Frenk & White 1997.
 
 """
 

### 2022-07-26 holodeck/sam.py

@@ -1,12 +1,5 @@
 """Holodeck - Semi Analytic Modeling submodule
 
-References
-----------
-* Chen, Sesana, Conselice 2019 = [Chen19]
-  Constraining astrophysical observables of galaxy and supermassive black hole binary mergers
-  using pulsar timing arrays
-  https://ui.adsabs.harvard.edu/abs/2019MNRAS.488..401C/abstract
-
 To-Do
 -----
 * Check that _GW_ frequencies and _orbital_ frequencies are being used in the correct places.
@@ -17,6 +10,10 @@ To-Do
   the log(dens) instead of just `dens`.
 * Should there be an additional dimension of parameter space for galaxy properties?  This way
   variance in scaling relations is incorporated directly before calculating realizations.
+
+References
+----------
+* [Chen2019]_ Chen, Sesana, Conselice 2019
 
 """
 

### 2022-07-26 docs/holodeck-paper/src/sam_text.tex
DOCS: start adding SAM description

@@
 
-    We consider the standard \citep[e.g.~][]{Sesana+2008, Chen+2019} semi-analytic model (SAM) formalism of MBH binary populations.
+    We consider the standard semi-analytic model (SAM) formalism of MBH binary populations \citep[e.g.~][]{Sesana+2008, Chen+2019}.
 In this style of calculation, $F$ is determined in a region of parameter space that can be observed/estimated, and this is evolved to find the distribution in a different region of parameter space that is of interest.
-In practice, the observed parameter space is galaxies and galaxy mergers, and the parameter space of interest is close separated MBH binaries that could be GW detectable.
-Thus we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts), and we set $S_{\!F}$.
+In practice, the observed parameter space is galaxies and galaxy mergers, and the parameter space of interest is closely separated MBH binaries that could be GW detectable.
+Thus we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts), and we set $S_{\!F} = 0$.
 
@@
 We assume that there is a one-to-one mapping from galaxy mass to MBH mass, such that the galaxy mass-function can still be used to uniquely define the mass distribution of MBHs.
-Typically the MBH--galaxy relation is given in terms of an `$M_\trt{BH}$--$M_\trt{bulge}$' relation \citep[e.g.][]{McConnell+Ma-2013}, which is a observationally derived relation between the mass of MBHs and the mass of the stellar bulge component of their host galaxy.
+Typically the MBH--galaxy relation is given in terms of an `$M_\trt{BH}$--$M_\trt{bulge}$' relation \citep[e.g.][]{Kormendy+Ho-2013}, which is an observationally derived relation between the mass of MBHs and the mass of the stellar bulge component of their host galaxy.
     In \citet{Chen+2019}, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$P = \int_{a_0}^{a_1} P_a \, da$.

### 2022-07-26 holodeck/population.py
DOCS: docstrings for 'relations.py'

@@ -488,7 +488,7 @@ class PM_Mass_Reset(Population_Modifier):
 
         Parameters
         ----------
-        mhost : class or instance of `holodeck.relations._MHost_Relation`
+        mhost : class or instance of `holodeck.relations._Host_Relation`
             The Mbh-MHost scaling relationship with which to reset population masses.
         scatter : bool, optional
             Include random scatter when resetting masses.
@@ -496,10 +496,10 @@ class PM_Mass_Reset(Population_Modifier):
 
         """
         # if `mhost` is a class (not an instance), then instantiate it; make sure its a subclass
-        # of `_MHost_Relation`
-        mhost = utils._get_subclass_instance(mhost, None, holo.relations._MHost_Relation)
+        # of `_Host_Relation`
+        mhost = utils._get_subclass_instance(mhost, None, holo.relations._Host_Relation)
         # store attributes
-        self.mhost = mhost         #: Scaling relationship between host and MBH (`holo.relations._MHost_relation`)
+        self.mhost = mhost         #: Scaling relationship between host and MBH (`holo.relations._Host_Relation`)
         self._scatter = scatter    #: Bool determining whether resampled masses should include statistical scatter
         return
 

### 2022-07-26 holodeck/evolution.py
DOCS: docstrings for 'relations.py'

@@ -24,12 +24,12 @@ To-Do
 
 References
 ----------
-* [Quinlan96]_ Quinlan 1996
-* [SHM06]_ Sesana, Haardt & Madau et al. 2006
-* [BBR80]_ Begelman, Blandford & Rees 1980
-* [Sesana10]_ Sesana 2010
-* [Kelley17]_ Kelley, Blecha & Hernquist 2017
-* [Chen17]_ Chen, Sesana, & Del Pozzo 2017
+* [Quinlan96]_ Quinlan 1996.
+* [Sesana2006]_ Sesana, Haardt & Madau et al. 2006.
+* [BBR80]_ Begelman, Blandford & Rees 1980.
+* [Sesana2010]_ Sesana 2010.
+* [Kelley17]_ Kelley, Blecha & Hernquist 2017.
+* [Chen2017]_ Chen, Sesana, & Del Pozzo 2017.
 
 """
 
@@ -655,7 +655,7 @@ class Hard_GW(_Hardening):
 class Sesana_Scattering(_Hardening):
     """Binary-Hardening Rates calculated based on the Sesana stellar-scattering model.
 
-    This module uses the stellar-scattering rate constants from the fits in [SHM06]_ using the
+    This module uses the stellar-scattering rate constants from the fits in [Sesana2006]_ using the
     `_SHM06` class.  Scattering is assumed to only be effective once the binary is bound.  An
     exponential cutoff is imposed at larger radii.
 
@@ -757,7 +757,7 @@ class Dynamical_Friction_NFW(_Hardening):
     Attenuation of the DF hardening rate is typically also included, to account for the inefficiency of DF once the
     binary enters the hardened regime.  This is calculated using the prescription from [BBR80]_.  The different
     characteristic radii, needed for the attenuation calculation, currently use a fixed Dehnen stellar-density profile
-    as in [Chen17]_, and a fixed scaling relationship to find the characteristic stellar-radius.
+    as in [Chen2017]_, and a fixed scaling relationship to find the characteristic stellar-radius.
 
     This module does not evolve eccentricity.
 
@@ -1467,7 +1467,9 @@ def _get_stellar_mass_halo_mass_relation(smhm=None):
 
 
 def _radius_stellar_characteristic_dabringhausen_2008(mstar, gamma=1.0):
[…]

### 2022-07-26 holodeck/relations.py

@@ -1,4 +1,4 @@
-"""Holodeck - Scaling Relations
+"""Empirical and phenomenological scaling relationships.
 
 To-Do
 -----
@@ -24,11 +24,13 @@ from holodeck import cosmo, utils, log
 from holodeck.constants import MSOL, NWTG, KMPERSEC
 
 
-class _MHost_Relation(abc.ABC):
+class _Host_Relation(abc.ABC):
+    """Base class for general relationships between MBHs and their host galaxies.
+    """
 
-    _PROPERTIES = []
+    _PROPERTIES = []    #: list of property names to retrieve from population instances.
 
-    def get_host_properties(self, pop, copy=True):
+    def get_host_properties(self, pop, copy=True) -> dict:
         """Get the host properties specified in the `_PROPERTIES` list of variable names.
 
         NOTE: if the `copy` flag is False, then values are returned by reference and the original
@@ -69,6 +71,8 @@ class _MHost_Relation(abc.ABC):
     def mbh_from_host(self, host):
         """Convert from host galaxy properties to blackhole mass.
 
+        BUG/FIX: IS THIS A GOOD INTERFACE??
+
         Units of [grams].
         """
         return
@@ -79,39 +83,107 @@ class _MHost_Relation(abc.ABC):
 # -----------------------------------------
 
 
-class _MMBulge_Relation(_MHost_Relation):
+class _MMBulge_Relation(_Host_Relation):
+    """Base class for implementing Mbh--Mbulge relationships, between MBH and their host galaxies.
+    """
 
     _PROPERTIES = ['mbulge']
 
     def mbulge_from_mstar(self, mstar):
+        """Calculate stellar bulge-mass given the total galaxy stellar mass.
+
+        Multiplies `mstar` by the result of `self.bulge_mass_frac(mstar)`.
+
+        Parameters
+        ----------
+        mstar : array_like,
+            Galaxy total stellar mass.
+
+        Returns
+        -------
+        array_like,
+            Stellar bulge mass.
+
+        """
         return self.bulge_mass_frac(mstar) * mstar
 
     # ---- Abstract Methods : must be overridden in subclasses  ----
 
     @abc.abstractmethod
-    def dmstar_dmbh(self, pop):
+    def dmstar_dmbh(self, mstar):
+        """The partial derivative of stellar mass versus black-hole mass.
+
+        Parameters
+        ----------
+        mstar : array_like,
[…]

### 2022-11-15 docs/holodeck-paper/ms.tex

--- 2022-07-19 docs/holodeck-paper/src/ms.tex
@@
 % =================================================================================================
-% Holodeck Paper Manuscript
-% -------------------------
+% Holodeck Methods Paper - Manuscript
+% -----------------------------------
 %
@@
 
+% =================================================================================================
+% ====   Header & Setup
+% =================================================================================================
+
 % general package imports and setup
-\input{packages}
+\input{src/packages}
 % definitions of custom commands
-\input{commands}
+\input{src/commands}
 % journal abbreviations for bibliography
-\input{abbrevs}
+\input{src/abbrevs}
 
@@
 
+    Massive black holes (MBHs) are known to occupy the centers of most if not all massive galaxies \needcite{}.
+    Binaries of MBHs are believed to form following the merger of massive galaxies \needcite{}.
+Such MBH Binaries (MBHBs) could then produce low-frequency gravitational waves (GWs) \needcite{LF-GWs}, in the nano-Hertz regime, that would soon be detectable by experiments called pulsar timing arrays (PTAs) \needcite{PTAs}.
+In this paper, we introduce the MBHB population synthesis code called \holodeck{}.
+The goal of this package is to provide a cohesive and comprehensive framework for:
+    \begin{enumerate}
+        \item
+        modeling the physics of MBHB formation and evolution while accurately capturing the true uncertainties in this modeling;
+        \item
+        rapidly generating large simulated populations of MBHBs;
+        \item
+        calculating the resulting GW signals from these populations.
+    \end{enumerate}
+    % (1) modeling the physics of MBHB formation and evolution while accurately capturing the true uncertainties in this modeling, (2) rapidly generating large simulated populations of MBHBs, and (3) calculating the resulting GW signals from these populations.
+
+    As in the case of high-frequency GWs detected by LIGO-Virgo \needcite{}, tremendous uncertainties exist in the physics of the formation and evolution of the binary sources.
+For individual MBHs, there are numerous, well-known scaling relationships between their properties and those of their host galaxies \needcite{MBH-Host relations}, but numerous questions remain such as where or not the observed sample of MBHs (i.e.~quasars and active galactic nuclei, AGN) accurately reflect the full population \needcite{MBH bias}, and how reliably these relationships extend to different host galaxies and redshift ranges.
+
+    For binary MBHs in particular, a variety of physical processes are required for the two MBHs to find the center of the post-merger host galaxy and eventually become gravitationally bound \needcite{BBR1990}.
+In particular, the two MBHs must interact with their host galaxy environments through dynamical friction, stellar scattering, accretion disk torques, and possibly three-body interactions with a third MBH component \needcite{}, before reaching small enough binary separations for GW emission to eventually dominate the final coalescence of the binary.
+Besides GW emission, none of these processes can be modeled exactly, necessitating the usage of simplified prescriptions which each containing numerous parametric uncertainties.
+The resolution of cosmological hydrodynamic simulations is far too coarse to resolve the MBH binary phase, with some simulations instantaneously merging nearby MBHs \needcite{illustris}, and others using sub-grid models to `resolve' down to the scale of $\sim kpc$ \needcite{Romulus/Tremmel, Ma+Hopkins}.
+
+    In addition to the theoretical uncertainties, there exists a substantial dearth of observational constraints.
+A growing number of \textit{candidate} MBHB systems have been identified through varying electromagnetic (EM) indicators in AGN \needcite{}, but a substantial fraction may be false positives \needcite{}.
+To date, no gravitationally bound MBH binary has yet to be confirmed, although examples exist of dual AGN at close projected separation \needcite{Rodriguez et al.
+2007}, and single AGN showing time-variability highly suggestive of a binary companion \needcite{OJ~287}.
+
+    GWs from MBHBs, particularly in the form of a stochastic gravitational wave background (GWB), have been forecast for decades.
+Only recently have the sensitivities of PTA experiments begin to encroach on astrophysically plausible GWB amplitudes \needcite{PTA limits} \needcite{c.f.~Shannon et al.}.
+Indeed, PTA upper-limits have already been used to place constraints on limited sub-spaces of MBHB parameters \needcite{11yr analysis, etc}.
+In the 12.5yr dataset of the North-American nano-Hertz Observatory for Gravitational waves \needcite{NANOGrav}, a `common process' signal was identified that is consistent with a GWB but not yet statistically definitive.
+Followup modelling suggests that the GW nature of the signal can be determined within the next few years \needcite{Pol}.
+While the interpretation is currently entirely conjectural, the two currently measurable parameters of the common process, the GW-strain amplitude and strain spectral index, each challenge the standard expectations of theoretical models.
+The \todo{possible} amplitude is larger than many models have predicted, and the spectral index significantly flatter than the canonical `-2/3' power-law \needcite{phinney} \needcite{c.f.~Sesana, Kocsis}.
+Still, both of these parameters can easily be fit with current models, and indeed, studies have already performed MBHB model inference using the amplitude and spectral index of the candidate \todo{(what's the word besides `candidate'?)} signal \needcite{middleton}.
+
+    Studies which have used PTA observations to constrain MBHB models have been focused on demonstrating the feasibility of performing such inference, and thus have relied on single classes of models (semi-analytic models, SAMs) that make significant simplifications of the physical processes, and only explore small sub-spaces of astrophysical parameters.
+The \holodeck{} framework, presented here, is structured to more fully capture the astrophysical uncertainties underlying our understanding of MBHB formation and evolution.
+\holodeck{} uses comprehensive and self-consistent models for binary evolution, in addition to allowing SAMs to be interchanged with populations derived from cosmological hydrodynamic simulations.
+In this way, \holodeck{} is designed to capture both statistical and systematic uncertainties by allowing the marginalization over not only specific astrophysical parameters, but also entire modeling methodologies.
+
+    In this paper, we present the simulation methodologies and their implementations in \holodeck{}.
+We also use the framework to explore in more detail the range of plausible GW-strain spectral indices that would be detectable by PTAs.
+\todo{expand this paragraph}.
+
 % =================================================================================================
@@
 
-    \input{sam_text}
+    \subsection{Source Distributions and Gravitational Wave Calculations}
+
+    Let $N(z)$ denote the number of binaries out to a redshift $z$, corresponding to a comoving volume $V_c$.
+    We can calculate the characteristic strain from an ensemble of GW sources as \citep[][Eqs.~5/8]{Phinney-2001},
+        \begin{align}
+            h_c^2(f) = \int_0^\infty \!\! dz \; \frac{d^2 N}{dz \, d\ln f_r} \; h_s^2\lr{f_r}.
+        \end{align}
+    In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$, extrapolated into an observer's entire past light-cone.
+These can be related as \citep[][Eq.~6]{Sesana+2008},
+        \begin{align}
+            \label{eq:num_num_dens}
+            \frac{d^2 N}{dz \, d\ln f_r} = \frac{d n_c}{dz} \frac{dz}{dt} \frac{dt}{d\ln f_r} \frac{d V_c}{dz}.
+        \end{align}
+    Note that the total number of binaries in a given frequency interval depends not only on the number density, but also their rate of frequency-evolution.
+The faster binaries evolve through a given frequency interval, the fewer binaries will occupy that interval on average.
+This is typically described by the binary `hardening' timescale (or `residence time'), given by the expression,
+    \begin{equation}
+        \thardf \equiv - \diffp{t}{{\ln\!f_r}} = - \frac{f_r}{\partial f_r / \partial t}.
+    \end{equation}
+    Plugging in the relevant cosmographic relationships (see \secref{sec:app_eqs}) we have,
+        \begin{align}
+            h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
+        \end{align}
+    The challenge in modeling astrophysical sources of GWs is in calculating their number density.  Numerous approaches are possible, described in the following sections.
+
+    \subsection{Semi-Analytic Modeling}
+
+    Define the distribution function of sources as $F(M,q,a,z) = d^2 n(M,q,a,z) / dM dq$.
+Here $M$ is the total mass of each systems, the mass-ratio is $q\equiv m_2/m_1 \leq 1$, $a$ is the binary separation, and $z$ is the redshift.
+    % Define the distribution function of sources as $F(M,q,f_r,z) = d^2 n(M,q,f_r,z) / dM dq$.
+Here $M$ is the total mass of each systems, the mass-ratio is $q\equiv m_2/m_1 \leq 1$, $f_r$ is the binary orbital frequency in the rest frame, and $z$ is the redshift.
+    We can write the conservation equation for binaries as of function of redshift as,
+    \begin{equation}
+        \label{eq:conservation}
+        \diffp{F}{z} +
+            \diffp{}{M} \lrs{F \diffp{M}{z}} +
+            \diffp{}{q} \lrs{F \diffp{q}{z}} +
+            \diffp{}{a} \lrs{F \diffp{a}{z}} = S_{\!F}(M, q, a, z).
+    \end{equation}
+    Here $S_{\!F}$ is a source/sink function that can account for the creation or destruction of binaries.
+
+    We consider the standard semi-analytic model (SAM) formalism of MBH binary populations \citep[e.g.~][]{Sesana+2008, Chen+2019}.
+In this style of calculation, $F$ is determined in a region of parameter space that can be observed/estimated, and this is evolved to find the distribution in a different region of parameter space that is of interest.
+In practice, the observed parameter space is galaxies and galaxy mergers, and the parameter space of interest is closely separated MBH binaries that could be GW detectable.
+Thus we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts), and we set $S_{\!F} = 0$.
+
+    We will express the distribution function as a product of a mass function, and a pair fraction:
+    \begin{equation}
+        \label{eq:dist_func}
+        F(M,q,z) = \frac{\Phi(M, z)}{M \ln\!10} \cdot P(M,q,z),
+    \end{equation}
+    where the mass function, $\Phi(M, q, z) \equiv \diffp{n_g}{{\logten{M}}}$, is calculated based on the number density of galaxies ($n_g$).
+We assume that there is a one-to-one mapping from galaxy mass to MBH mass, such that the galaxy mass-function can still be used to uniquely define the mass distribution of MBHs.
+Typically the MBH--galaxy relation is given in terms of an `$M_\trt{BH}$--$M_\trt{bulge}$' relation \citep[e.g.][]{Kormendy+Ho-2013}, which is an observationally derived relation between the mass of MBHs and the mass of the stellar bulge component of their host galaxy.
+    In \citet{Chen+2019}, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$P = \int_{a_0}^{a_1} P_a \, da$.
+
+    From $\eqref{eq:conservation}$, we use the chain rule to mix time and redshift evolution, and assume that the mass-change of binaries is negligible, i.e.~$\diffp{m}{t} = 0$ and $\diffp{q}{t} = 0$, giving:
+    \begin{equation}
+        \diffp{F}{z} = - \diffp{t}{z} \diffp{}{a} \lrs{F \diffp{a}{t}}.
+    \end{equation}
+
+    The binary population is assumed to be changing only in separation and redshift, which are related by $\partial a / \partial z = (\partial a / \partial t) (\partial t / \partial z)$.
+Because the overall number-density is conserved, we can take a finite step in separation and time/redshift, $a\rightarrow a'$ and $z\rightarrow z'$.
+Here the time it takes for a binary to go from $a \rightarrow a'$ is $T(M,q,a,z|a')$, which leads to a redshift at the later time of $z' = z'(t + T)$.
+So far, we have left the binary separation $a$ as implicit in the expression for F.
+To obtain the standard expression \citep[e.g.][~Eq.~5]{Chen+2019}, we make the approximation that,
+    \begin{equation}
+        \diffp{}{a} \lrs{F(M,q,z) \diffp{a}{t}} \approx \frac{F}{T(M,q,a,z|a')}.
+    \end{equation}
+    Thus giving,
+    \begin{equation}
+        \label{eq:cont_eq_result}
+        \diffp{F(M,q,a',z')}{{z'}} = \diffp{n}{{M}{q}{z'}} = - \diffp{t}{z} \frac{\Phi(M,z) \, P(M,q,z)}{T(M,q,a,z|a')}.
+    \end{equation}
+
+    Combining \eqref{eq:cont_eq_result} with \eqref{eq:num_num_dens}, we can finally write,
+    \begin{equation}
+        \label{eq:sam_final}
+        \diffp{N}{{M}{q}{z}{\ln\!f_r}} = \frac{\Phi(M,z) \, P(M,q,z)}{T(M,q,a,z|a')} \, \thardf \, \diffp{V_c}{z}.
+    \end{equation}
+
+    % From the SAM formalism, the number of binaries in each bin is calculated by `integrating' each finite element of \eqref{eq:sam_final}, i.e.,
+    % \begin{equation}
+    %     \diffp{N_{ijk}(f_r)}{{\ln\!f_r}} \equiv \diffp{N(M_i, q_j, z_k, {f_r})}{{M}{q}{z}{\ln\!f_r}} \, \Delta M \,\Delta q \, \Delta z.
+    % \end{equation}
+    % The characteristic strain of the GWB can then be calculated
 

### 2022-11-15 holodeck/utils.py
Fixed spacing bug in nyquist_freqs() in utils.py. Updated gen_lib_sams.py to use current API.  Created gen_spec_libs.sams.py to create just GWB spectrum and not full population.

@@ -551,7 +551,7 @@ def nyquist_freqs(
 
     """
     fmin = 1.0 / dur
-    fmax = 1.0 / cad
+    fmax = 1.0 / cad * 0.5
     # df = fmin / sample
     df = fmin
     freqs = np.arange(fmin, fmax + df/10.0, df)

### 2022-11-15 scripts/gen_spec_lib_sams.py

@@ -1,7 +1,7 @@
 """
 
 
-mpirun -n 14 --use-hwthread-cpus python ./scripts/gen_lib_sams.py output/test_2022-06-27
+mpirun -n 14 --use-hwthread-cpus python ./scripts/gen_spec_lib_sams.py output/test_2022-06-27
 
 
 """
@@ -107,7 +107,7 @@ BEG = datetime.now()
 
 parser = argparse.ArgumentParser()
 parser.add_argument('output', metavar='output', type=str,
-                    help='output path [created if doesnt exist')
+                    help='output path [created if doesnt exist]')
 
 # parser.add_argument('-r', '--reals', action='store', dest='reals', type=int,
 #                     help='number of realizations', default=10)
@@ -211,7 +211,7 @@ def main():
         indices = np.array_split(indices, comm.size)
         num_ind_per_proc = [len(ii) for ii in indices]
         # LOG.info(f"{npars=}, {nreals=}, total={npars*nreals} || ave runs per core = {np.mean(num_ind_per_proc)}")
-        LOG.info(f"{npars=}, {nreals=} || ave runs per core = {np.mean(num_ind_per_proc)}")
+        LOG.info(f"{npars=}, {nreals=} || avg runs per core = {np.mean(num_ind_per_proc)}")
     else:
         indices = None
     indices = comm.scatter(indices, root=0)
@@ -264,27 +264,26 @@ def main():
 
 def run_sam(pnum, real, path_output):
 
-    iterator = range(args.NUM_REALS)
-    if comm.rank == 0:
-        iterator = holo.utils.tqdm(iterator, leave=False)
+#     iterator = range(args.NUM_REALS)
+#     if comm.rank == 0:
+#         iterator = holo.utils.tqdm(iterator, leave=False)
 
-    for real in iterator:
+    
+    fname = f"lib_sams__p{pnum:06d}.npz"
+    fname = os.path.join(path_output, fname)
+    if os.path.exists(fname):
+        LOG.warning(f"File {fname} already exists.")
 
-        fname = f"lib_sams__p{pnum:06d}_r{real:03d}.npz"
-        fname = os.path.join(path_output, fname)
-        if os.path.exists(fname):
-            LOG.warning(f"File {fname} already exists.")
-            continue
[…]

### 2022-11-18 scripts/gen_lib_sams_hard.py
now use nyquist_freqs_edges

@@ -280,11 +280,13 @@ def run_sam(pnum, real, path_output):
     if os.path.exists(fname):
         LOG.warning(f"File {fname} already exists.")
 
-    _fobs = holo.utils.nyquist_freqs(args.PTA_DUR, args.PTA_CAD)
-
-    df = _fobs[0]/2
-    fobs_edges = _fobs - df
-    fobs_edges = np.concatenate([fobs_edges, [_fobs[-1] + df]])
+#    _fobs = holo.utils.nyquist_freqs(args.PTA_DUR, args.PTA_CAD)
+#
+#    df = _fobs[0]/2
+#    fobs_edges = _fobs - df
+#    fobs_edges = np.concatenate([fobs_edges, [_fobs[-1] + df]])
+    fobs_centers=holo.utils.nyquist_freqs(args.PTA_DUR, args.PTA_CAD)
+    fobs_edges=holo.utils.nyquist_freqs_edges(args.PTA_DUR, args.PTA_CAD)
 
     sam, hard = SPACE.sam_for_number(pnum)
 #         hard = holo.evolution.Hard_GW()
@@ -292,7 +294,8 @@ def run_sam(pnum, real, path_output):
 #         gff, gwf, gwb = holo.gravwaves._gws_from_samples(vals, weights, fobs)
     gwbspec = sam.gwb(fobs_edges, realize=args.NUM_REALS, hard=hard)
     legend = SPACE.param_dict_for_number(pnum)
-    np.savez(fname, fobs=_fobs, fobs_edges=fobs_edges, gwbspec=gwbspec, pnum=pnum, nreals=args.NUM_REALS, **legend)
+#    np.savez(fname, fobs=_fobs, fobs_edges=fobs_edges, gwbspec=gwbspec, pnum=pnum, nreals=args.NUM_REALS, **legend)
+    np.savez(fname, fobs=fobs_centers, fobs_edges=fobs_edges, gwbspec=gwbspec, pnum=pnum, nreals=args.NUM_REALS, **legend)
     LOG.info(f"Saved to {fname} after {(datetime.now()-BEG)} (start: {BEG})")
 
     return

### 2022-11-19 holodeck/gravwaves.py
stuff

@@ -478,8 +478,36 @@ def _gws_from_number_grid_centroids(edges, dnum, number, realize):
 def _gws_from_number_grid_integrated(edges, dnum, number, realize, integrate=True):
     """
 
-    # ! BUG: THIS ASSUMES THAT FREQUENCIES ARE NYQUIST SAMPLED !
-    # ! otherwise the conversion from hs to hc doesnt work !
+    #! --------------------------- !#
+    #! CHECK GW vs ORB FREQUENCIES !#
+    #! --------------------------- !#
+
+    Parameters
+    ----------
+    edges : (4,) list of 1darrays
+        A list containing the edges along each dimension.  The four dimensions correspond to
+        total mass, mass ratio, redshift, and rest-frame orbital frequency.
+        The length of each of the four arrays is M, Q, Z, F.
+    dnum : (M, Q, Z, F) ndarray
+        The differential number of binaries per unit bin-volume.  This is,
+        ``d^4 N / [dlog10(M) dq dz dln(f_r)]``.
+    number : (M-1, Q-1, Z-1, F-1) ndarray
+        The number of binaries in each bin of parameter space.  This is calculated by integrating
+        `dnum` over each bin.
+    realize : bool or int,
+        Specification of how to construct one or more discrete realizations.
+        If a `bool` value, then whether or not to construct a realization.
+        If an `int` value, then how many discrete realizations to construct.
+    integrate : bool,
+        Whether or not to integrate (sum) over axes {0, 1, 2}.
+
+    Returns
+    -------
+    hc : ndarray
+        Characteristic strain of the GWB.
+        The shape depends on whether `integrate` is true or false.
+        integrate = True:  shape is (F-1,)
+        integrate = False: shape is (M-1, Q-1, Z-1, F-1)
 
     """
     grid = np.meshgrid(*edges, indexing='ij')
@@ -491,6 +519,40 @@ def _gws_from_number_grid_integrated(edges, dnum, number, realize, integrate=Tru
     fr = utils.frst_from_fobs(grid[3], grid[2])
     hs = utils.gw_strain_source(mc, dc, fr)
 
+    print(f"{hs.shape=} {fr.shape=} {np.shape(grid)=}  aiouh79fe3")
+
+    if realize is True:
[…]

### 2022-11-19 holodeck/sam.py

@@ -36,6 +36,7 @@ To-Do
 
 References
 ----------
+* [Sesana2008]_ Sesana, Vecchio, Colacino 2008.
 * [Chen2019]_ Chen, Sesana, Conselice 2019.
 
 """
@@ -483,6 +484,10 @@ class Semi_Analytic_Model:
         density : (M, Q, Z) ndarray of float
             Number density of binaries, per unit redshift, mass-ratio, and log10 of mass.  Units of [Mpc^-3].
 
+        Notes
+        -----
+        This function effectively calculates Eq.21 & 5 of [Chen2019]_.
+
         """
         if self._density is None:
             dens = np.zeros(self.shape)
@@ -550,11 +555,11 @@ class Semi_Analytic_Model:
     def dynamic_binary_number(self, hard, fobs=None, sepa=None, limit_merger_time=None):
         """Calculate the differential number of binaries (per bin-volume, per log-freq interval).
 
-        d^3 n / dlog10(M) dq dz ===> d^4 N / [dlog10(M) dq dz dln(f)
+        d^4 N / [dlog10(M) dq dz dln(X)    <===    d^3 n / dlog10(M) dq dz
 
         The value returned is `d^4 N / [dlog10(M) dq dz dln(X)]`, where X is either:
-        *   separation (a) ::  if  `sepa` is passed in, or
-        *   observer-frame orbital-frequency (f_r) :: if `fobs` is passed in.
+        *   separation (a)                     ::  if `sepa` is passed in, or
+        *   rest-frame orbital-frequency (f_r) ::  if `fobs` is passed in.
 
         Parameters
         ----------
@@ -562,8 +567,10 @@ class Semi_Analytic_Model:
             Hardening instance for calculating binary evolution rate.
         fobs : ArrayLike
             observer-frame orbital-frequency in [1/sec].
+            NOTE: either `fobs` or `sepa` must be provided (and not both).
         sepa : ArrayLike
             Rest-frame binary separation in [cm].
+            NOTE: either `fobs` or `sepa` must be provided (and not both).
         limit_merger_time : None or scalar,
             Maximum allowed merger time in [sec].  If `None`, no maximum is imposed.
 
@@ -576,12 +583,15 @@ class Semi_Analytic_Model:
 
[…]

### 2023-01-20 docs/holodeck-paper/ms.tex

@@
 
-    Let $N(z)$ denote the number of binaries out to a redshift $z$, corresponding to a comoving volume $V_c$.
+        Let $N(M,q,z,\frst)$ denote the number of binaries up to a total mass $M$, mass ratio $q$, redshift $z$, and rest-frame orbital-frequency $\frst$.
         We can calculate the characteristic strain from an ensemble of GW sources as \citep[][Eqs.~5/8]{Phinney-2001},
             \begin{align}
-            h_c^2(f) = \int_0^\infty \!\! dz \; \frac{d^2 N}{dz \, d\ln f_r} \; h_s^2\lr{f_r}.
+                h_c^2 = \int_0^\infty \!\! dz \; \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} \; \hs^2\lr{\frst}.
             \end{align}
-    In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$, extrapolated into an observer's entire past light-cone.
+        In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$, for some comoving volume $V_c$, extrapolated into an observer's entire past light-cone.
 These can be related as \citep[][Eq.~6]{Sesana+2008},
@@
                 \label{eq:num_num_dens}
-            \frac{d^2 N}{dz \, d\ln f_r} = \frac{d n_c}{dz} \frac{dz}{dt} \frac{dt}{d\ln f_r} \frac{d V_c}{dz}.
+                \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} = \frac{d^3 n_c}{dM \, dq \, dz} \frac{dz}{dt} \frac{dt}{d\ln\!\frst} \frac{d V_c}{dz}.
             \end{align}
@@
         \begin{equation}
-        \thardf \equiv - \diffp{t}{{\ln\!f_r}} = - \frac{f_r}{\partial f_r / \partial t}.
+            \label{eq:hardening}
+            \thardf \equiv - \diffp{t}{{\ln\!\frst}} = - \frac{\frst}{\partial \frst / \partial t}.
         \end{equation}
-    Plugging in the relevant cosmographic relationships (see \secref{sec:app_eqs}) we have,
-        \begin{align}
-            h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
-        \end{align}
-    The challenge in modeling astrophysical sources of GWs is in calculating their number density.  Numerous approaches are possible, described in the following sections.
+
+        % Plugging in the relevant cosmographic relationships (see \secref{sec:app_eqs}) we have,
+        %     \begin{align}
+            %         h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
+            %     \end{align}
+            % The challenge in modeling astrophysical sources of GWs is in calculating their number density.  Numerous approaches are possible, described in the following sections.
+
+        Plugging in \eqref{eq:hardening} along with the relevant cosmographic relationships (see \secref{sec:app_eqs}) into \eqref{eq:num_num_dens} we have,
+        \begin{equation}
+            \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} =
+                \frac{d^3 n_c}{dM \, dq \, dz} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
+        \end{equation}
 
@@
 % =================================================================================================
+
 \section{Results}
@@
 % =================================================================================================
+
 \section{Discussion \& Conclusions}
@@
 \bibliographystyle{mnras}
-\bibliography{refs}
+\bibliography{src/refs}
 

### 2023-01-20 holodeck/cyutils.pyx
Comment-out debugging print statements

@@ -1062,18 +1062,18 @@ cdef double[:, :] _sam_calc_gwb_7(
 
                     # if `gwfr` is lower than lowest evolution point, continue to next frequency
                     if (gwfr < frst_evo_lo):
-                        printf(
-                            "\nLO: %d - %d %d - %d - %e %e %e\n",
-                            ff, aa, bb, ecc_idx, gwfr*MY_YR, frst_evo_lo*MY_YR, frst_evo_hi*MY_YR
-                        )
+                        # printf(
+                        #     "\nLO: %d - %d %d - %d - %e %e %e\n",
+                        #     ff, aa, bb, ecc_idx, gwfr*MY_YR, frst_evo_lo*MY_YR, frst_evo_hi*MY_YR
+                        # )
                         continue
 
                     # if `gwfr` is > highest evolution point, also be true for all following frequencies, so break
                     if (gwfr > frst_evo_hi):
-                        printf(
-                            "\nHI: %d - %d %d - %d - %e %e %e\n",
-                            ff, aa, bb, ecc_idx, gwfr*MY_YR, frst_evo_lo*MY_YR, frst_evo_hi*MY_YR
-                        )
+                        # printf(
+                        #     "\nHI: %d - %d %d - %d - %e %e %e\n",
+                        #     ff, aa, bb, ecc_idx, gwfr*MY_YR, frst_evo_lo*MY_YR, frst_evo_hi*MY_YR
+                        # )
                         break
 
                     # calculate slope M

### 2023-01-28 holodeck/cyutils.pyx
Documentation additions, start merging new text into holodeck manuscript draft

@@ -399,23 +399,25 @@ cdef double[:, :] _sam_calc_gwb_single_eccen(
 
     """
 
+    # Initialize sizes and shapes
     cdef int nfreqs = len(gwfobs)
     cdef int n_mtot = len(mtot_log10)
     cdef int n_mrat = len(mrat)
     cdef int n_redz = len(redz)
     cdef int n_eccs = len(sepa_evo_in)
-
     cdef int num_freq_harm = nfreqs * nharms
     cdef (int *)shape = <int *>malloc(2 * sizeof(int))
     shape[0] = nfreqs
     shape[1] = nharms
 
+    # Declare variables used later
     cdef int ii, nh, jj, kk, aa, bb, ff, ecc_idx, ecc_idx_beg, ii_mm, kk_zz
     cdef double m1, m2, mchirp, sa, qq, tau
     cdef double gwfr, zterm, dc_cm, dc_mpc, dc_term, gne, hterm
     cdef double weight_ik, weight, four_over_nh_squared, sa_fourth
     cdef double fe_ecc, mt, mt_sqrt, nd, hterm_pref, frst_evo_lo, frst_evo_hi
 
+    # Initialize constants
     cdef double four_pi_c_mpc = 4 * M_PI * (MY_SPLC / MY_MPC)
     cdef double kep_sa_term = MY_NWTG / pow(2.0*M_PI, 2)
     cdef double one_third = 1.0 / 3.0
@@ -424,61 +426,81 @@ cdef double[:, :] _sam_calc_gwb_single_eccen(
     cdef double three_fifths = 3.0 / 5.0
     cdef double six_fifths = 6.0 / 5.0
 
-    cdef double *sepa_evo = <double *>malloc(n_eccs * sizeof(double))
-    cdef double *eccen_evo = <double *>malloc(n_eccs * sizeof(double))
-    cdef double _freq_pref = (1.0/(2.0*M_PI)) * sqrt(MY_NWTG)
-    cdef double *frst_evo_pref = <double *>malloc(n_eccs * sizeof(double))
-    for ii in range(n_eccs):
-        sepa_evo[ii] = <double>sepa_evo_in[ii]
-        eccen_evo[ii] = <double>eccen_evo_in[ii]
-        frst_evo_pref[ii] = _freq_pref / pow(sepa_evo[ii], 1.5)
-
-    cdef (double *)freq_harms = <double *>malloc(num_freq_harm * sizeof(double))
[…]

### 2023-01-28 docs/holodeck-paper/ms.tex
Documentation additions, start merging new text into holodeck manuscript draft

@@
     \subsection{Source Distributions and Gravitational Wave Calculations}
+
+        It was shown by \citet[][Eq.~5]{Phinney-2001} that the GWB spectrum, can be expressed in terms of the GW characteristic strain ($\hc$),
+        \begin{equation}
+            % \label{eq:}
+            \hc^2(f) = \frac{4G}{\pi c^2 f} \int dz \; \frac{d\ndens}{dz} \, \scale{d\egw}{d f_p} \bigg|_{f_p=f\lr{1+z}}.
+        \end{equation}
+        We can convert from the GW energy-spectrum ($\frac{d\egw}{d f_p}$) to GW spectral strain using Eq.~\ref{eq:gw_energy_spectrum}~\&~\ref{eq:strain_lum}, and rewrite this as,
+        \begin{equation}
+            \label{eq:gwb_sa}
+            \hc^2(f) = \int_0^\infty \!\! dz \, \sum_n^\infty \; \frac{d\ndens}{dz} \, \hsn^2\lr{f_p} \cdot 4\pi c \, d_c^2 \cdot \lr{1+z} \cdot \thard(f_p) \, \bigg|_{f_p = f(1+z)/n}.
+        \end{equation}
+        In the case of purely GW-driven binary hardening, and circular binary orbits, this can be simplified to \citep[][Eq.~11]{Phinney-2001},
+        \begin{equation}
+            \label{eq:gwb_ideal}
+            \hc^2(f) = \frac{4 \pi}{3 c^2} \, \lr[-4/3]{2\pi f} \int_0^\infty \!\! dz \; \frac{d\ndens}{dz}  \, \frac{\lr[5/3]{G \mchirp}}{\lr[1/3]{1+z}}.
+        \end{equation}
+        This ideal, power-law GWB is often expressed as, \mbox{$\hc(f) = A_{\yr} \cdot \lr[-2/3]{f \cdot \yr}$}.
+
+        The redshift evolution of the comoving number density of sources ($\ndens \equiv dN/dV_c$) can be related to the number of sources in a given frequency interval as \citep[][Eq.~6]{Sesana+2008},
+        \begin{equation}
+            \label{eq:number_density_to_number_frequency}
+            \frac{d\ndens}{dz} = \frac{d^2N}{dz \, dV_c} = \frac{d^2N}{dz \, d\ln f_p} \frac{d \ln f_p}{dt_r} \frac{dt_r}{dz} \frac{dz}{dV_c}.
+        \end{equation}
+        Plugging \eqref{eq:number_density_to_number_frequency} into \eqref{eq:gwb_sa}, and utilizing the expressions from Eqs.~\ref{eq:gw_hard},~\ref{eq:hard_time},~\ref{eq:comvol}~\&~\ref{eq:redshift_time}, to write:
+        \begin{equation}
+            \label{eq:gwb_mc}
+                \hc^2(f) = \int dz \sum_n^\infty \frac{d^2 N}{dz \, d\ln f_p} \, \hsn^2(f_p).
+        \end{equation}
+
+        While Eq.~\ref{eq:gwb_sa} and Eq.~\ref{eq:gwb_mc} are formally equal, how they are used in practice can produce important differences.
+In particular, as pointed out by \citep[][Eq.~6]{Sesana+2008}, `monte-carlo' calculations that explicitly use Eq.~\ref{eq:gwb_mc} encode the intrinsic discreteness of binary sources.
+The `semi-analytic' calculation using Eq.~\ref{eq:gwb_sa}, naively implemented, assumes a smooth, continuous distribution of binary sources without taking into account their quantization.
+The effect of binary quantization on the GWB is to produce a steeping of the GWB spectrum relative to the idealized power-law \eqref{eq:gwb_ideal}.
+The frequency at which this becomes important is model dependent, but typically at \mbox{$f \gtrsim 1 - 10 \, \pyr$}.
+
+    \subsection{Discretization and Cosmic Variance}
+
+    \subsection{??Other??}
 
@@
 All figures were generated using \texttt{matplotlib}~\citep{matplotlib2007}.
-Kernel density estimation was performed using the \texttt{kalepy}{} package (\href{https://github.com/lzkelley/kalepy}{github.com/lzkelley/kalepy}) \citep{kalepy2020}.
+Kernel density estimation was performed using the \texttt{kalepy}{} package (\href{https://github.com/lzkelley/kalepy}{github.com/lzkelley/kalepy}) \citep{kalepy2021}.
 
@@
     \label{sec:app_eqs}
+
+    \subsection{Binaries and Gravitational Waves}
+
         A binary is described by its total mass $M$, mass ratio $q \equiv m_2 / m_1$ (s.t.~$q\leq1$), (proper) semi-major axis $a$, and rest-frame orbital frequency $f_r$.
@@
 
-    The GW strain from a circular binary is \citep[][Eq.~7; sky and polarization averaged]{Sesana+2008},
+        The time evolution of a binary's orbit due to the loss of energy and angular momentum through gravitational waves was calculated by \citep[][Eq.~5.6--5.8]{Peters1964}:
+        \begin{subequations}
+        \label{eq:gw_hard}
+        \begin{align}
+            \frac{da}{dt} = & -\frac{64 \, G^3}{5 \, c^5} \frac{M_1 \, M_2 \, M}{a^3} F(e), \\
+            \frac{de}{dt} = & -\frac{304 \, G^3}{15 \, c^5} \frac{M_1 \, M_2 \, M}{a^4} \lr{e + \frac{121}{304}e^3},
+        \end{align}
+        \end{subequations}
+        where the GW eccentricity function is,
         \begin{equation}
-        \label{eq:source_strain}
-        h_s = \frac{8}{10^{1/2}} \frac{\left(G\mchirp\right)^{5/3}}{c^4 \, d_L}
-            \left(2 \pi f_r \right)^{2/3},
+            \label{eq:ecc_func}
+            F(e) \equiv \frac{1 + \frac{73}{24} e^2 + \frac{37}{96} e^4}{\left( 1 - e^2\right)^{7/2}}.
         \end{equation}
-    for a luminosity distance $d_L$.
-Recall that the luminosity distance and comoving distance are related as $d_L = d_c \, (1+z)$, that the comoving volume of the Universe is given by \citep{Hogg-1999},
+        We define the binary `hardening' timescale with respect to frequency, instead of separation, as:
         \begin{equation}
-        \frac{d V_c}{dz} = 4\pi \frac{c}{H_0} \frac{d_c^2}{E(z)},
+            \label{eq:hard_time}
+            \thard \equiv dt/d\ln \frst = \frst / (d\frst/dt) = - \frac{2}{3} a / (da/dt).
         \end{equation}
-    and that redshift can be related to time (or age of the universe) as,
+        In general, the hardening rate can be driven by other mechanisms in addition to gravitational wave emission.
+
+        GW power is emitted at integer harmonics of the orbital frequency, $\frstn = n \cdot \frstorb$, which is also redshifted to reach the observer-frame, $\fobsn = n \cdot \frstorb / (1+z)$.
+The energy spectrum emitted in GWs is \citet[][Eq.~3.10]{enoki2007a},
+        \begin{subequations}
         \begin{align}
-        \frac{dz}{dt} = H_0 \lr{1+z} E(z).
+            \label{eq:gw_energy_spectrum}
+            \frac{d \egw}{d \frst} & = \int \frac{dt_r}{df_p} \, \sum_{n=1}^\infty \, \lgwn(f_p) \; \delta(\frst - n f_p) \; d f_p \\
+                & = \sum_{n=1}^\infty \left[ \lgwn(f_p) \cdot \frac{\thard(f_p)}{n f_p} \right]_{f_p = \frst/n},
         \end{align}
-    The evolution of the Hubble parameter over redshift can be calculated from the matter-energy constituents of the universe as,
+        \end{subequations}
+        where the power-radiated at each harmonic can be expressed as is \citep[][Eq.~2.2]{enoki2007a},
         \begin{align}
-        H(z) & = H_0(z) E(z), \\
-        E(z) & \equiv \lrs[1/2]{\Omega_R \lr[4]{1+z} + \Omega_m \lr[3]{1+z} + \Omega_k \lr[2]{1+z} + \Omega_\Lambda}.
+            \label{eq:lum_gw}
+            \lgw & = \sum_n \lgwn = \lgwc \sum_n g(n,e) = \lgwc \cdot F(e), \\
+            \lgwc & = \frac{32}{5 G c^5} \left(G\mchirp\right)^{10/3} \left( 2\pi \frstorb \right)^{10/3}.
+        \end{align}
+        Here, the GW frequency distribution function is,
+        \begin{equation}
+        \begin{split}
+            \label{eq:freq_dist_func}
+            g(n,e) \equiv \frac{n^4}{32} \Biggl(& \left[ J_{n-2}(ne) - 2eJ_{n-1}(ne) + \frac{2}{n} J_n (ne) + 2e J_{n+1} (ne) - J_{n+2} (ne) \vphantom{\frac{2}{n}}\right]^2 \\
+            &+ \left(1-e^2\right) \Bigl[J_{n-2}(ne) - 2eJ_n(ne) + J_{n+2}(ne)\Bigr]^2 + \frac{4}{3n^2}\Bigl[J_n(ne)\Bigr]^2 \Biggr).
+        \end{split}
+        \end{equation}
+        The GW spectral strain amplitude at a given harmonic can be related to the (rest-frame) luminosity as \citep[][Eq.~2.1]{Finn+Thorne-2000},
+        \begin{align}
+            \label{eq:strain_lum}
+            \hsn^2 & = \frac{G}{c^3} \, \scale[2]{2}{n} \, \frac{\lgwn}{\lr[2]{2 \pi \, \frstorb} \, \distcom^2}, \\
+                & = \frac{32}{5 c^8} \, \scale[2]{2}{n} \, \frac{\left(G\mchirp\right)^{10/3}}{\distcom^2} \lr[4/3]{2\pi\frstorb} \cdot g(n,e).
+        \end{align}
+        We always take the chirp mass to be an intrinsic (rest-frame) property of the binary, however an `observer-frame' chirp-mass can also be used.
+For example, in the case of a circular binary, the spectral strain is,
+        \begin{equation}
+            \label{eq:gw_strain_amp_circ}
+            \hscirc
+                = \frac{8}{10^{1/2}} \frac{\left(G\mchirp\right)^{5/3}}{c^4 \, \distcom} \left(2 \pi \frstorb \right)^{2/3}
+                = \frac{8}{10^{1/2}} \frac{\left(G\mchirp_o\right)^{5/3}}{c^4 \, \distlum} \left(2 \pi \fobsorb \right)^{2/3}.
+        \end{equation}
+
+    \subsection{Cosmology}
+        Some useful cosmographic relations are reproduced here \citep{Hogg-1999}.
+Recall that luminosity distance and comoving distance are related as $\distlum = (1+z) \distcom$.
+The differential comoving volume of the Universe is,
+        \begin{equation}
+            \label{eq:comvol}
+            \frac{d V_c}{dz} = 4\pi \, d_c^2 \, \frac{c}{H(z)},
+        \end{equation}
+        and time (or age of the universe) can be related to redshift through,
+        \begin{align}
+            \label{eq:redshift_time}
+            \frac{dz}{dt} = \lr{1+z} \; H(z).
+        \end{align}
+        The Hubble parameter can be calculated from the matter-energy constituents of the universe as,
+        \begin{align}
+            H(z) & = H_0(z) \cdot \lrs[1/2]{\Omega_R \lr[4]{1+z} + \Omega_m \lr[3]{1+z} + \Omega_k \lr[2]{1+z} + \Omega_\Lambda}.
         \end{align}

### 2023-01-31 docs/holodeck-paper/ms.tex
Transfer over a bunch of material for holodeck paper

@@
 
-    For binary MBHs in particular, a variety of physical processes are required for the two MBHs to find the center of the post-merger host galaxy and eventually become gravitationally bound \needcite{BBR1990}.
+    For binary MBHs in particular, a variety of physical processes are required for the two MBHs to find the center of the post-merger host galaxy and eventually become gravitationally bound \needcite{BBR1980}.
 In particular, the two MBHs must interact with their host galaxy environments through dynamical friction, stellar scattering, accretion disk torques, and possibly three-body interactions with a third MBH component \needcite{}, before reaching small enough binary separations for GW emission to eventually dominate the final coalescence of the binary.
@@
 
-    \subsection{??Other??}
+        The conversion from a continuous number-density of sources, to a discrete population of binaries is given by \eqref{eq:number_density_to_number_frequency}.
+This can also clearly be seen by comparing Eq.~\ref{eq:gwb_mc} with Eq.~\ref{eq:gwb_sa}, to identify:
+        \begin{equation}
+            \label{eq:pop_discrete_from_continuous}
+            \frac{d^2 N}{dz \, d\ln\!f_p} =
+                \frac{d \ndens}{dz} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf(f_p).
+        \end{equation}
 
-        Let $N(M,q,z,\frst)$ denote the number of binaries up to a total mass $M$, mass ratio $q$, redshift $z$, and rest-frame orbital-frequency $\frst$.
-        We can calculate the characteristic strain from an ensemble of GW sources as \citep[][Eqs.~5/8]{Phinney-2001},
-            \begin{align}
-                h_c^2 = \int_0^\infty \!\! dz \; \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} \; \hs^2\lr{\frst}.
-            \end{align}
-        In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$, for some comoving volume $V_c$, extrapolated into an observer's entire past light-cone.
+        In general, implementations of binary populations will be described over some finite grid, often in a parameter space of total mass, mass ratio, redshift, and orbital frequency.
+In this case the derivative over redshift in \eqref{eq:pop_discrete_from_continuous}, should be replaced with one over all parameters of interest, e.g.~\mbox{$d\ndens / dz \rightarrow d^3\ndens / dM \, dq \, dz$}.
+All additional parameters are treated the same in this context, so for brevity we will only explicitly include redshift.
+
+        Eventually the end goal of binary modeling is often to determine the number of binaries over some range of parameters.
+When calculating gravitational wave signals, for example, the number of binaries across a target frequency interval is desired.
+The discreteness of binary sources can be enforced by converting the differential value of \eqref{eq:pop_discrete_from_continuous} into a number of sources over a small bin, i.e., \mbox{$\Delta N \approx \lr{d^2 N / dz \, d\ln\!f_p} \, \cdot \Delta z \cdot \Delta \ln\!f_p$}.
+
+        In general, the properties of \textit{every} binary in the Universe are not known, and thus the MC calculation cannot be performed directly.
+The typical solution is to use some version of \eqref{eq:number_density_to_number_frequency} to convert from a number-density (e.g.~$d\ndens/dz$) into a discrete distribution of individual sources (e.g.~$d^2 N / dz \, d\ln f_r$).
+The number-density is often constructed from either (semi-)analytic models which specify $\ndens(M,q,z,\dots)$ explicitly, or from simulations/observations of sources in a finite volume.
+In the latter case, the number-density will be calculated as $\ndens \equiv N_\trt{finite}/V_\trt{c,finite}$, often over bins in some parameter space (e.g.~$\{M, q, z, \dots\}$).
+\\
+
+        For a finite population of binaries, the GWB spectrum is only defined over finite bins of frequency.
+The definition of characteristic strain is related to the GW energy across a logarithmic frequency bin, so \eqref{eq:gwb_mc} should really be written as,
+        \begin{equation}
+            \label{eq:gwb_discrete}
+            \hc^2(\Delta \ln f) = \sum_\textrm{bins} \sum_n^\infty \frac{\Delta N(f_p,M_a,q_b,z_c,\dots)}{\Delta \ln f_p} \, \hsn^2(f_p,M_a,q_b,z_c,\dots).
+        \end{equation}
+        Here we have written out explicitly that the number of binaries, $\Delta N(f_p,M_a,q_b,z_c,\dots)$, is the number sources with frequencies inside $\Delta \ln f_p$, typically over bins of mass, mass-ratio, redshift, etc.
+Discretizing and plugging into Eq.~\ref{eq:number_density_to_number_frequency} (or by examining Eq.~\ref{eq:gwb_sa}), we can then write,
+        \begin{equation}
+            \label{eq:expectation_number_of_binaries}
+            \langle \Delta N \rangle = \Delta \ndens \cdot 4\pi c \, \distcom^2 \cdot (1+z) \cdot \thard \cdot \Delta \ln f_p.
+        \end{equation}
+        Note that we have again dropped the written dependence on bin values $(M_a, q_b, \dots)$, but they are still implicitly present.
+This expression represents an \textit{expectation value} based on extrapolating the given number-density of binaries.
+It also will not give an integer, discrete number of binaries.
+To address this, we construct discrete realizations by drawing from a Poisson distribution: $\Delta N \in \poisson(\langle \Delta N \rangle)$.
+\\
+
+        The way \eqref{eq:expectation_number_of_binaries} is handled in practice will still vary somewhat depending on how $\ndens$ is specified.  If a semi-analytic model is being used, then often
+        $\Delta \ndens = \left[ d^3 \ndens(M,q,z) / dM \, dq \, dz \right] \cdot \Delta M \, \Delta q \, \Delta z$.
+In the case of a finite volume population, the procedure can be simplified.
+An analogous exercise would amount to a binning and un-binning of the sources, i.e.~$\Delta \ndens \big|_\textrm{finite} = \left[ \Delta N_\textrm{finite}(M,q,z) / \Delta M \, \Delta q \, \Delta z \, \Delta V_c \right] \cdot \Delta M \, \Delta q \, \Delta z$, which unnecessarily reduces accuracy.
+Instead, we can imagine shrinking each of our parameter-space bins such that each one captures a single source, in which case we can write,
+        \begin{equation}
+            \label{eq:expectation_number_of_single_binary}
+            \langle \Delta N \rangle_\textrm{source} = \frac{4\pi c \, \distcom^2 \, \thard}{V_c} \cdot (1+z) \cdot \Delta \ln f_p.
+        \end{equation}
+        In this case, $\thard$ and $\hsn$ in \eqref{eq:gwb_discrete}, are the hardening time and strain amplitude for each source, and $\langle \Delta N \rangle_\textrm{source}$ represents the expectation number of binaries in the Universe implied by each source in the finite volume.
+
+    % \subsection{??Other??}
+
+    %     Let $N(M,q,z,\frst)$ denote the number of binaries up to a total mass $M$, mass ratio $q$, redshift $z$, and rest-frame orbital-frequency $\frst$.
+    %     We can calculate the characteristic strain from an ensemble of GW sources as \citep[][Eqs.~5/8]{Phinney-2001},
+    %         \begin{align}
+    %             h_c^2 = \int_0^\infty \!\! dz \; \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} \; \hs^2\lr{\frst}.
+    %         \end{align}
+    %     In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$, for some comoving volume $V_c$, extrapolated into an observer's entire past light-cone.
 These can be related as \citep[][Eq.~6]{Sesana+2008},
-            \begin{align}
-                \label{eq:num_num_dens}
-                \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} = \frac{d^3 n_c}{dM \, dq \, dz} \frac{dz}{dt} \frac{dt}{d\ln\!\frst} \frac{d V_c}{dz}.
-            \end{align}
-        Note that the total number of binaries in a given frequency interval depends not only on the number density, but also their rate of frequency-evolution.
+    %         \begin{align}
+    %             \label{eq:num_num_dens}
+    %             \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} = \frac{d^3 n_c}{dM \, dq \, dz} \frac{dz}{dt} \frac{dt}{d\ln\!\frst} \frac{d V_c}{dz}.
+    %         \end{align}
+    %     Note that the total number of binaries in a given frequency interval depends not only on the number density, but also their rate of frequency-evolution.
 The faster binaries evolve through a given frequency interval, the fewer binaries will occupy that interval on average.
 This is typically described by the binary `hardening' timescale (or `residence time'), given by the expression,
-        \begin{equation}
-            \label{eq:hardening}
-            \thardf \equiv - \diffp{t}{{\ln\!\frst}} = - \frac{\frst}{\partial \frst / \partial t}.
-        \end{equation}
+    %     \begin{equation}
+    %         \label{eq:hardening}
+    %         \thardf \equiv - \diffp{t}{{\ln\!\frst}} = - \frac{\frst}{\partial \frst / \partial t}.
+    %     \end{equation}
 
-        % Plugging in the relevant cosmographic relationships (see \secref{sec:app_eqs}) we have,
-        %     \begin{align}
-            %         h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
-            %     \end{align}
-            % The challenge in modeling astrophysical sources of GWs is in calculating their number density.  Numerous approaches are possible, described in the following sections.
+    %     % Plugging in the relevant cosmographic relationships (see \secref{sec:app_eqs}) we have,
+    %     %     \begin{align}
+    %         %         h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
+    %         %     \end{align}
+    %         % The challenge in modeling astrophysical sources of GWs is in calculating their number density.  Numerous approaches are possible, described in the following sections.
 
-        Plugging in \eqref{eq:hardening} along with the relevant cosmographic relationships (see \secref{sec:app_eqs}) into \eqref{eq:num_num_dens} we have,
-        \begin{equation}
-            \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} =
-                \frac{d^3 n_c}{dM \, dq \, dz} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
-        \end{equation}
+    %     Plugging in \eqref{eq:hardening} along with the relevant cosmographic relationships (see \secref{sec:app_eqs}) into \eqref{eq:num_num_dens} we have,
+    %     \begin{equation}
+    %         \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} =
+    %             \frac{d^3 n_c}{dM \, dq \, dz} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
+    %     \end{equation}
 
@@
 
-    % From the SAM formalism, the number of binaries in each bin is calculated by `integrating' each finite element of \eqref{eq:sam_final}, i.e.,
-    % \begin{equation}
-    %     \diffp{N_{ijk}(f_r)}{{\ln\!f_r}} \equiv \diffp{N(M_i, q_j, z_k, {f_r})}{{M}{q}{z}{\ln\!f_r}} \, \Delta M \,\Delta q \, \Delta z.
-    % \end{equation}
-    % The characteristic strain of the GWB can then be calculated
+        From the SAM formalism, the number of binaries in each bin is calculated by `integrating' each finite element of \eqref{eq:sam_final}, i.e.,
+        \begin{equation}
+            \diffp{N_{ijk}(f_r)}{{\ln\!f_r}} \equiv \diffp{N(M_i, q_j, z_k, {f_r})}{{M}{q}{z}{\ln\!f_r}} \, \Delta M \,\Delta q \, \Delta z.
+        \end{equation}
+        The characteristic strain of the GWB can then be calculated
 
-    % % Fig -
-    % \begin{figure}
-    %     \centering
-    %     \includegraphics[width=1\columnwidth]{{{example-image-c}}}
-    %     \caption{\textbf{header.} body.}
-    %     \label{fig:fig1}
-    % \end{figure}
-    %
-    %
-    % \begin{equation}\begin{split}
-    %     \label{eq:eq1}
-    %     2+2 & = 4 \\
-    %         & = 10 - 6
-    % \end{split}\end{equation}
-    %
-    % % Fig -
-    % \begin{figure*}
-	% 	\centering
-    %     \subfloat[first]{{
-    %         \includegraphics[width=1.75\columnwidth]{{{example-image-a}}}
-    %     }}
-	% 	\\
-    %     % \qquad
-    %     \subfloat[second]{{
-    %     	\includegraphics[width=1.75\columnwidth]{{{example-image-b}}}
-    %     }}
-    %     \caption{\textbf{Head.}  Body.}
-    %     \label{fig:fig2}
-    % \end{figure*}
+    \subsection{Finite Volume Populations: the Illustris simulations}
+
+        \todo{Describe illustris models.}
 

### 2023-02-02 scripts/gen_lib_sams.py
New libraries

@@ -23,7 +23,7 @@ To-Do
     * Does this resolve irregularities between different LHS implementations?
 * Use subclassing to cleanup `Parameter_Space` object.  e.g. implement LHS as subclass of generic Parameter_Space class.
 * BUG: `lhs_grid` and `lhs_grid_idx` are currently storing the same thing
-
+* #! IMPORTANT: mark output directories as incomplete until all runs have been finished.  Merged libraries from incomplete directories should also get some sort of flag! !#
 
 """
 
@@ -490,6 +490,13 @@ class LHS_PSpace_Eccen_02(holo.librarian._LHS_Parameter_Space):
         'eccen_init',
         'gsmf_phi0',
         'gsmf_phiz',
+        'gpf_malpha',
+        'gpf_zbeta',
+        'gpf_qgamma',
+
+        'gmt_malpha',
+        'gmt_zbeta',
+        'gmt_qgamma',
         'mmb_amp',
         'mmb_plaw',
     ]
@@ -575,7 +582,7 @@ def setup_argparse():
     return args
 
 
-SPACE = LHS_PSpace_Eccen_01
+SPACE = LHS_PSpace_Eccen_02
 comm = MPI.COMM_WORLD
 
 args = setup_argparse() if comm.rank == 0 else None

### 2023-02-02 scripts/run_gen_lib_sams.sh

@@ -4,7 +4,7 @@
 # Setup for running on the `minos` computer
 # --------------------------------------------------------------------------------------------------
 
-OUTPUT="output/debug01b_2023-01-24_01_n1000_g1000_s40_r100_f10"
+OUTPUT="output/eccen-02_2023-01-31_02_n100_s30_r100_f20"
 #OUTPUT=$OUTPUT"__TEST"
 SCRIPT="scripts/gen_lib_sams.py"
 LOG_NAME="job_log"
@@ -15,6 +15,6 @@ cp $0 "$OUTPUT/"
 LOG_OUT="$LOG_NAME.out"
 LOG_ERR="$LOG_NAME.err"
 
-mpirun -n 19  python $SCRIPT $OUTPUT -n 1000 -s 40 -r 100 -f 10  1> $LOG_OUT 2> $LOG_ERR &
+mpirun -n 19  python $SCRIPT $OUTPUT -n 100 -s 30 -r 100 -f 20  1> $LOG_OUT 2> $LOG_ERR &
 # mpirun -n 2  python $SCRIPT $OUTPUT -n 4 -s 40 -r 100 -f 10  1> $LOG_OUT 2> $LOG_ERR &
 

### 2023-02-02 scripts/gen_lib_sams.py
Update eccen_01 pspace

@@ -447,6 +447,7 @@ class LHS_PSpace_Eccen_01(holo.librarian._LHS_Parameter_Space):
     _PARAM_NAMES = [
         'eccen_init',
         'gsmf_phi0',
+        'gpf_zbeta',
         'mmb_amp',
     ]
 
@@ -455,22 +456,23 @@ class LHS_PSpace_Eccen_01(holo.librarian._LHS_Parameter_Space):
     def __init__(self, log, nsamples, sam_shape, lhs_sampler, seed):
         super().__init__(
             log, nsamples, sam_shape, lhs_sampler, seed,
-            eccen_init=[0.0, +1.0],
+            eccen_init=[0.0, +0.975],
             gsmf_phi0=[-3.0, -2.0],
+            gpf_zbeta=[+0.0, +2.0],
             mmb_amp=[0.1e9, 1.0e9],
         )
 
     def sam_for_lhsnumber(self, lhsnum):
         param_grid = self.params_for_lhsnumber(lhsnum)
 
-        eccen, gsmf_phi0, mmb_amp = param_grid
+        eccen, gsmf_phi0, gpf_zbeta, mmb_amp = param_grid
         mmb_amp = mmb_amp*MSOL
 
         # favor higher values of eccentricity instead of uniformly distributed
         eccen = eccen ** (1.0/5.0)
 
         gsmf = holo.sam.GSMF_Schechter(phi0=gsmf_phi0)
-        gpf = holo.sam.GPF_Power_Law()
+        gpf = holo.sam.GPF_Power_Law(zbeta=gpf_zbeta)
         gmt = holo.sam.GMT_Power_Law()
         mmbulge = holo.relations.MMBulge_KH2013(mamp=mmb_amp)
 

### 2023-02-02 scripts/run_gen_lib_sams.sh
Update run script

@@ -4,7 +4,7 @@
 # Setup for running on the `minos` computer
 # --------------------------------------------------------------------------------------------------
 
-OUTPUT="output/debug01b_2023-01-24_01_n1000_g1000_s40_r100_f10"
+OUTPUT="output/eccen-01_2023-01-31_01_n1000_s40_r100_f40"
 #OUTPUT=$OUTPUT"__TEST"
 SCRIPT="scripts/gen_lib_sams.py"
 LOG_NAME="job_log"
@@ -15,6 +15,6 @@ cp $0 "$OUTPUT/"
 LOG_OUT="$LOG_NAME.out"
 LOG_ERR="$LOG_NAME.err"
 
-mpirun -n 19  python $SCRIPT $OUTPUT -n 1000 -s 40 -r 100 -f 10  1> $LOG_OUT 2> $LOG_ERR &
+mpirun -n 8  python $SCRIPT $OUTPUT  -n 100 -s 40 -r 100 -f 40  1> $LOG_OUT 2> $LOG_ERR &
 # mpirun -n 2  python $SCRIPT $OUTPUT -n 4 -s 40 -r 100 -f 10  1> $LOG_OUT 2> $LOG_ERR &
 

### 2023-02-02 holodeck/evolution.py
test pre-merge :: from origin/dev for dev_sam-merge

@@ -46,37 +46,6 @@ To-Do
     *   re-implement "magic" hardening models that coalesce in zero change-of-redshift or fixed
         amounts of time.
 
-*   Dynamical_Friction_NFW
-
-    *   Allow stellar-density profiles to also be specified (instead of using a hard-coded
-        Dehnen profile)
-    *   Generalize calculation of stellar characteristic radius.  Make self-consistent with
-        stellar-profile, and user-specifiable.
-
-*   Evolution
-
-    *   `_sample_universe()` : sample in comoving-volume instead of redshift
-
-*   Sesana_Scattering
-
-    *   Allow stellar-density profile (or otherwise the binding-radius) to be user-specified
-        and flexible.  Currently hard-coded to Dehnen profile estimate.
-
-*   _SHM06
-
-    *   Interpolants of hardening parameters return 1D arrays.
-
-*   Fixed_Time
-
-    *   Handle `rchar` better with respect to interpolation.  Currently not an interpolation
-        variable, which restricts it's usage.
-    *   This class should be separated into a generic `_Fixed_Time` class that can use any
-        functional form, and then a 2-power-law functional form that requires a specified
-        normalization.  When they're combined, it will produce the same effect.  Another good
-        functional form to implement would be GW + log-uniform hardening time, the same as the
-        current phenomenological model but with both power-laws set to 0.
-
-
 References
 ----------
 * [BBR1980]_ Begelman, Blandford & Rees 1980.
@@ -95,37 +64,17 @@ from __future__ import annotations
 # ! -- then call it again with a revision to the estimate, using right-edge values -- !
 # ! ===============================================================================================!
 
-import abc
-# import inspect
-import json
-import os
-# from typing import Union, TypeVar  # , Callable, Iterator
-import warnings
-
 import numpy as np
-import scipy as sp
-import scipy.interpolate   # noqa
 
 import kalepy as kale
 
 import holodeck as holo
-from holodeck import utils, cosmo, log, _PATH_DATA
-from holodeck.constants import GYR, NWTG, PC, MSOL, YR
-from holodeck import accretion
+from holodeck import utils, cosmo, log
+from holodeck.constants import PC
+from holodeck.hardening import _Hardening
+# from holodeck import accretion
 
[…]

### 2023-02-02 holodeck/hardening.py
test pre-merge :: from origin/dev for dev_sam-merge

@@ -1,0 +1,1415 @@
+"""
+To-Do
+-----
+
+*   Dynamical_Friction_NFW
+
+    *   Allow stellar-density profiles to also be specified (instead of using a hard-coded
+        Dehnen profile)
+    *   Generalize calculation of stellar characteristic radius.  Make self-consistent with
+        stellar-profile, and user-specifiable.
+
+*   Evolution
+
+    *   `_sample_universe()` : sample in comoving-volume instead of redshift
+
+*   Sesana_Scattering
+
+    *   Allow stellar-density profile (or otherwise the binding-radius) to be user-specified
+        and flexible.  Currently hard-coded to Dehnen profile estimate.
+
+*   _SHM06
+
+    *   Interpolants of hardening parameters return 1D arrays.
+
+*   Fixed_Time
+
+    *   Handle `rchar` better with respect to interpolation.  Currently not an interpolation
+        variable, which restricts it's usage.
+    *   This class should be separated into a generic `_Fixed_Time` class that can use any
+        functional form, and then a 2-power-law functional form that requires a specified
+        normalization.  When they're combined, it will produce the same effect.  Another good
+        functional form to implement would be GW + log-uniform hardening time, the same as the
+        current phenomenological model but with both power-laws set to 0.
+
+
+References
+----------
+* [BBR1980]_ Begelman, Blandford & Rees 1980.
+* [Chen2017]_ Chen, Sesana, & Del Pozzo 2017.
+* [Kelley2017a]_ Kelley, Blecha & Hernquist 2017.
+* [Quinlan1996]_ Quinlan 1996.
+* [Sesana2006]_ Sesana, Haardt & Madau et al. 2006.
+* [Sesana2010]_ Sesana 2010.
+
+"""
+from __future__ import annotations
+
+# ! ===============================================================================================!
+# ! -- UPDATE `_take_next_step` -- !
+# ! -- write a function to set right-edge values, call it with estimate from left edge -- !
+# ! -- then call it again with a revision to the estimate, using right-edge values -- !
+# ! ===============================================================================================!
+
+import abc
+# import inspect
+import json
+import os
+# from typing import Union, TypeVar  # , Callable, Iterator
+import warnings
+
+import numpy as np
+import scipy as sp
+import scipy.interpolate   # noqa
+
+import holodeck as holo
[…]

### 2023-03-11 holodeck/param_spaces.py
Narrow the parameter space for Astro-Tight-03 slighlty in hard_time

@@ -330,7 +330,7 @@ class PS_Astro_Tight_03(PS_Broad_Uniform_01):
         super(PS_Broad_Uniform_01, self).__init__(
             log, nsamples, sam_shape, seed,
 
-            hard_time=PD_Uniform(1.0, 10.0),   # [Gyr]
+            hard_time=PD_Uniform(1.0, 8.0),   # [Gyr]
             hard_gamma_inner=PD_Uniform(-1.5, -0.5),
 
             # from `sam-parameters.ipynb` fits to [Tomczak+2014] with 4x stdev values

### 2023-03-11 holodeck/gps/gp_utils.py
Flush print output to get immediate output on cluster jobs

@@ -172,7 +172,7 @@ def train_gp(spectra_file,
     spectra = h5py.File(spectra_file, "r")
 
     if VERBOSE:
-        print(f"Loaded spectra from {spectra_file}")
+        print(f"Loaded spectra from {spectra_file}", flush=True)
 
     # Get smoothed GWB
     gp_freqs, xobs, yerr, yobs, yobs_mean = get_smoothed_gwb(spectra, nfreqs,
@@ -181,13 +181,18 @@ def train_gp(spectra_file,
 
     pars = list(spectra.attrs["param_names"].astype(str))
 
-    # xobs = get_parameter_values(spectra, test_frac)
+
+    if VERBOSE:
+        print(f"Creating kernels", flush=True)
 
     gp_george, num_kpars = create_gp_kernels(gp_freqs, pars, xobs, yerr, yobs,
                                              kernel, kernel_opts)
 
     # Sample the posterior distribution of the kernel parameters
     # to find MAP value for each frequency.
+
+    if VERBOSE:
+        print(f"Fitting GPs", flush=True)
 
     fit_kernel_params(gp_freqs, yobs_mean, gp_george, num_kpars, nwalkers,
                       nsamples, burn_frac, mpi)
@@ -236,7 +241,7 @@ def get_smoothed_gwb(spectra, nfreqs, test_frac=0.0, center_measure="median"):
     xobs = spectra['sample_params']
     bads = np.any(np.isnan(gwb_spectra), axis=(1, 2))
     if VERBOSE:
-        print(f"Found {utils.frac_str(bads)} samples with NaN entries.  Removing them from library.")
+        print(f"Found {utils.frac_str(bads)} samples with NaN entries.  Removing them from library.", flush=True)
     # when sample points fail, all parameters are set to zero.  Make sure this is consistent
     if not np.all(xobs[bads] == 0.0):
         raise RuntimeError(f"NaN elements of `gwb` did not correspond to zero elements of `sample_params`!")
@@ -250,7 +255,7 @@ def get_smoothed_gwb(spectra, nfreqs, test_frac=0.0, center_measure="median"):
     # Cut out portion for test set later
     test_ind = int(gwb_spectra.shape[0] * test_frac)
     if VERBOSE:
-        print(f"setting aside {test_frac} of samples ({test_ind}) for testing, and choosing {nfreqs} frequencies")
[…]

### 2023-03-11 scripts/gp_trainer.py
Cleanup scripts, mark templates that users should make copies of for their own uses

@@ -6,6 +6,7 @@ import configparser
 import os
 import pickle
 import shutil
+import sys
 from datetime import datetime
 from pathlib import Path
 
@@ -100,6 +101,9 @@ def main(config):
 
 
 if __name__ == "__main__":
+    print()
+    print(__file__)
+
     args = setup_argparse()
     config = configparser.ConfigParser()
     config.read(args.config_path)

### 2023-03-11 scripts/run_gen_lib_sams.sh
Cleanup scripts, mark templates that users should make copies of for their own uses

@@ -1,47 +1,0 @@
-#!/bin/bash
-# --------------------------------------------------------------------------------------------------
-# Run parallel Semi-Analytic Model (SAM) library generation script in parallel.
-# Setup for running on the `minos` computer
-# --------------------------------------------------------------------------------------------------
-
-echo "================================================================================"
-echo "==================    HOLODECK - run_gen_lib_sams.sh    ========================"
-echo "================================================================================"
-echo ""
-
-# ====    setup simulation parameters    ====
-NAME="ps-test-pd-01_2023-03-01"
-PARS="n100_s61-81-101_r100_f40"
-
-# NAME="TEST__"$NAME
-
-SCRIPT="./scripts/gen_lib_sams.py"
-LOG_NAME="job_log_"$NAME
-
-OUTPUT="./output/"$NAME"_"$PARS
-
-mkdir -p $OUTPUT
-echo "Output directory: ${OUTPUT}"
-
-cp $0 "$OUTPUT/runtime_job-script.slurm"
-LOG_OUT="$LOG_NAME.out"
-LOG_ERR="$LOG_NAME.err"
-echo "logs: ${LOG_OUT} ${LOG_ERR}"
-
-# ====    run simulations    ====
-
-echo "PWD:"
-pwd
-ls $SCRIPT
-echo "Running `mpirun`"
-
-set -x
-
-# mpirun -np 80  python $SCRIPT $OUTPUT -n 4000 -r 100 -f 40  1> $LOG_OUT 2> $LOG_ERR
-# mpirun -n 4  python $SCRIPT $OUTPUT -n 8 -s 10 -r 20 -f 40  1> $LOG_OUT 2> $LOG_ERR
-python $SCRIPT $OUTPUT -n 8 -s 10 -r 20 -f 40  1> $LOG_OUT 2> $LOG_ERR
-
-SHARE_OUTPUT=$OUTPUT"_SHARE"
-mkdir -p $SHARE_OUTPUT
-cp $OUTPUT/{*.py,holodeck*.log,*.hdf5} $SHARE_OUTPUT/
-cp {$LOG_ERR,$LOG_OUT} $SHARE_OUTPUT/

### 2023-03-11 holodeck/gps/gp_utils.py
debugging output

@@ -181,7 +181,6 @@ def train_gp(spectra_file,
 
     pars = list(spectra.attrs["param_names"].astype(str))
 
-
     if VERBOSE:
         print(f"Creating kernels", flush=True)
 
@@ -241,21 +240,23 @@ def get_smoothed_gwb(spectra, nfreqs, test_frac=0.0, center_measure="median"):
     xobs = spectra['sample_params']
     bads = np.any(np.isnan(gwb_spectra), axis=(1, 2))
     if VERBOSE:
-        print(f"Found {utils.frac_str(bads)} samples with NaN entries.  Removing them from library.", flush=True)
+        print(f"Found {utils.frac_str(bads)} samps w/ NaNs. Removing from lib", flush=True)
     # when sample points fail, all parameters are set to zero.  Make sure this is consistent
     if not np.all(xobs[bads] == 0.0):
-        raise RuntimeError(f"NaN elements of `gwb` did not correspond to zero elements of `sample_params`!")
+        err = f"NaN elements of `gwb` did not correspond to zero elements of `sample_params`!"
+        raise RuntimeError(err)
     # Select valid spectra, and sample parameters
     gwb_spectra = gwb_spectra[~bads]
     xobs = xobs[~bads]
     # Make sure old/deprecated parameters are not in library
     if 'mmb_amp' in spectra.attrs['param_names']:
-        raise RuntimeError("Parameter `mmb_amp` should not be here!  Needs to be log-spaced (`mmb_amp_log10`)!")
+        err = "Parameter `mmb_amp` should not be here!  Needs to be log-spaced (`mmb_amp_log10`)!"
+        raise RuntimeError(err)
 
     # Cut out portion for test set later
     test_ind = int(gwb_spectra.shape[0] * test_frac)
     if VERBOSE:
-        print(f"setting aside {test_frac} of samples ({test_ind}) for testing, and choosing {nfreqs} frequencies", flush=True)
+        print(f"{test_frac}/{test_ind} for testing, and choosing {nfreqs} freqs", flush=True)
 
     gwb_spectra = gwb_spectra[test_ind:, :nfreqs, :]**2
     xobs = xobs[test_ind:, :]
@@ -443,7 +444,7 @@ def fit_kernel_params(gp_freqs, yobs_mean, gp_george, nkpars, nwalkers,
         mpi=mpi, processes=min(nwalkers // 2, cpu_count()))
     if VERBOSE:
[…]

### 2023-03-12 holodeck/cyutils.pyx
Fix outdated DOCS in cyutils

@@ -1,13 +1,14 @@
 """Module for methods implemented in cython.
 
-This module can be used by including the following code:
-    ```
-    import pyximport
-    pyximport.install(language_level=3, setup_args={"include_dirs": np.get_include()})
-    import holodeck.cyutils
-    ```
+To use and load this module, you will need to build the cython extension using:
 
-python setup.py build_ext -i
+$ python setup.py build_ext -i
+
+from the holodeck root directory (containing the `setup.py` file).
+
+And you still need to install holodeck in develop mode, using:
+
+$ python setup.py develop
 
 """
 

### 2023-05-21 docs/holodeck-paper/ms.tex
Add holodeck title
Updates to introduction

@@
 % =================================================================================================
+\documentclass[useAMS, usenatbib]{src/mnras}
 
-% general package imports and setup
+% % general package imports and setup
 \input{src/packages}
-% definitions of custom commands
+% % definitions of custom commands
 \input{src/commands}
-% journal abbreviations for bibliography
+% % journal abbreviations for bibliography
 \input{src/abbrevs}
@@
 
-\newtotcounter{citnum} % From the package documentation
-\def\oldbibitem{} \let\oldbibitem=\bibitem
-\def\bibitem{\stepcounter{citnum}\oldbibitem}
+% \newtotcounter{citnum} % From the package documentation
+% \def\oldbibitem{} \let\oldbibitem=\bibitem
+% \def\bibitem{\stepcounter{citnum}\oldbibitem}
 
-\newcommand{\orcidauthorA}{0000-0002-6625-6450} % For author A
+% \newcommand{\orcidauthorA}{0000-0002-6625-6450} % For author A
 
-\title[Short Title]{Very long and verbose title}
+\title[holodeck: MBHB populations]{Enter the \texttt{holodeck}: massive black hole binary population synthesis}
 \author[NANOGrav]{the NANOGrav collaboration}
-% \author[L.Z.~Kelley et al.]{
-%     Luke Zoltan Kelley$^{1}$\thanks{E-mail:lzkelley@northwestern.edu}\orcidA{} \\
-%     $^{1}$ \begin{minipage}[t]{\linewidth} Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA), and\\Department of Physics \& Astronomy, Northwestern University, Evanston, IL 60208
-%     \end{minipage}
-% }
 
@@
 
-    Massive black holes (MBHs) are known to occupy the centers of most if not all massive galaxies \needcite{}.
-    Binaries of MBHs are believed to form following the merger of massive galaxies \needcite{}.
-Such MBH Binaries (MBHBs) could then produce low-frequency gravitational waves (GWs) \needcite{LF-GWs}, in the nano-Hertz regime, that would soon be detectable by experiments called pulsar timing arrays (PTAs) \needcite{PTAs}.
-In this paper, we introduce the MBHB population synthesis code called \holodeck{}.
+    % Massive black holes (MBHs) are known to occupy the centers of most if not all massive galaxies \needcite{}.
+    Binaries of MBHs form following the merger of massive galaxies \needcite{}.
+These MBH Binaries (MBHBs) can then produce the loudest gravitational waves (GWs) in the Universe.
+MBHBs with total masses $M \gtrsim 10^8 \,\msol$ are detectable by experiments called pulsar timing arrays (PTAs) \needcite{PTAs} in the nanohertz regime \needcite{}, while more modest systems with $M \lesssim 10^7 \, \msol$ will be detectable by LISA \needcite{LISA} at millihertz \needcite{}.
+While accreting, MBHs are visible across the electromagnetic (EM) spectrum as active galactic nuclei (AGN) \needcite{}, suggesting that MBHBs could be promising multimessenger sources \needcite{}.
+A large number of studies have modeled cosmological populations of MBHBs, and made predictions for their GW signatures and EM counterparts, however no comprehensive framework has been released.
+In this paper, we introduce the MBHB population synthesis code \holodeck{}, developed within the astrophysics working group of the NANOGrav\footnote{North American nanohertz observatory for gravitational waves \citep{}: \url{https://www.nanograv.org}.}.
 The goal of this package is to provide a cohesive and comprehensive framework for:
@@
         \item
-        modeling the physics of MBHB formation and evolution while accurately capturing the true uncertainties in this modeling;
+        modeling the physics of MBHB formation and evolution, with parameterizations for uncertain physical processes;
         \item
-        rapidly generating large simulated populations of MBHBs;
+        rapidly generating large simulated populations of MBHBs; and,
         \item
-        calculating the resulting GW signals from these populations.
+        calculating the resulting GW signals and EM signatures from these populations.
     \end{enumerate}
-    % (1) modeling the physics of MBHB formation and evolution while accurately capturing the true uncertainties in this modeling, (2) rapidly generating large simulated populations of MBHBs, and (3) calculating the resulting GW signals from these populations.
-
-    As in the case of high-frequency GWs detected by LIGO-Virgo \needcite{}, tremendous uncertainties exist in the physics of the formation and evolution of the binary sources.
-For individual MBHs, there are numerous, well-known scaling relationships between their properties and those of their host galaxies \needcite{MBH-Host relations}, but numerous questions remain such as where or not the observed sample of MBHs (i.e.~quasars and active galactic nuclei, AGN) accurately reflect the full population \needcite{MBH bias}, and how reliably these relationships extend to different host galaxies and redshift ranges.
-
-    For binary MBHs in particular, a variety of physical processes are required for the two MBHs to find the center of the post-merger host galaxy and eventually become gravitationally bound \needcite{BBR1980}.
-In particular, the two MBHs must interact with their host galaxy environments through dynamical friction, stellar scattering, accretion disk torques, and possibly three-body interactions with a third MBH component \needcite{}, before reaching small enough binary separations for GW emission to eventually dominate the final coalescence of the binary.
-Besides GW emission, none of these processes can be modeled exactly, necessitating the usage of simplified prescriptions which each containing numerous parametric uncertainties.
-The resolution of cosmological hydrodynamic simulations is far too coarse to resolve the MBH binary phase, with some simulations instantaneously merging nearby MBHs \needcite{illustris}, and others using sub-grid models to `resolve' down to the scale of $\sim kpc$ \needcite{Romulus/Tremmel, Ma+Hopkins}.
-
-    In addition to the theoretical uncertainties, there exists a substantial dearth of observational constraints.
-A growing number of \textit{candidate} MBHB systems have been identified through varying electromagnetic (EM) indicators in AGN \needcite{}, but a substantial fraction may be false positives \needcite{}.
-To date, no gravitationally bound MBH binary has yet to be confirmed, although examples exist of dual AGN at close projected separation \needcite{Rodriguez et al.
-2007}, and single AGN showing time-variability highly suggestive of a binary companion \needcite{OJ~287}.
 
@@
 
-    In this paper, we present the simulation methodologies and their implementations in \holodeck{}.
-We also use the framework to explore in more detail the range of plausible GW-strain spectral indices that would be detectable by PTAs.
-\todo{expand this paragraph}.
+    Tremendous uncertainties exist across every stage of the formation and evolution MBHBs.
+Even the formation process(es) of individual MBHs is almost entirely unconstrained \needcite{}.
+The three leading seeding mechanisms---light seeds: the rapid growth of stellar mass black holes, heavy seeds: from the `direct collapse' of pristine gas clouds, or in between: the runaway mergers of stars or stellar remnants in dense clusters---all struggle to explain the high masses and high occurrence rates of quasars in the very early Universe \needcite{}.
+LISA, scheduled to be launched in the mid 2030s, can directly detect MBHBs at their seeding mass and redshift \needcite{}.
+Accurate models for LISA sources are required to translate GW detections (or the lack of detections) into meaningful astrophysical constraints.
+
+    PTA sources, the most massive MBHBs in the relatively local Universe, are not particularly sensitive to seeding model, as the early history of their formation is washed out \needcite{}.
+For a given seeding model to be viable, it must be able to reproduce the properties of well characterized MBHs and AGN in the local Universe \needcite{}.
+For nearby systems, numerous, well-known scaling relations connect the properties of MBHs to their host galaxies \needcite{MBH-Host relations}.
+Even then, questions remain such as whether or not the observed sample of MBHs, i.e.~luminous ones, accurately reflect the full population \needcite{MBH bias}, and how reliably these relationships extend to different mass and redshift ranges, and differences in host galaxy type.
+
+    The properties and occurrence rates of MBH binaries are almost entirely unconstrained, as no confirmed examples of gravitationally bound MBHBs exist.
+A large number of \textit{dual} systems, not dynamically interacting, have been identified---typically at kiloparsec separations \needcite{dual AGN}, but one example which may be within the sphere of influence \needcite{Rogriguez et al.}.
+Additionally, many \textit{candidate} binaries have been suggested in the literature \needcite{Graham, Charisi}, however, numerous lines of reasoning suggest that many or most are likely false positives \needcite{Eracleous, Sesana, Kelley}.
+Still, handful of compelling examples persist \needcite{OJ287, etc?
+BoBCat?}.
+
+    Despite the lack of observational constraints, the evolution of MBHBs has been explored for decades.
+Following galaxy merger, a variety of physical processes are required for the two MBHs to find the center of the post-merger host galaxy and eventually become gravitationally bound \needcite{BBR1980}.
+In particular, the two MBHs must interact with their host galaxy environments through dynamical friction ($\sim\Kpc$) and stellar scattering ($\sim \pc$), and possibly a circumbinary accretion disk ($\lesssim \pc$), possibly triply interactions with a third MBH component ($\sim \pc$) \needcite{Loeb, Ryu, Bonetti}.
+Only after these interactions can binaries reach the separations where GW emission will become important, and drive their final coalescence.
+A lower-limit to binary lifetimes can be estimates by assuming that environmental processes are highly efficient at extracting energy and angular momentum from the two MBHs, in which case their separation could e-fold roughly every dynamical time, leading to total binary lifetimes of $\sim 100 \Myr$.
+
+    On the other extreme, there is no physical necessity for environmental interactions to be effective, and all binaries could stall.
+The existence of stable satellite galaxies demonstrates that, at least for small mass-ratio galaxy mergers ($q \equiv m_2/m_1 \leq 1$), dynamical friction is not always effective.
+The stellar scattering regime is typically believed to be the most challenging for inspiral.
+In general, roughly a mass of stars equivalent to that of the secondary MBH must be scattered for each e-folding of binary separation.
+Analytic estimates, using symmetric phase-space distributions of stars that are able to interact with the binary (the `loss cone'), typically find that inspiral would take longer than the Hubble time---the `final-parsec' problem \needcite{Merritt, Milosaljevic}.
+More detailed calculations that account for the tumultuous nuclear environments of realistic galaxies, typically find that the loss-cone can be continually refilled and inspiral can be efficient \needcite{}.
+Even with a full loss-cone, inspiral can take multiple billions of years, and large fractions of systems can still stall \needcite{Kelley2017}.
+
+    Torques from circumbinary accretion disks have been suggested to aid in the inspiral process at sub-parsec scales \needcite{Cuadra, Haiman, etc}.
+Galaxy mergers are known to drive substantial volumes of gas to galactic nuclei \needcite{Hernquist, etc}, but whether this persists long enough for binaries to reach the parsec regime is unclear.
+Nuclear bursts of star formation following galaxy mergers may even preferentially drive out the gas in galactic nuclei roughly coincident with when binaries form \needcite{may not be any}.
+Even if high mass circumbinary disks form, recent results from idealized hydrodynamic simulations have shown that circumbinary disk toques can actually drive binaries towards out-spiral \needcite{Munoz}.
+This is likely not a problem for PTA sources, higher mass MBHBs, both because of the regimes in which out-spiral is expected to dominate \needcite{outspiral population study}, and because high-mass elliptical galaxy mergers tend to be much drier.
+
+    The long inspiral times of MBHBs, particularly in systems which tend towards stalling, can easily become longer than the typical time between galaxy mergers.
+This suggests that systems with three MBHs could become important \needcite{Loeb}.
+Cosmological simulations have shown that such occurrences can indeed to be somewhat common \needcite{Kelley, Sayeb, Romulus, Astrid}, but they lack the resolution to explore triple interactions in detail.
+Semi-analytic modeling has suggested that such triples could drive a substantial fraction of `stalled' systems to merger \needcite{Bonetti, Ryu}.
+However, accurately modeling the triple dynamics at the same time as dissipative environmental interactions is very challenging.
+
+    Besides GW emission, none of these processes can be modeled exactly, necessitating the usage of simplified prescriptions which each containing numerous parametric uncertainties.
+The resolution of cosmological hydrodynamic simulations is far too coarse to resolve the MBH binary phase, with some simulations instantaneously merging nearby MBHs \needcite{illustris}, and others using sub-grid models to `resolve' down to the scale of $\sim kpc$ \needcite{Romulus/Tremmel, Ma+Hopkins}.
 
@@
 
+        \subsubsection{Parameters}
+
+            \begin{itemize}
+                \item \textbf{GSMF}
+                \begin{itemize}
+                    \item \citep{Tomczak+2014} - single schechter fits, little variation over redshift\\
+                    \item two
+                \end{itemize}
+
+                \item \textbf{GPF}
+                \begin{itemize}
+                    \item one \\
+                    \item two
+                \end{itemize}
+
+            \end{itemize}
+
 % =================================================================================================

### 2023-05-21 docs/holodeck-paper/src/packages.tex
Updates to introduction

@@
-\documentclass[useAMS, usenatbib]{mnras}
 \usepackage{color}
-\usepackage{booktabs}   % Fancier tables
+% \usepackage{booktabs}   % Fancier tables
 \usepackage{tabularx}
 \usepackage{float}
-\usepackage{totcount}
-\usepackage{cuted}
+% \usepackage{totcount}
+% \usepackage{cuted}
 % adds colors to colorspace (https://en.wikibooks.org/wiki/LaTeX/Colors)
@@
 \usepackage{graphicx}
-\usepackage{multirow}
 % Force figures to stay in their sections (tex.stackexchange.com/q/279/22806)
 \usepackage[section]{placeins}
-\usepackage{subfig}
+% \usepackage{subfig}
+\usepackage{multirow}
 \usepackage{multicol}
-\usepackage{widetext}
-\usepackage{upgreek}    % allow alternative greek letters like $\uptau$
+% \usepackage{widetext}
+% \usepackage{upgreek}    % allow alternative greek letters like $\uptau$
 \usepackage{enumitem}   % customize enumerate symbols
@@
 
-\usepackage{lipsum} % generate dummy text
 \usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage[thinc]{esdiff}  % easier/convenient derivatives and partial derivatives
 
@@
 
+% easier/convenient derivatives and partial derivatives, for example diffp for partials
+\usepackage[thinc]{esdiff}
+
 \usepackage{xfrac}
 
+\usepackage{microtype}   % fix overfull/underfull hbox issues
+
 % For ORCID iDs
-\usepackage{tikz,xcolor,hyperref}
-\definecolor{lime}{HTML}{A6CE39}
-\DeclareRobustCommand{\orcidicon}{
-	\begin{tikzpicture}
-	\draw[lime, fill=lime] (0,0)
-	circle [radius=0.16]
-	node[white] {{\fontfamily{qag}\selectfont \tiny ID}};
-	\draw[white, fill=white] (-0.0625,0.095)
-	circle [radius=0.007];
-	\end{tikzpicture}
-	\hspace{-2mm}
-}
+% \usepackage{tikz,xcolor,hyperref}
+% \definecolor{lime}{HTML}{A6CE39}
+% \DeclareRobustCommand{\orcidicon}{
+% 	\begin{tikzpicture}
+% 	\draw[lime, fill=lime] (0,0)
+% 	circle [radius=0.16]
+% 	node[white] {{\fontfamily{qag}\selectfont \tiny ID}};
+% 	\draw[white, fill=white] (-0.0625,0.095)
+% 	circle [radius=0.007];
+% 	\end{tikzpicture}
+% 	\hspace{-2mm}
+% }
 
-\foreach \x in {A, ..., Z}{\expandafter\xdef\csname orcid\x\endcsname{\noexpand\href{https://orcid.org/\csname orcidauthor\x\endcsname}
-			{\noexpand\orcidicon}}
-}
+% \foreach \x in {A, ..., Z}{\expandafter\xdef\csname orcid\x\endcsname{\noexpand\href{https://orcid.org/\csname orcidauthor\x\endcsname}
+% 			{\noexpand\orcidicon}}
+% }
 
@@
 % shift around the text on the page so looks good on Letter paper
-\voffset-.6in
-\hoffset0.2in
+% \voffset-.6in
+% \hoffset0.2in

### 2023-05-21 scripts/gp_trainer.py
Add frequency CLI argument

@@ -25,6 +25,13 @@ def setup_argparse():
         help="The path to the ini configuration file.",
     )
 
+    parser.add_argument(
+        "-f",
+        "--freq-ind",
+        type=int,
+        help="The frequency index from the library to use for training."
+    )
+
     args = parser.parse_args()
 
     return args
@@ -94,5 +101,6 @@ if __name__ == "__main__":
     args = setup_argparse()
     config = configparser.ConfigParser()
     config.read(args.config_path)
+    config['Training Options']['nfreqs'] = str(args.freq_ind)
 
     main(config)

### 2023-05-21 scripts/gp_trainer.py
Add training data type into gp filename

@@ -81,8 +81,14 @@ def main(config):
     datestr = datetime.now().strftime('%Y%m%d_%H%M%S')
 
     # Save the trained GP as a pickle to be used with PTA data!
-    gp_file = Path("trained_gp_" + spectra_file.parent.name + "_" + datestr +
-                   ".pkl")
+    # Add some extra info for specific frequency and what the GP was trained on
+    y_is_variance = train_opts.getboolean('train_on_variance', False)
+    if y_is_variance:
+        type_str = "std"
+    else:
+        type_str = "med"
+    gp_file = Path(f"trained_gp_{spectra_file.parent.name}_{type_str}_freq"\
+                   f"{train_opts.getint('nfreqs', None)}_{datestr}.pkl")
     loc_gp_file = spectra_file.parent / gp_file
 
     with open(loc_gp_file, "wb") as gpf:

### 2023-05-21 holodeck/gps/gp_utils.py
Change to single frequency

@@ -307,7 +307,9 @@ def get_gwb(spectra, nfreqs, test_frac=0.0, center_measure="median"):
     if VERBOSE:
         utils.my_print(f"setting aside {test_frac} of samples ({test_ind}) for testing, and choosing {nfreqs} frequencies")
 
-    gwb_spectra = gwb_spectra[test_ind:, :nfreqs, :]**2
+    # Changed the next line to only select one frequency. I'm reusing the `nfreqs`
+    # variable, but really it should be renamed something like `freq_ind`
+    gwb_spectra = gwb_spectra[test_ind:, nfreqs:nfreqs+1, :]**2
     xobs = xobs[test_ind:, :]
 
     # Find all the zeros and set them to be h_c = 1e-20
@@ -338,7 +340,7 @@ def get_gwb(spectra, nfreqs, test_frac=0.0, center_measure="median"):
     # The "y" data are the medians or means and errors for the spectra at each point in parameter space
     yobs = center.copy()  # mean.copy()
     yerr = err.copy()
-    gp_freqs = spectra["fobs"][:nfreqs].copy()
+    gp_freqs = spectra["fobs"][nfreqs:nfreqs+1].copy()
     gp_freqs *= YR
 
     # Find mean in each frequency bin (remove it before analyzing with the GP)

### 2023-05-21 scripts/combine_gps.py
Add script to combine GPs

@@ -1,0 +1,68 @@
+#!/usr/bin/env python3
+"""
+Module to combine indvidual GPs.
+
+Usage: python combine_gps.py -i <path_to_pkl> <maybe_another> <as_many_as_you_have> -o <save_merged_here>
+
+The saved PKLs are of the same format as the input: List[GaussProc].
+"""
+
+import argparse
+import operator
+import pickle
+from functools import reduce
+from pathlib import Path
+
+
+def parse_cli_args():
+    """Parse CLI arguments."""
+    parser = argparse.ArgumentParser()
+    parser.add_argument(
+        "-i",
+        "--input",
+        dest="input_gps",
+        type=Path,
+        help="Paths to PKLs of GPs. These need to be in order of increasing frequency!",
+        nargs="+",
+        required=True,
+    )
+    parser.add_argument(
+        "-o",
+        "--output",
+        dest="merged_gps",
+        type=Path,
+        help="Path to output merged PKL",
+        default=Path(Path.cwd() / "merged_gps.pkl"),
+    )
+    return parser.parse_args()
+
+
+def main():
+    # Get CLI args
+    options = parse_cli_args()
+
+    # Make sure the args make sense
+    if len(options.input_gps) < 2:
+        raise ValueError("Need at least two input files!")
+
+    unpkld_gps = []
+
+    # Iterate over the supplied PKLs
+    for pkl in options.input_gps:
+        with open(pkl, "rb") as f:
+            data = pickle.load(f)
+            unpkld_gps.append(data)
+
+    # Reduce to a single list from the list-of-lists
+    # We could have just taken the zeroeth element above when loading, but this
+    # method accounts for cases where we "chunk" the frequencies, which we might
+    # implement in the near future.
+    unpkld_gps = reduce(operator.add, unpkld_gps)
+
+    # Write out to disk
+    with open(options.merged_gps.with_suffix(".pkl"), "wb") as f:
+        pickle.dump(unpkld_gps, f)
+
+
+if __name__ == "__main__":
+    main()

### 2023-05-21 holodeck/param_spaces.py
Add param spaces 07D and 07E --- rotations of 07A

@@ -541,6 +541,60 @@ class PS_Uniform_07C(PS_Generic_2):
         return super().model_for_params(params, sam_shape=sam_shape, new_def_params=new_def_params)
 
 
+class PS_Uniform_07D(PS_Generic_2):
+    """Same as PS_Uniform_07A with a rotation of -0.5
+    """
+
+    def __init__(self, log, nsamples, sam_shape, seed):
+        super().__init__(
+            log, nsamples, sam_shape, seed,
+            hard_time=PD_Uniform(0.1, 11.0),   # [Gyr]
+            gsmf_phi0=PD_Uniform(-3.5, -1.5),
+            gsmf_mchar0_log10=PD_Uniform(10.5, 12.5),   # [log10(Msol)]
+            mmb_mamp_log10=PD_Uniform(+7.6, +9.0),   # [log10(Msol)]
+            mmb_scatter_dex=PD_Uniform(+0.0, +0.9),
+        )
+
+    @classmethod
+    def model_for_params(cls, params, sam_shape=None, new_def_params={}):
+        # NOTE: these should be the same as the default case, just duplicating them here for clarity
+        new_def_params = dict(
+            hard_gamma_inner=-1.0 - 0.5,
+            hard_rchar=100.0,               # [pc]
+            hard_gamma_outer=+1.5 - 0.5,
+            hard_sepa_init=1e4,     # [pc]
+        )
+        return super().model_for_params(params, sam_shape=sam_shape, new_def_params=new_def_params)
+
+
+
+class PS_Uniform_07E(PS_Generic_2):
+    """Same as PS_Uniform_07A with a rotation of +0.5
+    """
+
+    def __init__(self, log, nsamples, sam_shape, seed):
+        super().__init__(
+            log, nsamples, sam_shape, seed,
+            hard_time=PD_Uniform(0.1, 11.0),   # [Gyr]
+            gsmf_phi0=PD_Uniform(-3.5, -1.5),
+            gsmf_mchar0_log10=PD_Uniform(10.5, 12.5),   # [log10(Msol)]
+            mmb_mamp_log10=PD_Uniform(+7.6, +9.0),   # [log10(Msol)]
+            mmb_scatter_dex=PD_Uniform(+0.0, +0.9),
+        )
+
+    @classmethod
+    def model_for_params(cls, params, sam_shape=None, new_def_params={}):
+        # NOTE: these should be the same as the default case, just duplicating them here for clarity
[…]

### 2023-06-05 scripts/gen_lib_ss.py
Recovered gen_lib_ss.py from prev commit.

@@ -193,7 +193,7 @@ def _setup_argparse(*args, **kwargs):
                         help='Number of frequency bins', default=DEF_NUM_FBINS)
     parser.add_argument('-s', '--shape', action='store', dest='sam_shape', type=int,
                         help='Shape of SAM grid', default=DEF_SAM_SHAPE)
-    parser.add_argument('-l', '--nloudest', action='store', dest='nloudest', type=int,
+    parser.add_argument('-l', '--loudest', action='store', dest='nloudest', type=int,
                         help='Number of loudest single sources', default=1)
     parser.add_argument('-p', '--pars', action='store', dest='get_pars', type=int,
                         help='Whether or not to get pars', default=0)

### 2023-06-08 holodeck/accretion.py
eccentricity evolution from Siwek+23b working

@@ -48,8 +48,8 @@ class Accretion:
             """ Indices where separation is less than or equal to a parsec """
             ind_sepa = evol.sepa[:, step] <= PC
         else:
-            """ Indices where separation is less than or equal to 10 kilo-parsec """
-            ind_sepa = evol.sepa[:, step] <= 10**4 * PC
+            """ Indices where separation is less than or equal to 100 kilo-parsec """
+            ind_sepa = evol.sepa[:, step] <= 10**5 * PC
 
         """ Set total accretion rates to 0 when separation is larger than 1pc or 10kpc,
             depending on subpc switch applied to accretion instance """

### 2023-06-08 holodeck/evolution.py

@@ -938,7 +938,6 @@ class Evolution:
         dedt = None if self.eccen is None else np.zeros_like(dadt)
 
         for ii, hard in enumerate(self._hard):
-            print("hard = ", hard)
             _hard_dadt, _ecc = hard.dadt_dedt(self, step)
             dadt[:] += _hard_dadt
             if self._debug:    # nocov
@@ -1256,9 +1255,9 @@ class CBD_Torques(_Hardening):
 
         if evo._acc is None:
             """ If no accretion modules is supplied, use an Eddington fraction for now """
-            total_mass = mass[:,step,0] + mass[:,step,1]
+            total_mass = mass[:,0] + mass[:,1]
             accretion_instance = holo.accretion.Accretion(f_edd = self.f_edd, subpc=self.subpc)
-            mdot = accretion_instance.mdot_eddington(total_mass)
+            mdot = accretion_instance.mdot_total(evo, step)
         if evo._acc is not None:
             """ An instance of the accretion class has been supplied,
                 and binary masses are evolved through accretion
@@ -1267,8 +1266,11 @@ class CBD_Torques(_Hardening):
 
         dadt, dedt = self._dadt_dedt(mass, sepa, eccen, mdot)
 
+        """ CURRENTLY WE CANNOT USE +ve dadt VALUES, SO WE SET THEM TO 0 """
         inds_dadt_pos = dadt > 0
         dadt[inds_dadt_pos] = 0.0
+        inds_dadt_nan = np.isnan(dadt)
+        dadt[inds_dadt_nan] = 0.0
 
         return dadt, dedt
 
@@ -1294,9 +1296,18 @@ class CBD_Torques(_Hardening):
 
         """
         mass = np.atleast_2d(mass)
+        mtot = mass[:,0] + mass[:,1]
+        """ MASS RATIO """
+        m1 = mass[:, 0]
+        m2 = mass[:, 1]
+        mrat = m2/m1
+        """ secondary and primary can swap indices. need to account for that and reverse the mass ratio """
+        inds_rev = mrat > 1
+        mrat[inds_rev] = 1./mrat[inds_rev]
+        """ SEPARATION """
         sepa = np.atleast_1d(sepa)
+        """ ECCENTRICITY """
[…]

### 2023-06-08 docs/holodeck-paper/ms.tex

--- 2023-01-31 docs/holodeck-paper/ms.tex
@@
 % =================================================================================================
+\documentclass[useAMS, usenatbib]{mnras}
 
@@
 
-\title[Short Title]{Very long and verbose title}
+\title[holodeck: MBH binary populations]{Enter the \texttt{holodeck}: massive black hole binary population synthesis for gravitational wave calculations}
 \author[NANOGrav]{the NANOGrav collaboration}
-% \author[L.Z.~Kelley et al.]{
-%     Luke Zoltan Kelley$^{1}$\thanks{E-mail:lzkelley@northwestern.edu}\orcidA{} \\
-%     $^{1}$ \begin{minipage}[t]{\linewidth} Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA), and\\Department of Physics \& Astronomy, Northwestern University, Evanston, IL 60208
-%     \end{minipage}
-% }
 
@@
 
-        It was shown by \citet[][Eq.~5]{Phinney-2001} that the GWB spectrum, can be expressed in terms of the GW characteristic strain ($\hc$),
+        It was shown by \citet[][]{Phinney-2001} that the GWB spectrum can be calculated by relating the current energy-density of GWs to the integrated GW energy produced by sources over all of cosmic time,
         \begin{equation}
-            % \label{eq:}
-            \hc^2(f) = \frac{4G}{\pi c^2 f} \int dz \; \frac{d\ndens}{dz} \, \scale{d\egw}{d f_p} \bigg|_{f_p=f\lr{1+z}}.
+            \label{eq:soltan_analog}
+            \hc^2(f) = \frac{4G}{\pi c^2 f} \int dz \; \diffp{\ndens}{{z}} \, \diffp{\egw}{{f_r}} \bigg|_{f_r=f\lr{1+z}}.
         \end{equation}
-        We can convert from the GW energy-spectrum ($\frac{d\egw}{d f_p}$) to GW spectral strain using Eq.~\ref{eq:gw_energy_spectrum}~\&~\ref{eq:strain_lum}, and rewrite this as,
+        Converting from the rest-frame frequency $f_r$ to observer-frame $f$ in the term in front of the integral cancels out the redshifting of gravitons between their emission and detection.
+Here the comoving, volumetric number density of sources is \mbox{$\ndens \equiv \diffp{N}{{V_c}}$}, and the GWB spectrum is described in terms of the characteristic strain $\hc$.
+We can convert from the GW energy-spectrum integrated over each sources lifetime, $\diffp{\egw}{{f_p}}$, to GW spectral strain using Eq.~\ref{eq:gw_energy_spectrum}~\&~\ref{eq:strain_lum}, and rewrite the GWB spectrum as,
         \begin{equation}
@@
 
-        A binary is described by its total mass $M$, mass ratio $q \equiv m_2 / m_1$ (s.t.~$q\leq1$), (proper) semi-major axis $a$, and rest-frame orbital frequency $f_r$.
+        A binary is described by its total mass $M$, mass ratio $q \equiv m_2 / m_1$ (s.t.~$q\leq1$), (proper) semi-major axis $a$, and rest-frame orbital frequency $f_p$.
 The characteristic gravitational wave mass is the chirp mass, $\mchirp \equiv M \, q^{3/5} \, \lr[-6/5]{1+q}$.

### 2023-06-08 holodeck/evolution.py
add required imports

@@ -84,17 +84,19 @@ import scipy as sp
 import scipy.interpolate   # noqa
 from scipy.interpolate import RectBivariateSpline
 
-
 import kalepy as kale
 
 import holodeck as holo
-from holodeck import utils, cosmo, log
-from holodeck.constants import PC
+from holodeck import utils, cosmo, log, _PATH_DATA
+from holodeck.constants import PC, GYR, NWTG, PC, MSOL, YR
 from holodeck.hardening import _Hardening
 # from holodeck import accretion
 
 _MAX_ECCEN_ONE_MINUS = 1.0e-6
-
+#: number of influence radii to set minimum radius for dens calculation
+_MIN_DENS_RAD__INFL_RAD_MULT = 10.0
+_DEF_TIME_DELAY = (5.0*GYR, 0.2)   #: default delay-time parameters, (mean, stdev)
+_SCATTERING_DATA_FILENAME = "SHM06_scattering_experiments.json"
 
 # =================================================================================================
 # ====    Evolution Class    ====

### 2023-06-14 holodeck/anisotropy.py
added anis.lib_anisotropy_split function that will calculate the nbest samples Cl's in 'split' chunks.

@@ -20,6 +20,43 @@ LMAX = 8
 HC_REF15_10YR = 11.2*10**-15 
 
 def healpix_map(hc_ss, hc_bg, nside=NSIDE):
+    """ Build mollview array of hc^2/dOmega for a healpix map
+    
+    Parameters
+    ----------
+    hc_ss : (F,R,L) NDarray
+        Characteristic strain of single sources.
+    hc_bg : (F,R) NDarray
+        Characteristic strain of the background.
+    nside : integer
+        number of sides for healpix map.
+
+    Returns
+    -------
+    moll_hc : (NPIX,) 1Darray
+        Array of h_c^2/dOmega at every pixel for a mollview healpix map.
+    
+    NOTE: Could speed up the for-loops, but it's ok for now.
+    """
+
+    npix = hp.nside2npix(nside)
+    nfreqs = len(hc_ss)
+    nreals = len(hc_ss[0])
+    nloudest = len(hc_ss[0,0])
+
+    # spread background evenly across pixels in moll_hc
+    moll_hc = np.ones((nfreqs,nreals,npix)) * hc_bg[:,:,np.newaxis]**2/(npix) # (frequency, realization, pixel)
+
+    # choose random pixels to place the single sources
+    pix_ss = np.random.randint(0, npix-1, size=nfreqs*nreals*nloudest).reshape(nfreqs, nreals, nloudest)
+    for ff in range(nfreqs):
+        for rr in range(nreals):
+            for ll in range(nloudest):
+                moll_hc[ff,rr,pix_ss[ff,rr,ll]] = (moll_hc[ff,rr,pix_ss[ff,rr,ll]] + hc_ss[ff,rr,ll]**2)
+                
+    return moll_hc
+
+def healpix_map_oldhc(hc_ss, hc_bg, nside=NSIDE):
     """ Build mollview array of strains for a healpix map
     
     Parameters
@@ -56,6 +93,44 @@ def healpix_map(hc_ss, hc_bg, nside=NSIDE):
                                                           + hc_ss[ff,rr,ll]**2)
                 
     return moll_hc
+
+
+def healpix_hcsq_map(hc_ss, hc_bg, nside=NSIDE):
+    """ Build mollview array of strains for a healpix map
+    
+    Parameters
+    ----------
+    hc_ss : (F,R,L) NDarray
+        Characteristic strain of single sources.
[…]

### 2023-06-14 scripts/anis_lib_script.py

@@ -15,6 +15,7 @@ DEF_NSIDE = anisotropy.NSIDE
 DEF_LMAX = anisotropy.LMAX
 DEF_NBEST = 100
 DEF_NREALS = 50
+DEF_SPLIT = 1
 
 def _setup_argparse():
     parser = argparse.ArgumentParser()
@@ -28,14 +29,16 @@ def _setup_argparse():
                         help='number of best ranked samples to calculate spherical harmonics for')
     parser.add_argument('-r', '--nreals', action='store', dest='nreals', type=int, default=DEF_NREALS,
                         help='number of realizations to use ranking and harmonics, must be less than or equal to nreals of library')
+    parser.add_argument('--split', action='store', dest='split', type=int, default=DEF_SPLIT,
+                        help='number of sections to split nbest calculations into')
     
     args = parser.parse_args()
     return args
 
 def main():
     args = _setup_argparse()
-    anisotropy.lib_anisotropy(args.lib_path, hc_ref_10yr=HC_REF_10YR, nbest=args.nbest,
-                              nreals=args.nreals, lmax=args.lmax, nside=args.nside)
+    anisotropy.lib_anisotropy_split(args.lib_path, hc_ref_10yr=HC_REF_10YR, nbest=args.nbest,
+                              nreals=args.nreals, lmax=args.lmax, nside=args.nside, split=args.split)
 
 if __name__ == "__main__":
     main()

### 2023-06-14 holodeck/anisotropy.py
Added split to saved npz array for lib anisotropy.

@@ -337,7 +337,7 @@ def lib_anisotropy_split(lib_path, hc_ref_10yr=HC_REF15_10YR, nbest=100, nreals=
         print('Saving npz file: ', output_name)
         np.savez(output_name,
                 nsort=nsort, fidx=fidx, hc_ref=hc_ref, ss_shape=shape,
-            moll_hc_best=moll_hc_best, Cl_best=Cl_best, nside=nside, lmax=lmax, fobs=fobs)
+            moll_hc_best=moll_hc_best, Cl_best=Cl_best, nside=nside, lmax=lmax, fobs=fobs, split=split)
     
 
         # ---- plot median Cl/C0

### 2023-06-17 docs/holodeck-paper/ms.tex
Updates to Methods sec 2

@@
 
-    Massive black holes (MBHs) are known to occupy the centers of most if not all massive galaxies \needcite{}.
+    Massive black holes (MBHs) occupy the centers of most if not all massive galaxies \needcite{}.
     Binaries of MBHs are believed to form following the merger of massive galaxies \needcite{}.
@@
 
-        It was shown by \citet[][]{Phinney-2001} that the GWB spectrum can be calculated by relating the current energy-density of GWs to the integrated GW energy produced by sources over all of cosmic time,
+        The GWB spectrum can be calculated by relating the energy-density of GWs in the local Universe to the integrated GW energy produced by sources over all of cosmic time \citet{Phinney-2001},
         \begin{equation}
             \label{eq:soltan_analog}
-            \hc^2(f) = \frac{4G}{\pi c^2 f} \int dz \; \diffp{\ndens}{{z}} \, \diffp{\egw}{{f_r}} \bigg|_{f_r=f\lr{1+z}}.
+            \hc^2(\fobs) = \frac{4G}{\pi c^2 \fobs} \int dz \; \diffp{\ndens}{{z}} \, \diffp{\egw}{{f_r}} \bigg|_{\frst=\fobs\lr{1+z}}.
         \end{equation}
-        Converting from the rest-frame frequency $f_r$ to observer-frame $f$ in the term in front of the integral cancels out the redshifting of gravitons between their emission and detection.
+        Converting from the rest-frame frequency $\frst$ to observer-frame $\fobs$ in front of the integral cancels out the redshifting of gravitons between their emission and detection.
 Here the comoving, volumetric number density of sources is \mbox{$\ndens \equiv \diffp{N}{{V_c}}$}, and the GWB spectrum is described in terms of the characteristic strain $\hc$.
-We can convert from the GW energy-spectrum integrated over each sources lifetime, $\diffp{\egw}{{f_p}}$, to GW spectral strain using Eq.~\ref{eq:gw_energy_spectrum}~\&~\ref{eq:strain_lum}, and rewrite the GWB spectrum as,
+We can convert from the GW energy-spectrum of each source, $\diffp{\egw}{{\frst}}$, to GW spectral strain using Eq.~\ref{eq:gw_energy_spectrum}~\&~\ref{eq:strain_lum}, and rewrite the GWB spectrum as,
         \begin{equation}
             \label{eq:gwb_sa}
-            \hc^2(f) = \int_0^\infty \!\! dz \, \sum_n^\infty \; \frac{d\ndens}{dz} \, \hsn^2\lr{f_p} \cdot 4\pi c \, d_c^2 \cdot \lr{1+z} \cdot \thard(f_p) \, \bigg|_{f_p = f(1+z)/n}.
+            \hc^2(f) = \int_0^\infty \!\!
+dz \, \sum_n^\infty \diffp{\ndens}{{z}} \, \hsn^2\lr{\frstorb} \cdot 4\pi c \, \distcom^2 \cdot \lr{1+z} \cdot \thardf(\frstorb) \, \bigg|_{\frstorb = \fobs(1+z)/n}.
         \end{equation}
-        In the case of purely GW-driven binary hardening, and circular binary orbits, this can be simplified to \citep[][Eq.~11]{Phinney-2001},
+        Here, $\distcom$ is the comoving distance to redshift $z$.
+The characteristic strain (left-hand side) \needcite{} is more appropriate for calculating detectability, while the GW spectral strain $\hsn{}$, is directly related to a source's GW luminosity \needcite{} (\eqref{eq:strain_lum}).
+The subscript $n$ denotes the frequency harmonic.
+Binaries in eccentric orbits emit GWs at all integer harmonics of their orbital frequency ($\frstorb$).
+For circular obits, binaries only emit at the $n=2$ harmonic.
+Eq.~\ref{eq:gwb_sa} also includes the frequency residence time / hardening time, \mbox{$\thardf \equiv \frstorb / \lr{\partial \frstorb / \partial t}$}.
+
+        In the case of purely GW-driven binary hardening\footnote{i.e.~using Eq.~\ref{eq:gw_hard}~\&~\ref{eq:hard_time}.}, and circular binary orbits\footnote{i.e.~using Eq.~\ref{eq:gw_strain_amp_circ}}, this can be simplified to \citep[][Eq.~11]{Phinney-2001},
         \begin{equation}
@@
         \end{equation}
-        This ideal, power-law GWB is often expressed as, \mbox{$\hc(f) = A_{\yr} \cdot \lr[-2/3]{f \cdot \yr}$}.
+        This expression motivates the common idealized approximation that,
+        \begin{equation}
+            \label{eq:char_strain_plaw}
+            \hc(\fobs) \approx A_\tr{ref} \cdot \scale[\alpha]{\fobs}{\fref},
+        \end{equation}
+        and $\alpha \approx -2/3$ for an idealized GWB characteristic strain spectrum\footnote{In terms of timing residual power spectral density, the idealized power-law index is $-13/3$, see Eq.~\ref{eq:psd_char_strain}.}.
+The applicability of this approximation is discussed below.
 
-        The redshift evolution of the comoving number density of sources ($\ndens \equiv dN/dV_c$) can be related to the number of sources in a given frequency interval as \citep[][Eq.~6]{Sesana+2008},
+        Instead of assuming an idealized shape of the GWB spectrum, we would like an expression which calculates the instantaneous number of binaries emitting at any arbitrary frequency interval.
+We can do this by first relating the comoving number density of sources ($\ndens \equiv dN/dV_c$) to the total number of binaries in the observer's past light cone \citep{Haehnelt-1994}, and then connecting redshift evolution to the time-evolution of binaries over orbital frequency \citep{Jaffe+Backer-2003}.
+We can then write \citep{Sesana+2008},
         \begin{equation}
@@
         \end{equation}
-        Plugging \eqref{eq:number_density_to_number_frequency} into \eqref{eq:gwb_sa}, and utilizing the expressions from Eqs.~\ref{eq:gw_hard},~\ref{eq:hard_time},~\ref{eq:comvol}~\&~\ref{eq:redshift_time}, to write:
+        This is the crucial expression for modeling a full population of evolving binaries.
+Plugging \eqref{eq:number_density_to_number_frequency} into \eqref{eq:gwb_sa}\footnote{And utilizing the expressions from Eqs.~\ref{eq:gw_hard},~\ref{eq:hard_time},~\ref{eq:comvol}~\&~\ref{eq:redshift_time},}, we can confirm that the local characteristic strain is indeed the total spectral strain in a logarithmic frequency interval, emitting by all binaries in the observer's past light cone:
         \begin{equation}
@@
         \end{equation}
+        In general, implementations of binary populations will utilize some fixed number of binaries in a finite simulation volume; or alternatively be described as a number density over a finite grid, e.g., total mass, mass ratio, redshift, and orbital frequency.
+In the latter case the derivative over redshift in \eqref{eq:pop_discrete_from_continuous} should be replaced with one over all parameters of interest, i.e.~\mbox{$d\ndens / dz \rightarrow d^3\ndens / dM \, dq \, dz$}.
+In either case, we will procedurally always be calculating the left-hand side of \eqref{eq:pop_discrete_from_continuous} from the right-hand side.
+            \todo{Introduce 'discrete' vs. 'continuous' terminology and SAMs??}
 
-        In general, implementations of binary populations will be described over some finite grid, often in a parameter space of total mass, mass ratio, redshift, and orbital frequency.
-In this case the derivative over redshift in \eqref{eq:pop_discrete_from_continuous}, should be replaced with one over all parameters of interest, e.g.~\mbox{$d\ndens / dz \rightarrow d^3\ndens / dM \, dq \, dz$}.
-All additional parameters are treated the same in this context, so for brevity we will only explicitly include redshift.
+        We will often be interested in quantities which depend on the number of binaries over ranges of parameter space.
+For example, for the GWB in the PTA, we are interested in the number of binaries over finite width frequency bins, $\deltalnf = \Delta \ln\!\frstorb$.
+These quantities will be calculated using some numerical integration procedure, for example, in the simplest case: \mbox{$\langle \Delta N \rangle \approx \lr{d^2 N / dz \, d\ln\!f_p} \, \cdot \Delta z \cdot \Delta \ln\!f_p$}.
+For distributions calculated over grids with additional parameters, those will also need to be `integrated' over (e.g.~including $1/\partial M$, $1/\partial q$, etc.
+and also $\Delta M$, $\Delta q$, etc.).
+We write this with angle-brackets to denote it as an expectation value for two reasons.
+First, it does not yet account for the intrinsic discreteness of binaries.
+Second, it does not account for random variance from how many binaries happen to exist in a given volume (either physical volume or parameter-space volume).
+We address both issues by drawing specific `realizations' of populations.
+Assuming they are uniformly distributed, they will be Poisson distributed such that, $\Delta N \in \poisson(\langle \Delta N \rangle)$.
 
-        Eventually the end goal of binary modeling is often to determine the number of binaries over some range of parameters.
+        For a discretized population of binaries, the GWB spectrum is then calculated as,
+        \begin{equation}
+            \label{eq:gwb_discrete}
+            \hc^2(\Delta \ln f) = \sum_\textrm{bins} \sum_n^\infty \frac{\Delta N(\frstorb,M_a,q_b,z_c,\dots)}{\Delta \ln \frstorb} \, \hsn^2(\frstorb,M_a,q_b,z_c,\dots).
+        \end{equation}
+        The sum over `bins' corresponds to the bins $a$ in total mass ($M_a$), bins $b$ in mass ratio ($q_b$), etc.
+        This is all that is needed for populations calculated from number-density distributions over grids of parameter space, as in the case of semi-analytic models (SAMs) which are described in the following section.
+An additional simplification can be made, however, in the case of discrete binaries in finite volume simulations, such as cosmological hydrodynamic simulations.
+The most analogous procedure would be to group such binaries into similar bins of parameter space, calculate a number density, and translate into the number of binaries in the observer's light cone using \eqref{eq:pop_discrete_from_continuous}.
+Because a finite range of parameters would still need to be considered before Poisson sampling, this amount to dividing by and then multiplying by, finite widths of parameter space.
+This procedure can be skipped, and instead calculate for each binary ($i$):
+        \begin{equation}
+            \label{eq:expectation_number_of_single_binary}
+            \langle \Delta N \rangle_i = \frac{4\pi c \, \distcom^2 \, \thardf}{V_c} \cdot (1+z_i) \cdot \Delta \ln \frstorb.
+        \end{equation}
+        In this case, $\thardf$ and $\hsn$ are the hardening time and strain amplitude for each source, and $\langle \Delta N \rangle_i$ represents the expectation number of binaries in the Universe implied by each source in the finite volume.
+        \todo{Is this a biased estimator?  Is there a way to correct for that?  Because we're estimating an expectation value, based on a single Poisson draw from that expectation value.}
+
+        % In the latter case the derivative over redshift in \eqref{eq:pop_discrete_from_continuous} should be replaced with one over all parameters of interest, i.e.~\mbox{$d\ndens / dz \rightarrow d^3\ndens / dM \, dq \, dz$}.
+And each specific function will generally also be dependent on similar parameters, e.g., $\hc = \hc(M,q,z,f)$.
+        % All additional parameters are treated the same in this context, so for brevity we will only explicitly include redshift.
+
+        % The end goal of binary modeling is typically to determine the number of binaries over certain ranges of parameters.
 When calculating gravitational wave signals, for example, the number of binaries across a target frequency interval is desired.
@@
 
-        In general, the properties of \textit{every} binary in the Universe are not known, and thus the MC calculation cannot be performed directly.
-The typical solution is to use some version of \eqref{eq:number_density_to_number_frequency} to convert from a number-density (e.g.~$d\ndens/dz$) into a discrete distribution of individual sources (e.g.~$d^2 N / dz \, d\ln f_r$).
+        % The typical solution is to use some version of \eqref{eq:number_density_to_number_frequency} to convert from a number-density (e.g.~$d\ndens/dz$) into a discrete distribution of individual sources (e.g.~$d^2 N / dz \, d\ln f_r$).
 The number-density is often constructed from either (semi-)analytic models which specify $\ndens(M,q,z,\dots)$ explicitly, or from simulations/observations of sources in a finite volume.
@@
 
-        For a finite population of binaries, the GWB spectrum is only defined over finite bins of frequency.
+        % For a finite population of binaries, the GWB spectrum is only defined over finite bins of frequency.
 The definition of characteristic strain is related to the GW energy across a logarithmic frequency bin, so \eqref{eq:gwb_mc} should really be written as,
-        \begin{equation}
-            \label{eq:gwb_discrete}
-            \hc^2(\Delta \ln f) = \sum_\textrm{bins} \sum_n^\infty \frac{\Delta N(f_p,M_a,q_b,z_c,\dots)}{\Delta \ln f_p} \, \hsn^2(f_p,M_a,q_b,z_c,\dots).
-        \end{equation}
-        Here we have written out explicitly that the number of binaries, $\Delta N(f_p,M_a,q_b,z_c,\dots)$, is the number sources with frequencies inside $\Delta \ln f_p$, typically over bins of mass, mass-ratio, redshift, etc.
+        % \begin{equation}
+        %     \label{eq:gwb_discrete}
+        %     \hc^2(\Delta \ln f) = \sum_\textrm{bins} \sum_n^\infty \frac{\Delta N(f_p,M_a,q_b,z_c,\dots)}{\Delta \ln f_p} \, \hsn^2(f_p,M_a,q_b,z_c,\dots).
+        % \end{equation}
+        % Here we have written out explicitly that the number of binaries, $\Delta N(f_p,M_a,q_b,z_c,\dots)$, is the number sources with frequencies inside $\Delta \ln f_p$, typically over bins of mass, mass-ratio, redshift, etc.
 Discretizing and plugging into Eq.~\ref{eq:number_density_to_number_frequency} (or by examining Eq.~\ref{eq:gwb_sa}), we can then write,
-        \begin{equation}
-            \label{eq:expectation_number_of_binaries}
-            \langle \Delta N \rangle = \Delta \ndens \cdot 4\pi c \, \distcom^2 \cdot (1+z) \cdot \thard \cdot \Delta \ln f_p.
-        \end{equation}
-        Note that we have again dropped the written dependence on bin values $(M_a, q_b, \dots)$, but they are still implicitly present.
+        % \begin{equation}
+        %     \label{eq:expectation_number_of_binaries}
+        %     \langle \Delta N \rangle = \Delta \ndens \cdot 4\pi c \, \distcom^2 \cdot (1+z) \cdot \thard \cdot \Delta \ln f_p.
+        % \end{equation}
+        % Note that we have again dropped the written dependence on bin values $(M_a, q_b, \dots)$, but they are still implicitly present.
 This expression represents an \textit{expectation value} based on extrapolating the given number-density of binaries.
@@
 
-        The way \eqref{eq:expectation_number_of_binaries} is handled in practice will still vary somewhat depending on how $\ndens$ is specified.  If a semi-analytic model is being used, then often
-        $\Delta \ndens = \left[ d^3 \ndens(M,q,z) / dM \, dq \, dz \right] \cdot \Delta M \, \Delta q \, \Delta z$.
+        % The way \eqref{eq:expectation_number_of_binaries} is handled in practice will still vary somewhat depending on how $\ndens$ is specified.  If a semi-analytic model is being used, then often
+        % $\Delta \ndens = \left[ d^3 \ndens(M,q,z) / dM \, dq \, dz \right] \cdot \Delta M \, \Delta q \, \Delta z$.
 In the case of a finite volume population, the procedure can be simplified.
@@
 Instead, we can imagine shrinking each of our parameter-space bins such that each one captures a single source, in which case we can write,
-        \begin{equation}
-            \label{eq:expectation_number_of_single_binary}
-            \langle \Delta N \rangle_\textrm{source} = \frac{4\pi c \, \distcom^2 \, \thard}{V_c} \cdot (1+z) \cdot \Delta \ln f_p.
-        \end{equation}
-        In this case, $\thard$ and $\hsn$ in \eqref{eq:gwb_discrete}, are the hardening time and strain amplitude for each source, and $\langle \Delta N \rangle_\textrm{source}$ represents the expectation number of binaries in the Universe implied by each source in the finite volume.
-
-    % \subsection{??Other??}
-
-    %     Let $N(M,q,z,\frst)$ denote the number of binaries up to a total mass $M$, mass ratio $q$, redshift $z$, and rest-frame orbital-frequency $\frst$.
-    %     We can calculate the characteristic strain from an ensemble of GW sources as \citep[][Eqs.~5/8]{Phinney-2001},
-    %         \begin{align}
-    %             h_c^2 = \int_0^\infty \!\! dz \; \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} \; \hs^2\lr{\frst}.
-    %         \end{align}
-    %     In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$, for some comoving volume $V_c$, extrapolated into an observer's entire past light-cone.
-These can be related as \citep[][Eq.~6]{Sesana+2008},
-    %         \begin{align}
-    %             \label{eq:num_num_dens}
-    %             \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} = \frac{d^3 n_c}{dM \, dq \, dz} \frac{dz}{dt} \frac{dt}{d\ln\!\frst} \frac{d V_c}{dz}.
-    %         \end{align}
-    %     Note that the total number of binaries in a given frequency interval depends not only on the number density, but also their rate of frequency-evolution.
-The faster binaries evolve through a given frequency interval, the fewer binaries will occupy that interval on average.
-This is typically described by the binary `hardening' timescale (or `residence time'), given by the expression,
         % \begin{equation}
-    %         \label{eq:hardening}
-    %         \thardf \equiv - \diffp{t}{{\ln\!\frst}} = - \frac{\frst}{\partial \frst / \partial t}.
+        %     \label{eq:expectation_number_of_single_binary}
+        %     \langle \Delta N \rangle_\textrm{source} = \frac{4\pi c \, \distcom^2 \, \thard}{V_c} \cdot (1+z) \cdot \Delta \ln f_p.
         % \end{equation}
-
-    %     % Plugging in the relevant cosmographic relationships (see \secref{sec:app_eqs}) we have,
-    %     %     \begin{align}
-    %         %         h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
-    %         %     \end{align}
-    %         % The challenge in modeling astrophysical sources of GWs is in calculating their number density.  Numerous approaches are possible, described in the following sections.
-
-    %     Plugging in \eqref{eq:hardening} along with the relevant cosmographic relationships (see \secref{sec:app_eqs}) into \eqref{eq:num_num_dens} we have,
-    %     \begin{equation}
-    %         \frac{d^4 N}{dM \, dq \, dz \, d\ln\!\frst} =
-    %             \frac{d^3 n_c}{dM \, dq \, dz} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
-    %     \end{equation}
+        % In this case, $\thard$ and $\hsn$ in \eqref{eq:gwb_discrete}, are the hardening time and strain amplitude for each source, and $\langle \Delta N \rangle_\textrm{source}$ represents the expectation number of binaries in the Universe implied by each source in the finite volume.
 
@@
         \begin{subequations}
+        \label{eq:gw_energy_spectrum}
         \begin{align}
-            \label{eq:gw_energy_spectrum}
             \frac{d \egw}{d \frst} & = \int \frac{dt_r}{df_p} \, \sum_{n=1}^\infty \, \lgwn(f_p) \; \delta(\frst - n f_p) \; d f_p \\
@@
         where the power-radiated at each harmonic can be expressed as is \citep[][Eq.~2.2]{enoki2007a},
+        \begin{subequations}
         \begin{align}
@@
         \end{align}
+        \end{subequations}
         Here, the GW frequency distribution function is,
@@
         The GW spectral strain amplitude at a given harmonic can be related to the (rest-frame) luminosity as \citep[][Eq.~2.1]{Finn+Thorne-2000},
+        \begin{subequations}
+        \label{eq:strain_lum}
         \begin{align}
-            \label{eq:strain_lum}
             \hsn^2 & = \frac{G}{c^3} \, \scale[2]{2}{n} \, \frac{\lgwn}{\lr[2]{2 \pi \, \frstorb} \, \distcom^2}, \\
@@
         \end{align}
+        \end{subequations}
         We always take the chirp mass to be an intrinsic (rest-frame) property of the binary, however an `observer-frame' chirp-mass can also be used.
@@
         \end{equation}
+
+        \todo{The spectral strain can be related to characteristic strain as...}
+
+        Instead of characteristic strain, the GWB spectrum at low-frequencies is often characterized in terms of the power spectral density (PSD) of timing residuals\footnote{The PSD is often denoted by $S_h$ in the GW literature.
+Unfortunately, the PSD of GW strain is also often used and referred to using the same symbols and terminology, while the physical quantity is quite distinct \needcite{}.} ($\psd$)
+        \begin{equation}
+            \label{eq:psd_char_strain}
+            \psd(f) = \frac{\hc^2(f)}{12 \pi^2 f^3}.
+        \end{equation}
+        Assuming a power-law form for the PSD gives an expression analogous to Eq.~\ref{eq:char_strain_plaw}
+        \begin{equation}
+            \label{eq:psd_char_strain}
+            \psd(f) \approx  A^2_\tr{ref} \cdot \scale[\gamma]{\fobs}{\fref} \scale[-3]{f}{\fref},
+        \end{equation}
+        such that the PSD spectral index is related to the characteristic strain spectral index as, $\gamma = 2 \alpha - 3$.
+In terms of PSD, the idealized, circular, GW-only GWB power-law index is then $\gamma = -13/3$.
+Note that the literature uses varying sign conventions for both $\alpha$ and $\gamma$ such that they may typically be positive or negative.
 

### 2023-06-19 docs/holodeck-paper/ms.tex
More methods updates

@@
 For example, for the GWB in the PTA, we are interested in the number of binaries over finite width frequency bins, $\deltalnf = \Delta \ln\!\frstorb$.
-These quantities will be calculated using some numerical integration procedure, for example, in the simplest case: \mbox{$\langle \Delta N \rangle \approx \lr{d^2 N / dz \, d\ln\!f_p} \, \cdot \Delta z \cdot \Delta \ln\!f_p$}.
+These quantities will be calculated using some numerical integration procedure, for example, in the simplest case a Riemann sum: \mbox{$\langle \Delta N \rangle \approx \lr{d^2 N / dz \, d\ln\!f_p} \, \cdot \Delta z \cdot \Delta \ln\!f_p$}.
 For distributions calculated over grids with additional parameters, those will also need to be `integrated' over (e.g.~including $1/\partial M$, $1/\partial q$, etc.
@@
 First, it does not yet account for the intrinsic discreteness of binaries.
-Second, it does not account for random variance from how many binaries happen to exist in a given volume (either physical volume or parameter-space volume).
+Second, it does not account for random variance from how many binaries happen to exist in a given volume---either physical volume or parameter-space volume.
 We address both issues by drawing specific `realizations' of populations.
-Assuming they are uniformly distributed, they will be Poisson distributed such that, $\Delta N \in \poisson(\langle \Delta N \rangle)$.
+Assuming that binaries are uniformly and independently\footnote{These assumptions, while likely accurate, are non trivial.
+For example, the Poisson assumption requires that correlated large-scale structure (e.g.~galaxy clusters) produces negligible excesses over the physical volumes of interest.} distributed, they are described by a Poisson distribution such that, $\Delta N \in \poisson(\langle \Delta N \rangle)$.
 
@@
         This is all that is needed for populations calculated from number-density distributions over grids of parameter space, as in the case of semi-analytic models (SAMs) which are described in the following section.
-An additional simplification can be made, however, in the case of discrete binaries in finite volume simulations, such as cosmological hydrodynamic simulations.
-The most analogous procedure would be to group such binaries into similar bins of parameter space, calculate a number density, and translate into the number of binaries in the observer's light cone using \eqref{eq:pop_discrete_from_continuous}.
-Because a finite range of parameters would still need to be considered before Poisson sampling, this amount to dividing by and then multiplying by, finite widths of parameter space.
-This procedure can be skipped, and instead calculate for each binary ($i$):
+
+        An additional simplification can be made in the case of discrete binaries in finite volume simulations, such as cosmological hydrodynamic simulations.
+The procedure most analogous to that of continuous distributions would be to group discrete binaries into bins of parameter space, calculate a number density, and translate into the number of binaries in the observer's light cone using \eqref{eq:pop_discrete_from_continuous}.
+Because a finite range of parameters would still need to be considered before Poisson sampling, this amounts to dividing by and then multiplying by, finite widths of parameter space.
+Both steps can be skipped, and instead calculate for each binary ($i$):
         \begin{equation}
@@
         In this case, $\thardf$ and $\hsn$ are the hardening time and strain amplitude for each source, and $\langle \Delta N \rangle_i$ represents the expectation number of binaries in the Universe implied by each source in the finite volume.
-        \todo{Is this a biased estimator?  Is there a way to correct for that?  Because we're estimating an expectation value, based on a single Poisson draw from that expectation value.}
-
-        % In the latter case the derivative over redshift in \eqref{eq:pop_discrete_from_continuous} should be replaced with one over all parameters of interest, i.e.~\mbox{$d\ndens / dz \rightarrow d^3\ndens / dM \, dq \, dz$}.
-And each specific function will generally also be dependent on similar parameters, e.g., $\hc = \hc(M,q,z,f)$.
-        % All additional parameters are treated the same in this context, so for brevity we will only explicitly include redshift.
-
-        % The end goal of binary modeling is typically to determine the number of binaries over certain ranges of parameters.
-When calculating gravitational wave signals, for example, the number of binaries across a target frequency interval is desired.
-The discreteness of binary sources can be enforced by converting the differential value of \eqref{eq:pop_discrete_from_continuous} into a number of sources over a small bin, i.e., \mbox{$\Delta N \approx \lr{d^2 N / dz \, d\ln\!f_p} \, \cdot \Delta z \cdot \Delta \ln\!f_p$}.
-
-        % The typical solution is to use some version of \eqref{eq:number_density_to_number_frequency} to convert from a number-density (e.g.~$d\ndens/dz$) into a discrete distribution of individual sources (e.g.~$d^2 N / dz \, d\ln f_r$).
-The number-density is often constructed from either (semi-)analytic models which specify $\ndens(M,q,z,\dots)$ explicitly, or from simulations/observations of sources in a finite volume.
-In the latter case, the number-density will be calculated as $\ndens \equiv N_\trt{finite}/V_\trt{c,finite}$, often over bins in some parameter space (e.g.~$\{M, q, z, \dots\}$).
-\\
-
-        % For a finite population of binaries, the GWB spectrum is only defined over finite bins of frequency.
-The definition of characteristic strain is related to the GW energy across a logarithmic frequency bin, so \eqref{eq:gwb_mc} should really be written as,
-        % \begin{equation}
-        %     \label{eq:gwb_discrete}
-        %     \hc^2(\Delta \ln f) = \sum_\textrm{bins} \sum_n^\infty \frac{\Delta N(f_p,M_a,q_b,z_c,\dots)}{\Delta \ln f_p} \, \hsn^2(f_p,M_a,q_b,z_c,\dots).
-        % \end{equation}
-        % Here we have written out explicitly that the number of binaries, $\Delta N(f_p,M_a,q_b,z_c,\dots)$, is the number sources with frequencies inside $\Delta \ln f_p$, typically over bins of mass, mass-ratio, redshift, etc.
-Discretizing and plugging into Eq.~\ref{eq:number_density_to_number_frequency} (or by examining Eq.~\ref{eq:gwb_sa}), we can then write,
-        % \begin{equation}
-        %     \label{eq:expectation_number_of_binaries}
-        %     \langle \Delta N \rangle = \Delta \ndens \cdot 4\pi c \, \distcom^2 \cdot (1+z) \cdot \thard \cdot \Delta \ln f_p.
-        % \end{equation}
-        % Note that we have again dropped the written dependence on bin values $(M_a, q_b, \dots)$, but they are still implicitly present.
-This expression represents an \textit{expectation value} based on extrapolating the given number-density of binaries.
-It also will not give an integer, discrete number of binaries.
-To address this, we construct discrete realizations by drawing from a Poisson distribution: $\Delta N \in \poisson(\langle \Delta N \rangle)$.
-\\
-
-        % The way \eqref{eq:expectation_number_of_binaries} is handled in practice will still vary somewhat depending on how $\ndens$ is specified.  If a semi-analytic model is being used, then often
-        % $\Delta \ndens = \left[ d^3 \ndens(M,q,z) / dM \, dq \, dz \right] \cdot \Delta M \, \Delta q \, \Delta z$.
-In the case of a finite volume population, the procedure can be simplified.
-An analogous exercise would amount to a binning and un-binning of the sources, i.e.~$\Delta \ndens \big|_\textrm{finite} = \left[ \Delta N_\textrm{finite}(M,q,z) / \Delta M \, \Delta q \, \Delta z \, \Delta V_c \right] \cdot \Delta M \, \Delta q \, \Delta z$, which unnecessarily reduces accuracy.
-Instead, we can imagine shrinking each of our parameter-space bins such that each one captures a single source, in which case we can write,
-        % \begin{equation}
-        %     \label{eq:expectation_number_of_single_binary}
-        %     \langle \Delta N \rangle_\textrm{source} = \frac{4\pi c \, \distcom^2 \, \thard}{V_c} \cdot (1+z) \cdot \Delta \ln f_p.
-        % \end{equation}
-        % In this case, $\thard$ and $\hsn$ in \eqref{eq:gwb_discrete}, are the hardening time and strain amplitude for each source, and $\langle \Delta N \rangle_\textrm{source}$ represents the expectation number of binaries in the Universe implied by each source in the finite volume.
+        \todo{Is this a biased estimator?  I don't think so; the maximum likelihood estimator of a Poisson distribution (i.e. the average over the samples, even if n samples is one) is unbiased.}
 
@@
 
-        Define the distribution function of sources as $F(M,q,a,z) = d^2 n(M,q,a,z) / dM dq$.
-Here $M$ is the total mass of each systems, the mass-ratio is $q\equiv m_2/m_1 \leq 1$, $a$ is the binary separation, and $z$ is the redshift.
-        % Define the distribution function of sources as $F(M,q,f_r,z) = d^2 n(M,q,f_r,z) / dM dq$.
-Here $M$ is the total mass of each systems, the mass-ratio is $q\equiv m_2/m_1 \leq 1$, $f_r$ is the binary orbital frequency in the rest frame, and $z$ is the redshift.
-        We can write the conservation equation for binaries as of function of redshift as,
+        We start by defining the distribution function of sources as a volumetric number density such that,
+        \begin{equation}
+            F(M,q,z,a) = \diffp{n(M,q,z,a)}{{M}{q}}.
+        \end{equation}
+        Here, $M = m_1 + m_2$ is the total mass of each systems, the mass-ratio is $q\equiv m_2/m_1 \leq 1$, $z$ is the redshift, and $a$ is the binary separation.
+Note that we are using the term `binary' to include `pairs' of MBHs at large separations ($a \gg 1 \, \pc$), in addition to true gravitationally-bound systems ($a \lesssim 10s \pc$).
+We can write the conservation equation for binaries over redshift as,
         \begin{equation}
@@
                 \diffp{}{q} \lrs{F \diffp{q}{z}} +
-                \diffp{}{a} \lrs{F \diffp{a}{z}} = S_{\!F}(M, q, a, z).
+                \diffp{}{a} \lrs{F \diffp{a}{z}} = S_{\!F}(M,q,z,a).
         \end{equation}
@@
 
-        We consider the standard semi-analytic model (SAM) formalism of MBH binary populations \citep[e.g.~][]{Sesana+2008, Chen+2019}.
-In this style of calculation, $F$ is determined in a region of parameter space that can be observed/estimated, and this is evolved to find the distribution in a different region of parameter space that is of interest.
+        In the standard implementations of semi-analytic models (SAMs) for MBH binaries \citep{Rajagopal+Romani-1995, Jaffe+Backer-2003, Sesana+2008, Chen+2019}, $F$ is determined in a region of parameter space that can be observed/estimated, and this is evolved to find the distribution in a different region of parameter space that is of interest.
 In practice, the observed parameter space is galaxies and galaxy mergers, and the parameter space of interest is closely separated MBH binaries that could be GW detectable.
-Thus we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts), and we set $S_{\!F} = 0$.
+We assume that in between these two parts of parameter space there is no additional creation or destruction of binaries, and we set $S_{\!F} = 0$.
 
-        We will express the distribution function as a product of a mass function, and a pair fraction:
+        From $\eqref{eq:conservation}$, we use the chain rule to mix time and redshift evolution (where $z = z(t)$), and assume that the mass-change of binaries is negligible, i.e.~$\diffp{m}{t} = 0$ and $\diffp{q}{t} = 0$, giving:
+        \begin{equation}
+            \label{eq:conservation_fixed_mass}
+            \diffp{F}{z} = - \diffp{t}{z} \diffp{}{a} \lrs{F \diffp{a}{t}}.
+        \end{equation}
+        At this point the binary population is assumed to be changing only in separation and redshift, which are related by \mbox{$\partial a / \partial z = (\partial a / \partial t) (\partial t / \partial z)$}.
+
+        We consider two different SAM parameterizations, one in terms of a galaxy merger rate (GMR, $R_g$), and another in terms of a galaxy pair fraction (GPF, $P_g$) and galaxy merger time (GMT, $T_g$), such that $R_g \approx F_g/T_g$.
+In the former case, we consider the GMR as a flux: the rate at which galaxy pairs reach a given part of parameter space, $R_g = R_g(m_1,q,z,a)$.
+Because we are assuming masses to be fixed, this amounts to pairs of a given mass, mass-ratio, and redshift reaching a certain separation.
+We have written the GMR as a function of the primary mass $m_1$ because this is typically more analogous to parameterizations derived from either observations or simulations, but any pair of $M,q$ can be mapped to a corresponding $m_1,q$.
+We additionally introduce the galaxy mass function as the number-density of individual galaxies,
+        \begin{equation}
+            \Phi(m_1,z) \equiv \diffp{n(m_1,z)}{{m_1}},
+        \end{equation}
+        which allows us to rewrite \eqref{eq:conservation_fixed_mass} as,
+        \begin{equation}
+            \label{eq:sam_gmr}
+            \diffp{F}{z} = \diffp{n}{{M}{q}{z}} = - \diffp{t}{z} \, \Phi(m_1,z) \, R(m_1,q,z,a).
+        \end{equation}
+
+        Galaxy merger rates are particularly convenient when using finite-volume cosmological simulations \citep[e.g.][]{Lacey+Cole-1993}.
+In this case, an ensemble of galaxies (or dark matter halos) can be tracked, and the merger rate calculated from the number of mergers divided by the time span.
+This amounts to marginalizing over the separations, i.e.~~\mbox{$R_g = \int_{a_1}^{a_2} R_g(a) \, da$}.
+From observational data, specific galaxies cannot be tracked over time, and instead the occurrence rate of galaxy pairs is much more natural.
+This motivates expressing the distribution function of pairs as a product of the number density of singles (i.e.~mass function) with a pair fraction,
         \begin{equation}
             \label{eq:dist_func}
-            F(M,q,z) = \frac{\Phi(M, z)}{M \ln\!10} \cdot P(M,q,z),
+            F(M,q,z,a) = \Phi(m_1, z) \cdot P(m_1,q,z,a) \big|_{M=m_1(1+q)},
         \end{equation}
-        where the mass function, $\Phi(M, q, z) \equiv \diffp{n_g}{{\logten{M}}}$, is calculated based on the number density of galaxies ($n_g$).
-We assume that there is a one-to-one mapping from galaxy mass to MBH mass, such that the galaxy mass-function can still be used to uniquely define the mass distribution of MBHs.
-Typically the MBH--galaxy relation is given in terms of an `$M_\trt{BH}$--$M_\trt{bulge}$' relation \citep[e.g.][]{Kormendy+Ho-2013}, which is an observationally derived relation between the mass of MBHs and the mass of the stellar bulge component of their host galaxy.
-        In \citet{Chen+2019}, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$P = \int_{a_0}^{a_1} P_a \, da$.
+        \begin{equation}
+            \label{eq:pair_frac}
+            P(m_1,q,z) \equiv \diffp{}{q} \lrs{\frac{N_\tr{galaxy-pairs}(m_1,q,z)}{N_\tr{galaxies}(m_1,z)}}.
+        \end{equation}
+        It's important to note that the observational identification of pairs is both challenging and subtle \needcite{}.
+In a given survey, pairs can only be robustly associated with each other within some maximum separation, and they will only be distinguishable above some minimum separation.
+The GPF is often marginalized over this range of separations \citep[e.g.][]{Chen+2019}, i.e.~\mbox{$P_g = \int_{a_1}^{a_2} P(a) \, da$}.
+Due to these constraints, the pair fraction is typically not meaningful by itself and needs to be compared to a characteristic timescale over which the pairs would be observable.
 
-        From $\eqref{eq:conservation}$, we use the chain rule to mix time and redshift evolution, and assume that the mass-change of binaries is negligible, i.e.~$\diffp{m}{t} = 0$ and $\diffp{q}{t} = 0$, giving:
+        The number-density of binaries is conserved, which allows us to take finite steps in separation and time/redshift without approximation: $a\rightarrow a'$ and $z\rightarrow z'$.
+The time it takes for a binary to go between separations $a_1 \rightarrow a_2$ is
+        % \mbox{$T \equiv \int_{a_1}^{a_2} \frac{da}{\partial a / \partial t}$}.
+        \mbox{$T \equiv \int_{a_1}^{a_2} da / \lr{\partial a / \partial t}$}.
+        This leads to a redshift at the later time of $z' = z'(t + T)$.  This allows us to approximate the term from \eqref{eq:conservation_fixed_mass} as,
         \begin{equation}
-            \diffp{F}{z} = - \diffp{t}{z} \diffp{}{a} \lrs{F \diffp{a}{t}}.
+            \label{eq:}
+            \diffp{}{a} \lrs{F(M,q,z,a) \diffp{a}{t}} \approx \frac{F(M,q,z,a)}{T_g(m_1,q,z)}.
         \end{equation}
+        The approximation is tied to the marginalization over binary separation in the GMT term.
+Because the number density of binaries $n$ is not explicitly dependent on the separation, we can still write exactly that
+        \begin{equation}
+            \label{eq:sam_gpf_fmt}
+            \diffp{n}{{M}{q}{z'}} = - \diffp{t}{z} \frac{\Phi_g(m_1,z) \, P_g(m_1,q,z)}{T_g(m_1,q,z)}.
+        \end{equation}
+        Putting this all together, we can write our SAM number of binaries as either,
+        \begin{equation}
+            \begin{split}
+                \label{eq:sam_final_gmr}
+                \diffp{N}{{M}{q}{z'}{\ln \! f_r}} & = \Phi_g(m_1,z) \, R_g(m_1,q,z) \, \thardf(M,q,z',f) \, \diffp{{V_c}}{{z'}}, \\
+                & = \Phi_g(m_1,z) \, \frac{P_g(m_1,q,z)}{T_g(m_1,q,z)} \, \thardf(M,q,z',f) \, \diffp{V_c}{{z'}}.
+            \end{split}
+        \end{equation}
+        \todo{Should $R_g$ use $z$ or $z'$ (redshift-prime)?}
 
-        The binary population is assumed to be changing only in separation and redshift, which are related by $\partial a / \partial z = (\partial a / \partial t) (\partial t / \partial z)$.
-Because the overall number-density is conserved, we can take a finite step in separation and time/redshift, $a\rightarrow a'$ and $z\rightarrow z'$.
-Here the time it takes for a binary to go from $a \rightarrow a'$ is $T(M,q,a,z|a')$, which leads to a redshift at the later time of $z' = z'(t + T)$.
-So far, we have left the binary separation $a$ as implicit in the expression for F.
-To obtain the standard expression \citep[e.g.][~Eq.~5]{Chen+2019}, we make the approximation that,
-        \begin{equation}
-            \diffp{}{a} \lrs{F(M,q,z) \diffp{a}{t}} \approx \frac{F}{T(M,q,a,z|a')}.
-        \end{equation}
-        Thus giving,
-        \begin{equation}
-            \label{eq:cont_eq_result}
-            \diffp{F(M,q,a',z')}{{z'}} = \diffp{n}{{M}{q}{z'}} = - \diffp{t}{z} \frac{\Phi(M,z) \, P(M,q,z)}{T(M,q,a,z|a')}.
-        \end{equation}
-
-        Combining \eqref{eq:cont_eq_result} with \eqref{eq:num_num_dens}, we can finally write,
-        \begin{equation}
-            \label{eq:sam_final}
-            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = \frac{\Phi(M,z) \, P(M,q,z)}{T(M,q,a,z|a')} \, \thardf \, \diffp{V_c}{z}.
-        \end{equation}
-
-        From the SAM formalism, the number of binaries in each bin is calculated by `integrating' each finite element of \eqref{eq:sam_final}, i.e.,
-        \begin{equation}
-            \diffp{N_{ijk}(f_r)}{{\ln\!f_r}} \equiv \diffp{N(M_i, q_j, z_k, {f_r})}{{M}{q}{z}{\ln\!f_r}} \, \Delta M \,\Delta q \, \Delta z.
-        \end{equation}
-        The characteristic strain of the GWB can then be calculated
-
-    \subsection{Finite Volume Populations: the Illustris simulations}
-
-        \todo{Describe illustris models.}
+    % \subsection{Finite Volume Populations: the Illustris simulations}
+    %     \todo{Describe illustris models.}
 

### 2023-06-19 docs/holodeck-paper/ms.tex

@@
 
-\newtotcounter{citnum} % From the package documentation
-\def\oldbibitem{} \let\oldbibitem=\bibitem
-\def\bibitem{\stepcounter{citnum}\oldbibitem}
+% \newtotcounter{citnum} % From the package documentation
+% \def\oldbibitem{} \let\oldbibitem=\bibitem
+% \def\bibitem{\stepcounter{citnum}\oldbibitem}
 
-\newcommand{\orcidauthorA}{0000-0002-6625-6450} % For author A
+% \newcommand{\orcidauthorA}{0000-0002-6625-6450} % For author A
 
@@
 The goal of this package is to provide a cohesive and comprehensive framework for:
+    % Massive black holes (MBHs) are known to occupy the centers of most if not all massive galaxies \needcite{}.
+    Binaries of MBHs form following the merger of massive galaxies \needcite{}.
+These MBH Binaries (MBHBs) can then produce the loudest gravitational waves (GWs) in the Universe.
+MBHBs with total masses $M \gtrsim 10^8 \,\msol$ are detectable by experiments called pulsar timing arrays (PTAs) \needcite{PTAs} in the nanohertz regime \needcite{}, while more modest systems with $M \lesssim 10^7 \, \msol$ will be detectable by LISA \needcite{LISA} at millihertz \needcite{}.
+While accreting, MBHs are visible across the electromagnetic (EM) spectrum as active galactic nuclei (AGN) \needcite{}, suggesting that MBHBs could be promising multimessenger sources \needcite{}.
+A large number of studies have modeled cosmological populations of MBHBs, and made predictions for their GW signatures and EM counterparts, however no comprehensive framework has been released.
+In this paper, we introduce the MBHB population synthesis code \holodeck{}, developed within the astrophysics working group of the NANOGrav\footnote{North American nanohertz observatory for gravitational waves \citep{}: \url{https://www.nanograv.org}.}.
+The goal of this package is to provide a cohesive and comprehensive framework for:
     \begin{enumerate}
         \item
-        modeling the physics of MBHB formation and evolution while accurately capturing the true uncertainties in this modeling;
+        modeling the physics of MBHB formation and evolution, with parameterizations for uncertain physical processes;
         \item
-        rapidly generating large simulated populations of MBHBs;
+        rapidly generating large simulated populations of MBHBs; and,
         \item
-        calculating the resulting GW signals from these populations.
+        calculating the resulting GW signals and EM signatures from these populations.
     \end{enumerate}
@@
             \diffp{}{a} \lrs{F(M,q,z,a) \diffp{a}{t}} \approx \frac{F(M,q,z,a)}{T_g(m_1,q,z)}.
-        \end{equation}
+        \end{equation}`'
         The approximation is tied to the marginalization over binary separation in the GMT term.

### 2023-06-19 holodeck/__init__.py
Move SAMs stuff to submodule/directory, separate components into dedicated file; add galaxy merger rate formalism

@@ -93,7 +93,7 @@ from . import param_spaces    # noqa
 from . import plot            # noqa
 from . import population      # noqa
 from . import relations       # noqa
-from . import sam             # noqa
+from . import sams            # noqa
 from . import utils           # noqa
 
 # ---- Handle version

### 2023-06-19 holodeck/sams/__init__.py

@@ -1,0 +1,7 @@
+"""Holodeck - Semi-Analytic Models
+"""
+
+from . import sam
+from . import comps
+from sam import Semi_Analytic_Model
+from comps import (GSMF_Schechter, GPF_Power_Law, GMT_Power_Law, GMR_Illustris)

### 2023-06-19 holodeck/sams/sam.py

@@ -36,7 +36,6 @@ References
 
 """
 
-import abc
 from datetime import datetime
 
 import numpy as np
@@ -46,358 +45,20 @@ import scipy.interpolate  # noqa
 import kalepy as kale
 
 import holodeck as holo
-from holodeck import cosmo, utils
-from holodeck.constants import GYR, SPLC, MSOL, MPC
+from holodeck import cosmo, utils, log
+from holodeck.constants import SPLC, MSOL, MPC
 from holodeck import relations, gravwaves, single_sources
 import holodeck.sam_cython
+from holodeck.sams.comps import (
+    _Galaxy_Pair_Fraction, _Galaxy_Stellar_Mass_Function, _Galaxy_Merger_Time, _Galaxy_Merger_Rate,
+    GSMF_Schechter, GPF_Power_Law, GMT_Power_Law, GMR_Illustris
+)
 
 REDZ_SAMPLE_VOLUME = True    #: get redshifts by sampling uniformly in 3D spatial volume, and converting
 
 GSMF_USES_MTOT = False       #: the mass used in the GSMF is interpretted as M=m1+m2, otherwise use primary m1
 GPF_USES_MTOT = False        #: the mass used in the GPF  is interpretted as M=m1+m2, otherwise use primary m1
 GMT_USES_MTOT = False        #: the mass used in the GMT  is interpretted as M=m1+m2, otherwise use primary m1
-
-
-
-# ==============================
-# ====    SAM Components    ====
-# ==============================
-
-
-# ----    Galaxy Stellar-Mass Function    ----
-
-
-class _Galaxy_Stellar_Mass_Function(abc.ABC):
-    """Galaxy Stellar-Mass Function base-class.  Used to calculate number-density of galaxies.
-
-
-    """
-
-    @abc.abstractmethod
-    def __init__(self, *args, **kwargs):
-        return
-
-    @abc.abstractmethod
-    def __call__(self, mstar, redz):
-        """Return the number-density of galaxies at a given stellar mass, per log10 interval of stellar-mass.
-
-        i.e. Phi = dn / dlog10(M)
-
-        Parameters
-        ----------
-        mstar : scalar or ndarray
-            Galaxy stellar-mass in units of [grams]
-        redz : scalar or ndarray
-            Redshift.
-
-        Returns
-        -------
-        rv : scalar or ndarray
-            Number-density of galaxies in units of [Mpc^-3]
-
-        """
-        return
-
[…]

### 2023-06-19 holodeck/sams/comps.py
Move SAMs stuff to submodule/directory, separate components into dedicated file; add galaxy merger rate formalism

@@ -1,0 +1,557 @@
+"""Semi-Analytic Model - Components
+"""
+
+import abc
+# from datetime import datetime
+
+import numpy as np
+
+import holodeck as holo
+from holodeck import cosmo, utils
+from holodeck.constants import GYR, MSOL
+
+
+# ----    Galaxy Stellar-Mass Function    ----
+
+
+class _Galaxy_Stellar_Mass_Function(abc.ABC):
+    """Galaxy Stellar-Mass Function base-class.  Used to calculate number-density of galaxies.
+
+
+    """
+
+    @abc.abstractmethod
+    def __init__(self, *args, **kwargs):
+        return
+
+    @abc.abstractmethod
+    def __call__(self, mstar, redz):
+        """Return the number-density of galaxies at a given stellar mass, per log10 interval of stellar-mass.
+
+        i.e. Phi = dn / dlog10(M)
+
+        Parameters
+        ----------
+        mstar : scalar or ndarray
+            Galaxy stellar-mass in units of [grams]
+        redz : scalar or ndarray
+            Redshift.
+
+        Returns
+        -------
+        rv : scalar or ndarray
+            Number-density of galaxies in units of [Mpc^-3]
+
+        """
+        return
+
+    def mbh_mass_func(self, mbh, redz, mmbulge, scatter=None):
+        """Convert from the GSMF to a MBH mass function (number density), using a given Mbh-Mbulge relation.
+
+        Parameters
+        ----------
+        mbh : array_like
+            Blackhole masses at which to evaluate the mass function.
+        redz : array_like
+            Redshift(s) at which to evaluate the mass function.
+        mmbulge : `relations._MMBulge_Relation` subclass instance
+            Scaling relation between galaxy and MBH masses.
+        scatter : None, bool, or float
+            Introduce scatter in masses.
+            * `None` or `True` : use the value from `mmbulge._scatter_dex`
+            * `False` : do not introduce scatter
+            * float : introduce scatter with this amplitude (in dex)
+
+        Returns
+        -------
+        ndens : array_like
+            Number density of MBHs, in units of [Mpc^-3]
+
+        """
+        if scatter in [None, True]:
+            scatter = mmbulge._scatter_dex
+
[…]

### 2023-06-19 holodeck/__init__.py
Change default log level to WARNING

@@ -72,7 +72,7 @@ for cp in _check_paths:
 # ---- Load logger
 
 from . import logger   # noqa
-log = logger.get_logger(__name__, logging.INFO)       #: global root logger from `holodeck.logger`
+log = logger.get_logger(__name__, logging.WARNING)       #: global root logger from `holodeck.logger`
 
 # ---- Load cosmology instance
 

### 2023-06-19 holodeck/utils.py
Add new 'pta_freqs' method and add deprecation warning for nyquist_freqs (and edges)

@@ -73,6 +73,30 @@ class _Modifier(abc.ABC):
 # =================================================================================================
 
 
+def deprecated_warn(msg, exc_info=True):
+    """Decorator for functions that will be deprecated, add warning, but still execute function.
+    """
+
+    def decorator(func):
+        nonlocal msg
+
+        @functools.wraps(func)
+        def wrapper(*args, **kwargs):
+            nonlocal msg
+            old_name = func.__name__
+            _frame = inspect.currentframe().f_back
+            file_name = inspect.getfile(_frame.f_code)
+            fline = _frame.f_lineno
+            msg = f"{file_name}({fline}):{old_name} is deprecated!" + " | " + msg
+            warnings.warn_explicit(msg, category=DeprecationWarning, filename=file_name, lineno=fline)
+            log.warning(f"DEPRECATION: {msg}", exc_info=exc_info)
+            return func(*args, **kwargs)
+
+        return wrapper
+
+    return decorator
+
+
 def deprecated_pass(new_func, msg="", exc_info=True):
     """Decorator for functions that have been deprecated, warn and pass arguments to new function.
     """
@@ -779,83 +803,17 @@ def ndinterp(xx, xvals, yvals, xlog=False, ylog=False):
     return ynew
 
 
-def nyquist_freqs(
-    dur: float = 15.0*YR, cad: float = 0.1*YR, trim: Optional[Tuple[float, float]] = None
-) -> np.ndarray:
-    """Calculate Nyquist frequencies for the given timing parameters.
-
-    Parameters
-    ----------
-    dur : float,
-        Duration of observations
-    cad : float,
-        Cadence of observations
-    trim : (2,) or None,
-        Specification of minimum and maximum frequencies outside of which to remove values.
-        `None` can be used in place of either boundary, e.g. [0.1, None] would mean removing
-        frequencies below `0.1` (and not trimming values above a certain limit).
-
-    Returns
-    -------
-    freqs : ndarray,
-        Nyquist frequencies
-
-    """
+def pta_freqs(dur=16.03*YR, num=40, cad=None):
     fmin = 1.0 / dur
-    fmax = 1.0 / cad * 0.5
-    # df = fmin / sample
-    df = fmin
-    freqs = np.arange(fmin, fmax + df/10.0, df)
-    if trim is not None:
[…]

### 2023-06-19 holodeck/sams/__init__.py
Update SAM notebook to include GMR demo

@@ -3,5 +3,5 @@
 
 from . import sam
 from . import comps
-from sam import Semi_Analytic_Model
-from comps import (GSMF_Schechter, GPF_Power_Law, GMT_Power_Law, GMR_Illustris)
+from .sam import Semi_Analytic_Model
+from .comps import (GSMF_Schechter, GPF_Power_Law, GMT_Power_Law, GMR_Illustris)

### 2023-06-19 holodeck/detstats.py
Move sam_cython.pyx to sams/cyutils.pyx

@@ -12,28 +12,29 @@ from sympy import nsolve, Symbol
 import holodeck as holo
 from holodeck import utils, cosmo, log
 from holodeck.constants import MPC, YR
+from holodeck.sams import cyutils as sam_cyutils
 
 import hasasia.sensitivity as hsen
 import hasasia.sim as hsim
 import hasasia as has
 
 GAMMA_RHO_GRID_PATH = '/Users/emigardiner/GWs/holodeck/output/rho_gamma_grids'
-HC_REF15_10YR = 11.2*10**-15 
+HC_REF15_10YR = 11.2*10**-15
 
 
 ###################### Overlap Reduction Function ######################
 
 def _gammaij_from_thetaij(theta_ij):
     """ Calcualte gamma_ij for two pulsars of relative angle theta_ij.
-    
+
     Parameters
     ----------
-    theta_ij : scalar 
+    theta_ij : scalar
         Relative angular position between the ith and jth pulsars.
 
     Returns
     -------
-    gamma_ij : scalar 
+    gamma_ij : scalar
         [1 - cos(theta_ij)]/2
 
     """
@@ -48,7 +49,7 @@ def _dirac_delta(i,j):
     j : int
 
     Returns
-    ------- 
+    -------
     dirac_ij : int
         Dirac delta function of i and j
 
@@ -58,33 +59,33 @@ def _dirac_delta(i,j):
 
 def _relative_angle(theta_i, phi_i, theta_j, phi_j):
     """ Calcualte relative angle between two pulsars i and j.
-    
+
     Parameters
     ----------
-    theta_i : scalar 
+    theta_i : scalar
         Polar (latitudinal) angular position in the sky of the ith pulsar.
     phi_i : scalar
         Azimuthal (longitudinal) angular position in the sky of the ith pulsar.
-    theta_j : scalar 
+    theta_j : scalar
         Polar (latitudinal) angular position in the sky of the jth pulsar.
     phi_j : scalara
         Azimuthal (longitudinal) angular position in the sky of the jth pulsar.
 
     Returns
     -------
-    theta_ij : scalar 
+    theta_ij : scalar
         Relative angular position between the ith and jth pulsar.
 
     """
-    
[…]

### 2023-06-19 holodeck/sams/cyutils.pyx

@@ -425,11 +425,16 @@ def dynamic_binary_number_at_fobs(fobs_orb, sam, hard, cosmo):
     # ---- Fixed_Time_2pwl_SAM
 
     if isinstance(hard, holo.hardening.Fixed_Time_2PL_SAM):
+        gmt_time = sam._gmt_time
+        # if `sam` is using galaxy merger rate (GMR), then `gmt_time` will be `None`
+        if gmt_time is None:
+            sam._log.info("`gmt_time` not calculated in SAM.  Setting to zeros.")
+            gmt_time = np.zeros(sam.shape)
 
         _dynamic_binary_number_at_fobs_2pwl(
             fobs_orb, hard._sepa_init, hard._num_steps,
             hard._norm, hard._rchar, hard._gamma_inner, hard._gamma_outer,
-            dens, sam.mtot, sam.mrat, sam.redz, sam._gmt_time,
+            dens, sam.mtot, sam.mrat, sam.redz, gmt_time,
             cosmo._grid_z, cosmo._grid_dcom, cosmo._grid_age,
             # output:
             redz_final, diff_num
@@ -438,9 +443,15 @@ def dynamic_binary_number_at_fobs(fobs_orb, sam, hard, cosmo):
     # ---- Hard_GW
 
     elif isinstance(hard, holo.hardening.Hard_GW) or issubclass(hard, holo.hardening.Hard_GW):
+        redz_prime = sam._redz_prime
+        # if `sam` is using galaxy merger rate (GMR), then `redz_prime` will be `None`
+        if redz_prime is None:
+            sam._log.info("`redz_prime` not calculated in SAM.  Setting to `redz` (initial) values.")
+            redz_prime = sam.redz[np.newaxis, np.newaxis, :] * np.ones(sam.shape)
+
         _dynamic_binary_number_at_fobs_gw(
             fobs_orb,
-            dens, sam.mtot, sam.mrat, sam.redz, sam._redz_prime,
+            dens, sam.mtot, sam.mrat, sam.redz, redz_prime,
             cosmo._grid_z, cosmo._grid_dcom,
             # output:
             redz_final, diff_num
@@ -782,7 +793,7 @@ def gamma_of_rho_interp(rho, rsort, rho_interp_grid, gamma_interp_grid):
 @cython.nonecheck(False)
 @cython.cdivision(True)
 cdef int _gamma_of_rho_interp(
[…]

### 2023-06-19 holodeck/librarian.py
Move sam_cython.pyx to sams/cyutils.pyx

@@ -22,7 +22,7 @@ from scipy.stats import qmc
 
 import holodeck as holo
 import holodeck.single_sources
-import holodeck.sam_cython
+from holodeck.sams import cyutils as sam_cyutils
 from holodeck import utils, cosmo
 from holodeck.constants import YR
 
@@ -574,11 +574,11 @@ def run_sam_at_pspace_num(args, space, pnum):
             log.exception(err)
             raise RuntimeError(err)
 
-        redz_final, diff_num = holo.sam_cython.dynamic_binary_number_at_fobs(
+        redz_final, diff_num = sam_cyutils.dynamic_binary_number_at_fobs(
             fobs_orb_cents, sam, hard, cosmo
         )
         edges = [sam.mtot, sam.mrat, sam.redz, fobs_orb_edges]
-        number = holo.sam_cython.integrate_differential_number_3dx1d(edges, diff_num)
+        number = sam_cyutils.integrate_differential_number_3dx1d(edges, diff_num)
 
         log.debug(f"{utils.stats(number)=}")
 
@@ -677,12 +677,12 @@ def run_model(sam, hard, nreals, nfreqs, nloudest=5,
 
     data = dict(fobs_cents=fobs_cents, fobs_edges=fobs_edges)
 
-    redz_final, diff_num = holo.sam_cython.dynamic_binary_number_at_fobs(
+    redz_final, diff_num = sam_cyutils.dynamic_binary_number_at_fobs(
         fobs_orb_cents, sam, hard, cosmo
     )
     use_redz = redz_final
     edges = [sam.mtot, sam.mrat, sam.redz, fobs_orb_edges]
-    number = holo.sam_cython.integrate_differential_number_3dx1d(edges, diff_num)
+    number = sam_cyutils.integrate_differential_number_3dx1d(edges, diff_num)
     if details_flag:
         data['static_binary_density'] = sam.static_binary_density
         data['number'] = number

### 2023-06-19 holodeck/population.py

@@ -614,7 +614,7 @@ class PM_Mass_Reset(_Population_Modifier):
         """
         # if `mhost` is a class (not an instance), then instantiate it; make sure its a subclass
         # of `_Host_Relation`
-        mhost = utils._get_subclass_instance(mhost, None, holo.relations._Host_Relation)
+        mhost = utils.get_subclass_instance(mhost, None, holo.relations._Host_Relation)
         # store attributes
         self.mhost = mhost         #: Scaling relationship between host and MBH (`holo.relations._Host_Relation`)
         self._scatter = scatter    #: Bool determining whether resampled masses should include statistical scatter

### 2023-06-19 holodeck/sams/sam.py

@@ -47,8 +47,8 @@ import kalepy as kale
 import holodeck as holo
 from holodeck import cosmo, utils, log
 from holodeck.constants import SPLC, MSOL, MPC
-from holodeck import relations, gravwaves, single_sources
-import holodeck.sam_cython
+from holodeck import relations, single_sources
+from . import cyutils as sam_cyutils
 from holodeck.sams.comps import (
     _Galaxy_Pair_Fraction, _Galaxy_Stellar_Mass_Function, _Galaxy_Merger_Time, _Galaxy_Merger_Rate,
     GSMF_Schechter, GPF_Power_Law, GMT_Power_Law, GMR_Illustris
@@ -113,18 +113,18 @@ class Semi_Analytic_Model:
 
         # ---- Process SAM components
 
-        gsmf = utils._get_subclass_instance(gsmf, None, _Galaxy_Stellar_Mass_Function)
-        mmbulge = utils._get_subclass_instance(mmbulge, None, relations._MMBulge_Relation)
-        # gpf = utils._get_subclass_instance(gpf, None, _Galaxy_Pair_Fraction)
-        # gmt = utils._get_subclass_instance(gmt, None, _Galaxy_Merger_Time)
+        gsmf = utils.get_subclass_instance(gsmf, None, _Galaxy_Stellar_Mass_Function)
+        mmbulge = utils.get_subclass_instance(mmbulge, None, relations._MMBulge_Relation)
+        # gpf = utils.get_subclass_instance(gpf, None, _Galaxy_Pair_Fraction)
+        # gmt = utils.get_subclass_instance(gmt, None, _Galaxy_Merger_Time)
         if gmr is None:
             # if GMR is None, then we need both GMT and GPF
-            gmt = utils._get_subclass_instance(gmt, GMT_Power_Law, _Galaxy_Merger_Time)
-            gpf = utils._get_subclass_instance(gpf, GPF_Power_Law, _Galaxy_Pair_Fraction)
+            gmt = utils.get_subclass_instance(gmt, GMT_Power_Law, _Galaxy_Merger_Time)
+            gpf = utils.get_subclass_instance(gpf, GPF_Power_Law, _Galaxy_Pair_Fraction)
         else:
-            gmr = utils._get_subclass_instance(gmr, GMR_Illustris, _Galaxy_Merger_Rate)
+            gmr = utils.get_subclass_instance(gmr, GMR_Illustris, _Galaxy_Merger_Rate)
             # if GMR is given, GMT can still be used - for calculating GMT stalling
[…]

### 2023-06-19 holodeck/utils.py

@@ -294,7 +294,7 @@ def path_name_ending(path, ending):
     return fname
 
 
-def _get_subclass_instance(value, default, superclass):
+def get_subclass_instance(value, default, superclass, allow_none=False):
     """Convert the given `value` into a subclass instance.
 
     `None` ==> instance from `default` class
@@ -326,6 +326,10 @@ def _get_subclass_instance(value, default, superclass):
     # Set `value` to a default, if needed and it is given
     if (value is None) and (default is not None):
         value = default
+
+    # if `value` is not set, and there is no default, and `None` is allowed... just return None
+    if (value is None) and allow_none:
+        return value
 
     # If `value` is a class (constructor), then construct an instance from it
     if inspect.isclass(value):
@@ -2360,5 +2364,52 @@ def nyquist_freqs_edges(
             freqs_edges = freqs_edges[freqs_edges < trim[1]]
 
     return freqs_edges
+
+
+@deprecated_pass(get_subclass_instance)
+def _get_subclass_instance(value, default, superclass):
+    """Convert the given `value` into a subclass instance.
+
+    `None` ==> instance from `default` class
+    Class ==> instance from that class
+    instance ==> check that this is an instance of a subclass of `superclass`, error if not
+
+    Parameters
+    ----------
+    value : object,
+        Object to convert into a class instance.
+    default : class,
+        Default class constructor to use if `value` is None.
+    superclass : class,
+        Super/parent class to compare against the class instance from `value` or `default`.
+        If the class instance is not a subclass of `superclass`, a ValueError is raised.
+
+    Returns
+    -------
+    value : object,
+        Class instance that is a subclass of `superclass`.
+
+    Raises
+    ------
+    ValueError : if the class instance is not a subclass of `superclass`.
+
+    """
+    import inspect
+
+    # Set `value` to a default, if needed and it is given
+    if (value is None) and (default is not None):
+        value = default
+
+    # If `value` is a class (constructor), then construct an instance from it
+    if inspect.isclass(value):
+        value = value()
+
+    # Raise an error if `value` is not a subclass of `superclass`
+    if not isinstance(value, superclass):
+        err = f"argument ({value}) must be an instance or subclass of `{superclass}`!"
+        log.error(err)
+        raise ValueError(err)
+
+    return value
 
 

### 2023-06-19 holodeck/sams/comps.py
Add basic functionality tests for sams/comps.py

@@ -16,8 +16,6 @@ from holodeck.constants import GYR, MSOL
 
 class _Galaxy_Stellar_Mass_Function(abc.ABC):
     """Galaxy Stellar-Mass Function base-class.  Used to calculate number-density of galaxies.
-
-
     """
 
     @abc.abstractmethod
@@ -186,17 +184,6 @@ class _Galaxy_Merger_Rate(abc.ABC):
 
         """
         return
-
-    '''
-    def zprime(self, mass, mrat, redz, **kwargs):
-        """Return the redshift after merger (i.e. input `redz` delayed by merger time).
-        """
-        # tau0 = self(mass, mrat, redz, **kwargs)  # sec
-        # # Find the redshift of  t(z) + tau
-        # redz_prime = utils.redz_after(tau0, redz=redz)
-        # return redz_prime, tau0
-        return np.copy(redz), np.zeros_like(redz)
-    '''
 
 
 class GMR_Power_Law(_Galaxy_Merger_Rate):

### 2023-06-19 holodeck/sams/tests/test_comps.py

@@ -1,0 +1,135 @@
+"""
+"""
+
+import numpy as np
+import pytest
+
+import holodeck as holo
+from holodeck.constants import MSOL
+
+from holodeck.sams import comps
+
+
+def _test_abc_parent(parent_class):
+
+    # Parent = comps._Galaxy_Pair_Fraction
+
+    # Make sure the abc cannot be instantiated directly
+    with pytest.raises(TypeError):
+        parent_class()
+
+    # make sure a subclass cannot be instatiated directly
+    class Child_Fail(parent_class):
+        pass
+
+    with pytest.raises(TypeError):
+        Child_Fail()
+
+    # make sure that a subclass which overrides `__init__` and `__call__` CAN be instantiated
+    class Child_Good(parent_class):
+
+        def __init__(self):
+            pass
+
+        def __call__(self):
+            pass
+
+    Child_Good()
+
+    return
+
+
+# ==============================================================================
+# ====    Galaxy Stellar-Mass Function    ====
+# ==============================================================================
+
+
+def test_gsmf_base():
+    _test_abc_parent(comps._Galaxy_Stellar_Mass_Function)
+    return
+
+
+def test_GSMF_Schechter():
+    gsmf = comps.GSMF_Schechter()
+    SIZE = 100
+    mstar = MSOL * (10.0**np.random.uniform(10, 15, SIZE))
+    redz = np.random.uniform(0.0, 10.0, SIZE)
+
+    vals = gsmf(mstar, redz)   # noqa
+    return
+
+
+# ==============================================================================
+# ====    Galaxy Pair Fraction    ====
+# ==============================================================================
+
+
+def test_gpf_base():
+    _test_abc_parent(comps._Galaxy_Pair_Fraction)
+    return
+
+
+def test_GPF_Power_Law():
+    gpf = comps.GPF_Power_Law()
+    SIZE = 100
+    mstar = MSOL * (10.0**np.random.uniform(10, 15, SIZE))
+    mrat = 10.0 ** np.random.uniform(-4, 0.0, SIZE)
+    redz = np.random.uniform(0.0, 10.0, SIZE)
+
+    vals = gpf(mstar, mrat, redz)   # noqa
+    return
+
+
+# ==============================================================================
+# ====    Galaxy Merger Time    ====
[…]

### 2023-06-20 holodeck/detstats.py
Added ev back to library calculation, because it requires gamma_ssi which is too large to store.

@@ -1581,6 +1581,9 @@ def detect_lib(hdf_name, output_dir, npsrs, sigma, nskies, thresh=DEF_THRESH,
         Fraction of realizations with a single source detection, for each sample.
     df_bg : (N,) 1Darray
         Fraction of realizations with a background detection, for each sample.
+    ev_ss : (N,) 1Darray
+        Expectation number of single source detections, averaged across realizations,
+        for each sample.
 
     """
 
@@ -1623,7 +1626,7 @@ def detect_lib(hdf_name, output_dir, npsrs, sigma, nskies, thresh=DEF_THRESH,
     snr_bg = np.zeros((nsamp, nfreqs, nreals))
     df_ss = np.zeros(nsamp)
     df_bg = np.zeros(nsamp)
-    # ev_ss = np.zeros(nsamp)
+    ev_ss = np.zeros(nsamp)
     gamma_ssi = np.zeros((nsamp, nfreqs, nreals, nskies, nloudest))
 
     # # one time calculations
@@ -1642,7 +1645,7 @@ def detect_lib(hdf_name, output_dir, npsrs, sigma, nskies, thresh=DEF_THRESH,
         # df_ss[nn] = np.sum(dp_ss[nn]>thresh)/(nreals*nskies)
         # df_bg[nn] = np.sum(dp_bg[nn]>thresh)/(nreals)
         df_ss[nn], df_bg[nn] = detfrac_of_reals(dp_ss[nn], dp_bg[nn], thresh)
-        # ev_ss[nn] = expval_of_ss(gamma_ssi)
+        ev_ss[nn] = expval_of_ss(gamma_ssi)
 
         if plot:
             fig = plot_sample_nn(fobs, hc_ss[nn], hc_bg[nn],
@@ -1662,7 +1665,7 @@ def detect_lib(hdf_name, output_dir, npsrs, sigma, nskies, thresh=DEF_THRESH,
     np.savez(output_dir+'/detstats.npz', dp_ss=dp_ss, dp_bg=dp_bg,
              df_ss=df_ss, df_bg=df_bg, snr_ss=snr_ss, snr_bg=snr_bg, gamma_ssi=gamma_ssi)
 
-    return dp_ss, dp_bg, df_ss, df_bg, snr_ss, snr_bg #, ev_ss
+    return dp_ss, dp_bg, df_ss, df_bg, snr_ss, snr_bg, ev_ss
 
 
 def _build_skies(nfreqs, nskies, nloudest):

### 2023-06-20 holodeck/detstats.py
Updated detect_lib to have option of returning dict, otherwise return nothing. Also has option to calculate and save gamma_ssi, which takes up huge amounts of space.

@@ -1538,7 +1538,8 @@ def detect_ss_pta(pulsars, cad, dur, fobs, dfobs, hc_ss, hc_bg,
 
 def detect_lib(hdf_name, output_dir, npsrs, sigma, nskies, thresh=DEF_THRESH,
                    dur=None, cad=None, dfobs=None, plot=True, debug=False,
-                   grid_path=GAMMA_RHO_GRID_PATH, snr_cython = True):
+                   grid_path=GAMMA_RHO_GRID_PATH, snr_cython = True, 
+                   save_ssi=False, ret_dict=False):
     """ Calculate detection statistics for an ss library output.
 
     Parameters
@@ -1559,6 +1560,12 @@ def detect_lib(hdf_name, output_dir, npsrs, sigma, nskies, thresh=DEF_THRESH,
         Whether or not to make and save plots.
     debug : Bool
         Whether to print info along the way.
+    grid_path : string
+        Path to snr interpolation grid
+    snr_cython : Bool
+        Whether to use cython interpolation for ss snr calculation.
+    save_ssi : Bool
+        Whether to store gamma_ssi in npz arrays
 
     Returns
     -------
@@ -1627,7 +1634,7 @@ def detect_lib(hdf_name, output_dir, npsrs, sigma, nskies, thresh=DEF_THRESH,
     df_ss = np.zeros(nsamp)
     df_bg = np.zeros(nsamp)
     ev_ss = np.zeros(nsamp)
-    gamma_ssi = np.zeros((nsamp, nfreqs, nreals, nskies, nloudest))
+    if save_ssi: gamma_ssi = np.zeros((nsamp, nfreqs, nreals, nskies, nloudest))
 
     # # one time calculations
     # Num = nfreqs * nloudest # number of single sources in a single strain realization (F*L)
@@ -1641,11 +1648,10 @@ def detect_lib(hdf_name, output_dir, npsrs, sigma, nskies, thresh=DEF_THRESH,
                                                 gamma_cython=True, snr_cython=snr_cython,
                                                 theta_ss=theta_ss, phi_ss=phi_ss, Phi0_ss=Phi0_ss,
[…]

### 2023-06-20 holodeck/detstats.py
Debugged and tested detect_lib with dict and ev returned, with and without gamma_ssi in detect_lib_ev.ipynb

@@ -1618,6 +1618,7 @@ def detect_lib(hdf_name, output_dir, npsrs, sigma, nskies, thresh=DEF_THRESH,
     phis = np.random.uniform(0, 2*np.pi, size = npsrs)
     thetas = np.random.uniform(np.pi/2, np.pi/2, size = npsrs)
     # sigmas = np.ones_like(phis)*sigma
+    if debug: print(f"{phis.shape=}, {thetas.shape=}, {dur=}, {cad=}, {sigma=}")
     psrs = hsim.sim_pta(timespan=dur/YR, cad=1/(cad/YR), sigma=sigma,
                     phi=phis, theta=thetas)
 
@@ -1650,7 +1651,7 @@ def detect_lib(hdf_name, output_dir, npsrs, sigma, nskies, thresh=DEF_THRESH,
                                                 iota_ss=iota_ss, psi_ss=psi_ss, grid_path=grid_path)
         dp_ss[nn,:,:], snr_ss[nn,...] = vals_ss[0], vals_ss[1],
         df_ss[nn], df_bg[nn] = detfrac_of_reals(dp_ss[nn], dp_bg[nn], thresh)
-        ev_ss[nn] = expval_of_ss(gamma_ssi)
+        ev_ss[nn] = expval_of_ss(vals_ss[2])
         if save_ssi: gamma_ssi[nn] = vals_ss[2]
 
         if plot:
@@ -1735,7 +1736,8 @@ def expval_of_ss(gamma_ssi,):
         Expected number of single source detection (dp_ss>thresh) averaged across all strain and sky realizations.
     
     """
-    nfreqs, nreals, nskies, nloudest = [*gamma_ssi.size]
+    print(f"{gamma_ssi.shape=}, {[*gamma_ssi.shape]}")
+    nfreqs, nreals, nskies, nloudest = [*gamma_ssi.shape]
     ev_ss = np.sum(gamma_ssi)/(nreals*nskies)
     return ev_ss
     # df_bg[nn] = np.sum(dp_bg[nn]>thresh)/(nreals)

### 2023-06-20 scripts/detect_lib_script.py
Checked detect_libe_script results, gamma_ssi and ev look good.

@@ -13,6 +13,8 @@ DEF_SIGMA = 1e-6
 
 DEF_NSKIES = 25
 DEF_THRESH = 0.5
+DEF_SNR_CYTHON = True
+DEF_SAVE_SSI = False
 
 GAMMA_RHO_GRID_PATH = '/Users/emigardiner/GWs/holodeck/output/rho_gamma_grids' # modify for system
 
@@ -42,8 +44,10 @@ def _setup_argparse():
                         help='produce plots for each simulation configuration')
     parser.add_argument('--debug', action='store_true', default=False,
                         help='print steps along the way')
-    parser.add_argument('--snr_cython', action='store_true', default=False,
+    parser.add_argument('--snr_cython', action='store_true', default=DEF_SNR_CYTHON,
                         help='Use cython for ss snr clculations')
+    parser.add_argument('--save_ssi', action='store_true', default=DEF_SAVE_SSI,
+                        help="Save 'gamma_ssi', the detprob of each single source.")
     
     args = parser.parse_args()
     return args
@@ -80,7 +84,7 @@ def main():
     ds.detect_lib(hdf_name, output_dir, args.npsrs, args.sigma, 
                         nskies=args.nskies, thresh=args.thresh, plot=args.plot, debug=args.debug,
                         dur=dur, cad=cad, dfobs=dfobs, grid_path=args.grid_path, 
-                        snr_cython=args.snr_cython)
+                        snr_cython=args.snr_cython, save_ssi=args.save_ssi)
   
 
 if __name__ == "__main__":

### 2023-06-20 ecg-notebooks/parameter_investigation/anatomy.py
Moved three model plotting functions to anatomy.py in ecg-notebooks, saved all threemodel figs for GW and 09A full shape.

@@ -1,0 +1,355 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import h5py
+
+
+from holodeck import plot, detstats
+import holodeck.single_sources as sings
+from holodeck.constants import YR, MSOL, MPC
+import holodeck as holo
+
+import hasasia.sim as hsim
+
+
+
+NPSRS = 50
+SIGMA = 3.55e-6
+NSKIES = 25
+THRESH = 0.5
+DUR = holo.librarian.DEF_PTA_DUR
+
+def detect_pspace_model(data, dur=DUR,
+                        npsrs=NPSRS, sigma=SIGMA, nskies=NSKIES, thresh=THRESH):
+    fobs_cents = data['fobs_cents']
+    hc_ss = data['hc_ss']
+    hc_bg = data['hc_bg']
+    dsdata = detstats.detect_pspace_model(fobs_cents, hc_ss, hc_bg, dur,
+                        npsrs, sigma, nskies, thresh)
+    return dsdata
+
+
+def draw_sample_text(fig, params, param_names, 
+                     xx=0.1, yy=-0.025, fontsize=10, color='k'):
+    text = ''
+    for pp, name in enumerate(param_names):
+        text = text+"'%s'=%.2e, " % (name, params[name])
+        # if pp == int(len(param_names)/2):
+        #     text = text+'\n'
+    fig.text(xx, yy, text, fontsize=fontsize, color=color, alpha=0.75)
+
+
+############################################################
+#### Draw hc vs par
+############################################################
+
+def draw_hc_vs_par(ax, xx_ss=None, yy_ss=None, xx_bg=None, yy_bg=None, color_ss='r', color_bg='k', fast_ss=True, 
+                   show_medians = False, show_ci=False, show_reals=True):
+    if show_reals:
+        if (xx_ss is not None) and (yy_ss is not None):
+            if fast_ss:
+                ax.scatter(xx_ss.flatten(), yy_ss.flatten(), marker='o', s=15, alpha=0.1, color=color_ss)
+            else:
+                colors = cm.rainbow(np.linspace(0,1,len(yy_ss[0])))
[…]

### 2023-06-20 holodeck/gravwaves.py
Add example of eccentric GWB calculation for illustris population in 'discrete_illustris' notebook

@@ -21,7 +21,7 @@ _CALC_MC_PARS = ['mass', 'sepa', 'dadt', 'scafa', 'eccen']
 
 class Grav_Waves:
 
-    def __init__(self, bin_evo, fobs_gw, nharms=30, nreals=100):
+    def __init__(self, bin_evo, fobs_gw, nharms=103, nreals=100):
         self.fobs_gw = fobs_gw
         self.nharms = nharms
         self.nreals = nreals

### 2023-06-20 holodeck/gravwaves.py
Include 'discrete_hardening_models' in test notebooks

@@ -179,7 +179,6 @@ def _gws_harmonics_at_evo_fobs(fobs_gw, dlnf, evo, harm_range, nreals, box_vol, 
     num_binaries = _lambda_fact * dlnf
 
     shape = (num_binaries.size, nreals)
-    # num_pois = np.random.poisson(num_binaries[:, np.newaxis], shape)
     num_pois = poisson_as_needed(num_binaries[:, np.newaxis] * np.ones(shape))
 
     # --- Calculate GW Signals
